---
title: "AI Spellcheck"
output: null
date: 12-27-2025
---

```{r}
#install.packages("ellmer")
library(ellmer)
library(tidyverse)
library(glue)
library(janitor)

Sys.setenv(GOOGLE_API_KEY = "")

```

#Check Gemini is working

```{r}
chat <- chat_google_gemini()

chat$chat("Tell me three jokes about journalists")
```

## Load Data
```{r echo = F}
precivil_complete <- read_csv("data/precivil_lynching_complete_12_22.csv") |> 
  mutate(index1 = as.numeric(str_remove(filename, "_.*\\.txt"))) |> 
  mutate(index2 = coalesce(index, index1)) |> 
  select(-c(index1,index, matches, attempted_y_n)) |> 
  rename(index = index2)

```

# new using html tags

```{r}
library(dplyr)
library(stringr)
library(purrr)
library(ellmer)

# 1. CLEANING (Keep it simple, leave the heavy lifting to Gemini)
processed_llm_test <- precivil_complete %>%
  mutate(original_text = gsub("\\s+", " ", article)) %>%
  group_by(index) %>%
  summarize(original_text = paste(na.omit(original_text), collapse = " "), .groups = "drop") %>%
  mutate(batch_id = ntile(row_number(), 9)) # 9 batches for 65 articles

# 2. THE CHAT OBJECT (Requesting Tags, not JSON)
chat_robust <- chat_google_gemini(
  system_prompt = "You are a data processing assistant. 
  For each article, you must provide the output wrapped in these specific tags:
  <article>
    <index>ID HERE</index>
    <original>TEXT HERE</original>
    <fixed>CORRECTED TEXT HERE</fixed>
  </article>
  
  Instructions:
  - Do not use JSON or Markdown tables.
  - Do not worry about escaping quotes. 
  - Simply place the text inside the tags.
  - Fix OCR errors, spelling, and spacing in the <fixed> section.
  - The article text provided here was extracted from newspaper page images through an imperfect OCR process. Do your best to correct any flaws introduced in this process, without changing meaning of the article. You should spellcheck the text and correct spelling errors, standardize capitalization, fix extraneous spaces, remove newline characters and random slashes, separate words that have obviously been concatenated in error, remove non alphabetic or standard punctuation characters. "
)

# 3. THE EXTRACTION FUNCTION (The "Secret Sauce")
parse_tagged_response <- function(raw_text) {
  # This uses regex to find everything between the tags
  indices <- str_match_all(raw_text, "<index>(.*?)</index>")[[1]][,2]
  originals <- str_match_all(raw_text, "<original>(.*?)</original>")[[1]][,2]
  fixed <- str_match_all(raw_text, "<fixed>(.*?)</fixed>")[[1]][,2]
  
  # Combine into a dataframe
  tibble(
    index = indices,
    original_text = originals,
    spellchecked_text = fixed
  )
}

# 4. THE BATCH PROCESSOR
process_single_batch_robust <- function(batch_data) {
  batch_input <- batch_data %>%
    mutate(formatted = paste0("ID: ", index, "\nCONTENT: ", original_text)) %>%
    pull(formatted) %>%
    paste(collapse = "\n\n---\n\n")
  
  message(paste("Processing Batch ID:", unique(batch_data$batch_id)))
  
  # Get response from Gemini
  response <- chat_robust$chat(paste("Process these articles:\n", batch_input))
  
  # Parse with our robust regex function
  return(parse_tagged_response(response))
}

# 5. EXECUTION
final_results <- processed_llm_test %>%
  group_split(batch_id) %>%
  map_dfr(~process_single_batch_robust(.x))

# Save results immediately
#write.csv(final_results, "lynching_ocr_corrected.csv", row.names = FALSE)
```

```{r}
library(dplyr)
library(stringr)
library(purrr)
library(tidyr)

# 1. IMPROVED PARSER: Processes article-by-article
parse_tagged_response <- function(raw_text) {
  
  # Step A: Extract each full <article> block
  # We use (?s) to allow the dot to match newlines
  blocks <- str_extract_all(raw_text, "(?s)<article>.*?</article>")[[1]]
  
  if (length(blocks) == 0) {
    message("Warning: No <article> blocks found in this batch.")
    return(tibble())
  }
  
  # Step B: Extract fields from each block individually
  map_dfr(blocks, function(block) {
    tibble(
      index = str_match(block, "<index>(.*?)</index>")[,2],
      original_text = str_match(block, "<original>(.*?)</original>")[,2],
      spellchecked_text = str_match(block, "<fixed>(.*?)</fixed>")[,2]
    )
  })
}

# 2. IMPROVED PROCESSOR: Adds a 'Safe' wrapper
process_single_batch_robust <- function(batch_data) {
  
  batch_input <- batch_data %>%
    mutate(formatted = paste0("ID: ", index, "\nCONTENT: ", original_text)) %>%
    pull(formatted) %>%
    paste(collapse = "\n\n---\n\n")
  
  message(paste("Processing Batch ID:", unique(batch_data$batch_id), "..."))
  
  # API call with error handling
  response <- tryCatch({
    chat_robust$chat(paste("Process these articles:\n", batch_input))
  }, error = function(e) {
    message("API Error: ", e$message)
    return(NULL)
  })
  
  if (is.null(response)) return(tibble())
  
  # Parse the response
  parsed_data <- parse_tagged_response(response)
  
  # Check if we lost any articles
  if (nrow(parsed_data) != nrow(batch_data)) {
    message(sprintf("Warning: Batch %s returned %s articles but %s were sent.", 
                    unique(batch_data$batch_id), nrow(parsed_data), nrow(batch_data)))
  }
  
  return(parsed_data)
}

# 5. EXECUTION
final_results <- processed_llm_test %>%
  group_split(batch_id) %>%
  map_dfr(~process_single_batch_robust(.x))

final_results2 <-
  final_results |> 
  distinct(index, .keep_all = TRUE) |> 
  mutate(index = as.numeric(index))

#missing 7 entries and the vermont entry is duplicated four times

```


### rejoin to df

```{r}
precivil_complete2 <- precivil_complete |> 
  left_join(final_results2, by=c("index"))


write_csv(precivil_complete2, "data/precivil_complete_dec_28.csv")
```

#-------------------------


# failed json Gemini spellcheck

```{r}
library(dplyr)
library(stringr)
library(purrr)
library(ellmer)
library(jsonlite)

# 1. CLEANING & PREPARATION
# Note: I removed the line that deletes all spaces/punctuation 
# because Gemini needs that context to fix OCR errors.
process_articles <- function(df) {
  df %>%
    mutate(
      sentence = gsub("\n", " ", article),
      sentence = gsub("\\s+", " ", sentence)
    ) %>%
    group_by(index) %>%
    summarize(
      original_text = paste(na.omit(sentence), collapse = " "),
      date = first(date),
      .groups = "drop"
    )
}

# Process the raw data
processed_llm_test <- process_articles(precivil_complete)

# 2. CREATE BATCHES (approx 7-9 articles per batch)
n_batches <- 9 
processed_llm_test <- processed_llm_test %>%
  mutate(batch_id = ntile(row_number(), n_batches))

# 3. DEFINE THE CHAT OBJECT
chat2 <- chat_google_gemini(
  system_prompt = "You are an expert in historical newspaper OCR correction. 
  You will receive a list of articles. For each, correct the OCR errors and spellcheck.
  Return the data as a JSON array of objects. 
  Each object MUST have: 'index', 'original_text', and 'spellchecked_text'.
  Do not include any markdown formatting or extra text.
  
  You are a JSON-only data generator. 
  For every article provided, return a JSON object in an array.
  
  CRITICAL RULES:
  1. Output MUST be valid JSON. 
  2. ESCAPE all internal double quotes within the text fields using a backslash (e.g., \\\"quote\\\").
  3. Do not include any conversational text, markdown code blocks, or explanations. 
  4. Return ONLY the raw JSON array.
  5. Text fields: 'index', 'original_text', 'spellchecked_text'."
)

# 4. EXECUTE BATCH PROCESSING
# This function sends one batch to Gemini
process_single_batch <- function(batch_data) {
  
  batch_input <- batch_data %>%
    mutate(formatted = paste0("INDEX: ", index, "\nTEXT: ", original_text)) %>%
    pull(formatted) %>%
    paste(collapse = "\n\n---NEXT ARTICLE---\n\n")
  
  message(paste("Processing Batch ID:", unique(batch_data$batch_id)))
  
  # Call Gemini
  response <- chat2$chat(paste(
    "Return the JSON array for these articles:",
    batch_input
  ))
  
  # --- ROBUST CLEANING STEP ---
  # 1. Remove Markdown code blocks
  clean_json <- gsub("```json|```", "", response)
  
  # 2. Extract only the part between the first [ and the last ]
  # (This ignores any text Gemini accidentally puts before or after the JSON)
  start_pos <- regexpr("\\[", clean_json)
  end_pos <- regexpr("\\](?=[^\\]]*$)", clean_json, perl = TRUE)
  
  if (start_pos == -1 || end_pos == -1) {
    warning("Could not find JSON array in response. Returning empty data.")
    return(data.frame())
  }
  
  clean_json <- substr(clean_json, start_pos, end_pos)
  
  # 3. Attempt to parse
  out <- tryCatch({
    fromJSON(clean_json)
  }, error = function(e) {
    message("JSON parsing failed for this batch. Check for unescaped quotes.")
    # Print the response to console so you can see exactly what broke
    cat("Problematic Response Snippet:", substr(clean_json, 1, 500), "...\n")
    return(data.frame()) 
  })
  
  return(out)
}

# 5. RUN THE LOOP
# This safely iterates through each batch and binds the results together
final_results <- processed_llm_test %>%
  group_split(batch_id) %>%
  map_dfr(~process_single_batch(.x))

# View final output
print(head(final_results))
```


### rejoin to df

```{r}
precivil_complete2 <- precivil_complete |> 
  left_join(final_results, by=c("index"))
```


#-------------------------------------------------------------------------

#---earlier attempts---------------------------------------------------------------------------
#-------------------------------------------------------------------------

```{r}
#65 articles
llm_df <- precivil_complete

llm_df <- llm_df |> 
  mutate(sentence = gsub("\n", " ", article)) |> 
  mutate(sentence = gsub("\\s+", " ", article))

```
```{r}
process_articles <- function(df) {
  # First group by article identifiers and concatenate sentences
  # Using filename as the identifier, but you might need to use other columns
  compiled_df <- df %>%
    group_by(index) %>%
    summarize(
      sentence = paste(na.omit(sentence), collapse = " "),
      index = first(index),
      date = first(date)
    )
  
  # Remove punctuation and spaces from the sentence column
  compiled_df <- compiled_df %>%
    mutate(sentence = str_replace_all(sentence, "[[:punct:][:space:]]", "")) |> 
    mutate(sentence = tolower(sentence))
  
  return(compiled_df)
}

# Make sure to load the necessary libraries
library(dplyr)
library(stringr)

# Apply the function to your dataframe
processed_llm_test <- process_articles(llm_df)
```
```{r}
# Prepare data with filenames and content
articles_for_analysis <- processed_llm_test %>%
  select(index, sentence) %>%
  mutate(article_data = paste("FILENAME:", index, "\nCONTENT:", sentence))

# Combine with clear separators
combined_text_with_filenames <- paste(articles_for_analysis$article_data, 
                                    collapse = "\n\n---ARTICLE SEPARATOR---\n\n")
```

# LLM Prompt


```{r}
chat2 <- chat_google_gemini(
system_prompt = "You are an expert in classifying historical newspaper articles about lynchings in the United States before the Civil War. You always follow instructions.
 Please do the following:
 -- The article text provided here was extracted from newspaper page images through an imperfect OCR process. Do your best to correct any flaws introduced in this process, without changing meaning of the article. You should spellcheck the text and correct spelling errors, standardize capitalization, fix extraneous spaces, remove newline characters and random slashes, separate words that have obviously been concatenated in error, remove non alphabetic or standard punctuation characters. Of special importance is to correct any errors that will prevent the json from being parsed correctly later.
 Format your response as a dataframe with these exact fields:
 {
     \"index\": \"string, unchanged from input\",
      \"sentence\": \"string, original text of sentence, unchanged from input\",
     \"spellchecked_text\": \"string, corrected spelling of sentence\",

 }

 Important formatting rules:
 - Use double quotes for all strings
 - No line breaks in text fields
 - No trailing commas
 - No comments or additional text
 - No markdown formatting
 - Escape all quotes within text using single quotes
 - Remove any \\r or \\n characters from text"
)

```





```{r}
library(dplyr)
library(stringr)
library(jsonlite)

# Split into individual articles
articles <- str_split(combined_text_with_filenames, "FILENAME: ")[[1]]
articles <- articles[articles != ""]  # Remove empty elements

cat("Total articles:", length(articles), "\n")

# Create batches of 10 articles each (62 articles = 7 batches)
batch_size <- 10
batches <- split(articles, ceiling(seq_along(articles) / batch_size))

cat("Total batches:", length(batches), "\n")

# Set up chat with max tokens
chat2 <- chat_google_gemini(
  model = "gemini-2.5-flash",
  params = params(max_tokens = 8192)
)

# Process each batch
all_results <- list()

for(i in seq_along(batches)) {
  cat("Processing batch", i, "of", length(batches), "...\n")
  
  # Combine batch back into text
  batch_text <- paste("FILENAME:", batches[[i]], collapse = "\n\n")
  
  # Your spellcheck prompt
  prompt <- paste0(
    "You are an expert in classifying historical newspaper articles about lynchings in the United States before the Civil War. You always follow instructions.
 Please do the following:
 -- The article text provided here was extracted from newspaper page images through an imperfect OCR process. Do your best to correct any flaws introduced in this process, without changing meaning of the article. You should spellcheck the text and correct spelling errors, standardize capitalization, fix extraneous spaces, remove newline characters and random slashes, separate words that have obviously been concatenated in error, remove non alphabetic or standard punctuation characters. Of special importance is to correct any errors that will prevent the json from being parsed correctly later.
 Format your response as a dataframe with these exact fields:
 {
     \"index\": \"string, unchanged from input\",
      \"sentence\": \"string, original text of sentence, unchanged from input\",
     \"spellchecked_text\": \"string, corrected spelling of sentence\",

 }

 Important formatting rules:
 - Use double quotes for all strings
 - No line breaks in text fields
 - No trailing commas
 - No comments or additional text
 - No markdown formatting
 - Escape all quotes within text using single quotes
 - Remove any \\r or \\n characters from text",
    batch_text
  )
  
  all_results[[i]] <- chat2$chat(prompt)
  
  Sys.sleep(2)  # Rate limiting - be respectful to API
}
library(stringr)
library(jsonlite)
library(dplyr)

# Function to extract JSON objects from incomplete/truncated responses
extract_json_objects <- function(response) {
  # Remove markdown code blocks if present
  response <- gsub("```json\\n|```\\n|```", "", response)
  response <- trimws(response)
  
  # Try to parse as complete JSON first
  result <- tryCatch({
    parsed <- fromJSON(response, simplifyDataFrame = TRUE)
    if (is.data.frame(parsed)) {
      return(parsed)
    } else if (is.list(parsed)) {
      return(bind_rows(lapply(parsed, as.data.frame)))
    }
  }, error = function(e) NULL)
  
  if (!is.null(result)) return(result)
  
  # If that fails, try to extract individual complete objects
  # More precise regex to find complete JSON objects
  pattern <- '\\{\\s*"index"\\s*:\\s*"[^"]*"\\s*,\\s*"sentence"\\s*:\\s*"[^"]*"\\s*,\\s*"spellchecked_text"\\s*:\\s*"[^"]*"\\s*\\}'
  matches <- str_extract_all(response, pattern)[[1]]
  
  if (length(matches) == 0) {
    # Try even more lenient pattern for multiline strings
    pattern2 <- '\\{[^}]*?"index"[^}]*?"sentence"[^}]*?"spellchecked_text"[^}]*?\\}'
    matches <- str_extract_all(response, pattern2)[[1]]
  }
  
  # Parse each matched object
  result_list <- lapply(matches, function(obj) {
    tryCatch({
      parsed <- fromJSON(obj)
      data.frame(
        index = as.character(parsed$index),
        sentence = as.character(parsed$sentence),
        spellchecked_text = as.character(parsed$spellchecked_text),
        stringsAsFactors = FALSE
      )
    }, error = function(e) {
      message("Failed to parse: ", substr(obj, 1, 100))
      return(NULL)
    })
  })
  
  # Filter out NULLs and combine
  result_list <- result_list[!sapply(result_list, is.null)]
  
  if (length(result_list) > 0) {
    return(bind_rows(result_list))
  } else {
    return(data.frame(
      index = character(0),
      sentence = character(0),
      spellchecked_text = character(0)
    ))
  }
}

# Re-process your results
df_list <- lapply(all_results, function(result) {
  tryCatch({
    extract_json_objects(result)
  }, error = function(e) {
    message("Error processing batch: ", e$message)
    return(NULL)
  })
})

df_spellchecked <- bind_rows(df_list)
cat("Total rows processed:", nrow(df_spellchecked), "\n")
head(df_spellchecked)
```


# Send to the LLM
```{r}
spellchecked <- chat2$chat(combined_text_with_filenames)


```




```{r}
library(jsonlite)

process_llm_response_to_df <- function(response) {
  # If response is a vector, collapse into single string
  if (length(response) > 1) {
    response <- paste(response, collapse = "\n")
  }
  
  # Remove markdown code block markers if present
  response <- gsub("```json\\n|```\\n|```", "", response)
  
  # Try to parse as JSON
  tryCatch({
    # Parse the JSON
    parsed <- fromJSON(response, simplifyDataFrame = TRUE)
    
    # If it's already a dataframe, return it
    if (is.data.frame(parsed)) {
      return(parsed)
    }
    
    # If it's a list, convert to dataframe
    if (is.list(parsed)) {
      df <- as.data.frame(do.call(rbind, lapply(parsed, function(x) {
        data.frame(
          index = as.character(x$index),
          sentence = as.character(x$sentence),
          spellchecked_text = as.character(x$spellchecked_text),
          stringsAsFactors = FALSE
        )
      })))
      return(df)
    }
    
  }, error = function(e) {
    message("Error parsing JSON: ", e$message)
    # Return empty dataframe with correct structure
    return(data.frame(
      index = character(0),
      sentence = character(0),
      spellchecked_text = character(0),
      stringsAsFactors = FALSE
    ))
  })
}

# Process your data
df_spellchecked <- process_llm_response_to_df(spellchecked)


df_spellchecked <- df_spellchecked |> 
  mutate(index= as.numeric(index))

```

```{r}
# Process each response individually
df_list <- lapply(spellchecked, function(x) {
  tryCatch({
    process_llm_response_to_df(x)
  }, error = function(e) {
    message("Skipping invalid response: ", substr(x, 1, 100))
    return(NULL)
  })
})

# Combine all valid dataframes
df_spellchecked <- bind_rows(df_list)
```


#-------------------------------------


ver 10-20-2025
```{r}
chat2 <- chat_gemini(
  system_prompt = "You are an academic researcher performing context analysis on selected news articles. Analyze each article and classify it based on these criteria,
  
  1. POLITICS classification: Classify an article as 'politics' if it contains two or more mentions of electoral politics.

'Electoral politics' is defined as:
- Upcoming elections (primaries, general elections, special elections)
- Election results and outcomes
- Political campaigns and campaign strategies
- Candidate announcements, debates, or endorsements
- Election polling and predictions
- Political party activities related to elections
- **Terms**: Classify an article as 'politics' if it contains the issues above and also mentions terms such as 'vote' or 'precinct' or 'primary results'

EXCLUSIONS:
- Policy debates or legislation without electoral context → classify as 'other'
- Appointed officials or bureaucratic actions → not electoral politics unless tied to an election


2. VOTER TURNOUT Classification: 
Classify an article as 'voter_turnout' if it contains two or more mentions of voter participation or turnout-related issues.

'Voter turnout' is defined as:
- Voter registration drives or registration statistics
- Early voting, absentee voting, or mail-in ballot usage
- Voter turnout rates or participation levels
- Get-out-the-vote (GOTV) campaigns or efforts
- Barriers to voting or voter access issues (polling locations, ID requirements, wait times)
- Demographic breakdowns of voter participation
- Voter suppression
- **Terms**: Classify an article as 'voter_turnout' if it contains the issues above and also mentions terms such as 'ballot box'

  3. ECONOMY classification: Classify an article as 'economy' if it contains two or more mentions of macroeconomic issues or economic policy.

'Economy' is defined as:
- Macroeconomic indicators (GDP, inflation, unemployment rates, economic growth)
- Monetary policy (Federal Reserve actions, interest rates)
- Trade policy (tariffs, trade agreements, import/export regulations)
- Economic forecasts or recession/expansion predictions
- Currency markets and exchange rates
- National or global economic conditions and trends

EXCLUSIONS:
- Individual company financial performance → classify as 'business'
- Industry-specific trends without macroeconomic context → classify as 'business'
- Stock market movements alone (unless tied to broader economic analysis) → classify as 'business'
- Fiscal policy debates over taxation and spending in Congress  → classify as 'politics'

  4. BUSINESS classification: Classify an article as 'business' if it contains two or more mentions of business issues.
  
BUSINESS ISSUES is defined as articles that primarily discuss:
- **Companies**: Corporate entities, their operations, or strategic decisions
- **Executives**: Business leaders, appointments, departures, or leadership actions
- **Business Activity**: Including but not limited to:
  - Product launches and announcements
  - Marketing campaigns
  - Regulatory compliance matters
  - Mergers and acquisitions
  - Partnerships and collaborations
  - Corporate strategy changes
- **Financial Dealings**: Earnings, investments, funding, stock performance, financial results
- **Industry**: Sector trends, market analysis, competitive dynamics
- **Terms**: Classify an article as 'business' if it contains the issues above and also mentions such terms as 'private enterprise', 'company', 'firm', 'business leaders', 'businessmen',

EXCLUSIONS:
- Labor issues (strikes, unionization, worker disputes) should be classified separately as LABOR, not business issues
- Articles that only mention a company tangentially (e.g., 'Apple CEO attended a charity gala') without discussing business matters

  5.LABOR CLASSIFICATION: Classify an article as 'labor' if it contains two or more mentions of organized labor issues.

'Organized labor' is defined as:
- Strikes or work stoppages
- Unionization efforts or union elections
- Worker disputes involving unions or collective bargaining
- Public policy issues focusing on organized labor (e.g., labor law reforms, right-to-work legislation)
- Statements or actions by union leaders or labor organizations

- **Terms**: Classify an article as 'labor' if it contains the issues above and also mentions terms such as 'AFL' and or 'CIO' (Congress of Industrial Organizations),

IMPORTANT:
- The threshold is two or more mentions to avoid misclassifying articles that only reference labor tangentially
- General workplace issues (individual employee disputes, non-union worker complaints, HR policies) do NOT count as organized labor unless unions are involved

6. OVERLAPPING CLASSIFICATIONS:
An article should receive multiple classifications if it meets the threshold 
(two or more mentions) for more than one category.

- Assign all applicable classifications separated by semicolons
- List classifications in alphabetical order for consistency
- Example: An article with 3 mentions of electoral politics AND 2 mentions 
  of business issues would be classified as 'business;politics'


COUNTING METHODOLOGY:
- Count distinct mentions, not repeated references to the same fact
- A single sentence can contain multiple mentions if it discusses 
  multiple distinct aspects of a category
- When in doubt, count conservatively


EDGE CASES:
- If an article has exactly 2 mentions but you're uncertain if they're 
  distinct enough, err on the side of inclusion
- If an article discusses the same election/company/policy throughout, 
  count each substantive new point, not repetitions
- Introductory context-setting doesn't count toward thresholds

**Terms**: An article must contain BOTH the substantive issues described 
above AND at least one of these terms to qualify for this classification.

IMPORTANT: Each category must independently meet its own threshold of 
two or more mentions to be included in the classification.
  
  7. If an article meets none of the criteria, classify it as 'other'
  
 REQUIRED OUTPUT FORMAT:
Return ONLY a CSV with exactly THREE columns and NO additional text:

Column 1: new_filename (the article identifier)
Column 2: category (one or more classifications separated by semicolons)
Column 3: explanation (brief justification citing specific counts)

Rules:
- Include a header row: new_filename,category,explanation
- No quotes around values unless they contain commas
- Multiple categories in alphabetical order separated by semicolons
- Explanation must specify the count for each assigned category
- Maximum 15 words per explanation

Examples:
article1.txt,politics,`3 mentions of election campaigns, 2 candidate endorsements`
article2.txt,business;economy,`4 company strategy mentions, 2 GDP references'
article3.txt,other, `No categories met 2-mention threshold` 
  No additional text, explanations, or summary counts.")


```
  
# new prompt 10-21-2025

```{r}
chat2 <- chat_gemini(
  system_prompt = "You are an academic researcher performing content analysis on newspaper articles. 
Your task is to classify each article based on specific criteria defined below.

COUNTING METHODOLOGY:
- Count distinct mentions, not repeated references to the same fact
- A single sentence can contain multiple mentions if it discusses multiple distinct aspects of a category
- When in doubt, count conservatively
- Introductory context-setting does not count toward thresholds
- Each substantive new point counts; mere repetition does not

=============================================================================
CLASSIFICATION CATEGORIES (alphabetically)
=============================================================================

1. BUSINESS CLASSIFICATION

Classify an article as 'business' if it contains two or more mentions of business issues.

Business issues include articles that primarily discuss:
- Companies: Corporate entities, their operations, or strategic decisions
- Executives: Business leaders, appointments, departures, or leadership actions
- Business Activity:
  * Product launches and announcements
  * Marketing campaigns
  * Regulatory compliance matters
  * Mergers and acquisitions
  * Partnerships and collaborations
  * Corporate strategy changes
- Financial Dealings: Earnings, investments, funding, stock performance, financial results
- Industry: Sector trends, market analysis, competitive dynamics

**Indicator Terms**: 'private enterprise', 'company', 'firm', 'business leaders', 'businessmen', 'corporation', 'CEO', 'revenue'

EXCLUSIONS:
- Labor issues (strikes, unionization, worker disputes) → classify as 'labor'
- Articles that only mention a company tangentially without discussing business matters
- Macroeconomic policy discussions → classify as 'economy'

---

2. ECONOMY CLASSIFICATION

Classify an article as 'economy' if it contains two or more mentions of macroeconomic issues or economic policy.

Economy issues include:
- Macroeconomic indicators: GDP, inflation, unemployment rates, economic growth
- Monetary policy: Federal Reserve actions, interest rates, money supply
- Trade policy: Tariffs, trade agreements, import/export regulations
- Economic forecasts or recession/expansion predictions
- Currency markets and exchange rates
- National or global economic conditions and trends
- Economic crises or systemic economic issues

**Indicator Terms**: 'GDP', 'inflation', 'Federal Reserve', 'interest rates', 'recession', 'economic growth', 'unemployment rate'

EXCLUSIONS:
- Individual company financial performance → classify as 'business'
- Industry-specific trends without macroeconomic context → classify as 'business'
- Stock market movements alone (unless tied to broader economic analysis) → classify as 'business'
- Fiscal policy debates over taxation and spending in Congress → classify as 'politics'

---

3. LABOR CLASSIFICATION

Classify an article as 'labor' if it contains two or more mentions of organized labor issues.

Organized labor issues include:
- Strikes or work stoppages
- Unionization efforts or union elections
- Worker disputes involving unions or collective bargaining
- Labor contract negotiations
- Public policy issues focusing on organized labor (e.g., labor law reforms, right-to-work legislation)
- Statements or actions by union leaders or labor organizations
- Picket lines, labor protests, or organizing campaigns

**Indicator Terms**: 'union', 'strike', 'collective bargaining', 'picket line', 'AFL-CIO', 'labor organization', 'workers' rights', 'unionization'

Note: `CIO` refers to Congress of Industrial Organizations (historical labor federation, now part of AFL-CIO)

EXCLUSIONS:
- General workplace issues (individual employee disputes, non-union worker complaints, HR policies) without union involvement
- Executive/management labor issues → classify as 'business'

---

4. POLITICS CLASSIFICATION

Classify an article as 'politics' if it contains two or more mentions of electoral politics.

Electoral politics includes:
- Upcoming elections (primaries, general elections, special elections, referendums)
- Election results and outcomes
- Political campaigns and campaign strategies
- Candidate announcements, debates, or endorsements
- Election polling and predictions
- Political party activities related to elections
- Voter mobilization efforts tied to specific elections
- Campaign finance in electoral context

**Indicator Terms**: 'vote', 'election', 'candidate', 'primary', 'ballot', 'campaign', 'precinct', 'primary results', 'polling'

EXCLUSIONS:
- Policy debates or legislation without electoral context → classify as 'other'
- Appointed officials or bureaucratic actions → not electoral politics unless tied to an election
- Governance and policy implementation → classify as 'other'

---

5. VOTER TURNOUT CLASSIFICATION

Classify an article as 'voter_turnout' if it contains two or more mentions of voter participation or turnout-related issues.

Voter turnout issues include:
- Voter registration drives or registration statistics
- Early voting, absentee voting, or mail-in ballot usage
- Voter turnout rates or participation levels
- Get-out-the-vote (GOTV) campaigns or efforts
- Barriers to voting or voter access issues (polling locations, ID requirements, wait times)
- Voter suppression allegations or concerns
- Demographic breakdowns of voter participation
- Voting infrastructure (polling places, voting machines, ballot access)

**Indicator Terms**: 'ballot box', 'voter registration', 'turnout rate', 'early voting', 'absentee ballot', 'polling place', 'voter access'

=============================================================================
OVERLAPPING CLASSIFICATIONS
=============================================================================

An article may qualify for multiple classifications if it independently meets 
the two-mention threshold for more than one category.

PROCESS:
a) Evaluate the article against each classification independently
b) Count mentions for each category separately
c) Assign ALL categories that meet the threshold (2+ mentions)
d) List multiple classifications alphabetically, separated by semicolons

EXAMPLES:
- Article discusses 3 campaign events + 2 corporate mergers → 'business;politics'
- Article discusses 4 election issues + 3 GDP mentions + 2 strikes → 'economy;labor;politics'
- Article discusses 1 election + 5 business issues → 'business' (politics doesn't meet threshold)
- Article discusses 3 voter registration drives + 2 election debates → 'politics;voter_turnout'

IMPORTANT: Do not blend counts across categories. Each category stands alone.

=============================================================================
EDGE CASES
=============================================================================

- If an article has exactly 2 mentions but you're uncertain if they're distinct enough, err on the side of inclusion
- If an article discusses the same election/company/policy throughout, count each substantive new point, not repetitions
- If multiple categories could apply to the same content, count it for each relevant category independently

=============================================================================
DEFAULT CLASSIFICATION
=============================================================================

If an article meets none of the criteria above (does not reach the two-mention 
threshold for any category), classify it as 'other'.

=============================================================================
REQUIRED OUTPUT FORMAT
=============================================================================

Return ONLY a CSV with exactly THREE columns and NO additional text:

Column 1: new_filename (the article identifier)
Column 2: category (one or more classifications separated by semicolons)
Column 3: explanation (brief justification citing specific counts)

FORMATTING RULES:
- Include a header row: new_filename,category,explanation
- No quotes around values unless they contain commas
- Multiple categories in alphabetical order separated by semicolons (no spaces)
- Explanation must specify the count for each assigned category
- Maximum 20 words per explanation
- Do not include any summary statistics, commentary, or additional text

EXAMPLES:

new_filename,category,explanation
article1.txt,politics,3 mentions of election campaigns and 2 candidate endorsements
article2.txt,business;economy,4 company strategy mentions and 2 GDP references
article3.txt,voter_turnout,2 voter registration drives and 3 early voting statistics
article4.txt,labor;politics,3 strike references and 2 election campaign mentions
article5.txt,other,No categories met 2-mention threshold

=============================================================================

Begin classification now.")
```




# Send to the LLM
```{r}
politics <- chat2$chat(combined_text_with_filenames)
```

#Process results

```{r}
process_llm_response_to_df <- function(response) {
  # Extract lines
  lines <- strsplit(response, "\n")[[1]]
  
  # Remove markdown code block markers if present
  lines <- lines[!grepl("^```", lines)]
  
  # Initialize vectors for data
  filenames <- c()
  categories <- c()
  explanations <- c()
  
  # Flag to track if we're processing data (after header)
  header_found <- FALSE
  
  for (line in lines) {
    # Skip empty lines
    if (trimws(line) == "") next
    
    # Skip row numbers or other artifacts (lines with asterisks)
    if (grepl("\\*\\*", line)) next
    
    # Check if this is the header line
    if (grepl("filename.*category.*explanation", line, ignore.case = TRUE)) {
      header_found <- TRUE
      next
    }
    
    # Process data lines (only after header is found)
    if (header_found || grep("batch.*\\.txt", line)) {
      # Parse each row with better handling of commas within explanations
      parts <- strsplit(line, ",\\s*", perl = TRUE, fixed = FALSE)[[1]]
      
      if (length(parts) >= 1) {
        filename <- parts[1]
        category <- ifelse(length(parts) >= 2, parts[2], "")
        
        # Handle the explanation (everything after second comma)
        if (length(parts) >= 3) {
          explanation <- paste(parts[3:length(parts)], collapse = ", ")
        } else {
          explanation <- NA
        }
        
        filenames <- c(filenames, filename)
        categories <- c(categories, category)
        explanations <- c(explanations, explanation)
      }
    }
  }
  
  # Create and return the dataframe
  if (length(filenames) > 0) {
    data.frame(
      filename = filenames,
      category = categories,
      explanation = explanations,
      stringsAsFactors = FALSE
    )
  } else {
    # Return empty dataframe with correct structure if no data found
    data.frame(
      filename = character(0),
      category = character(0),
      explanation = character(0),
      stringsAsFactors = FALSE
    )
  }
}
lat_1956_results <- process_llm_response_to_df(politics)
write.csv(lat_1956_results, "./output/ai_classified_lat_1956_oct_20_2025.csv")
```


count the results
```{r}
lat_1956_results |> 
  count(category) |> 
  arrange(desc(n))

```

#Compare the results
```{r}
july24 <- rio::import("https://docs.google.com/spreadsheets/d/1Y-IWU_AgbyE-2G2jyT15MIO2jtv_Qpgi1imRwKTjPHQ/edit?gid=1373544130#gid=1373544130", which = "July24_results", header = TRUE) |> 
  clean_names()

oct20 <- rio::import("https://docs.google.com/spreadsheets/d/1Y-IWU_AgbyE-2G2jyT15MIO2jtv_Qpgi1imRwKTjPHQ/edit?gid=1553901874#gid=1553901874", which = "Oct20_results", header = TRUE) |> 
  clean_names() 


```


### Process to join
```{r}
july24_clean <- july24 |> 
  select(filename, category, explanation, correct, human_comments) |> 
  rename(july24_category = category, july24_explanation = explanation, july24_correct = correct, july24_human_comments = human_comments)
oct20_clean <- oct20 |> 
  select(filename, category, explanation)

combo <- oct20_clean |> 
  right_join(july24_clean, by ="filename")

#split crime1 into three new columns
combo <- separate(data = combo, col = category, into = c("category1", "category2"), sep = ";", extra = "merge", fill = "right")
combo <- separate(data = combo, col = july24_category, into = c("july24_category1", "july24_category2"), sep = ";", extra = "merge", fill = "right")
```

### Analyze differences
```{r}
diffs <- combo |> 
  mutate(changed = case_when(
    category1 == july24_category1 ~ "same",
    category1 != july24_category1 ~ "different",
     category2 == july24_category2 ~ "same",
    category2 != july24_category2 ~ "different",
    .default = NA_character_
  )) |> 
  select(changed, category1, july24_category1, category2, july24_category2, july24_correct, july24_human_comments, everything())

write_csv(diffs, "./output/oct20_july24_diffs.csv")
```


### Count all category1 and category2 together
```{r}

current_counts <- combo |>
  select(filename, category1, category2) |>
  pivot_longer(cols = c(category1, category2), 
               names_to = "category_type", 
               values_to = "category") |>
  filter(!is.na(category)) |>
  count(category, name = "current_count")

# Count all july24_category1 and july24_category2 together
july24_counts <- combo |>
  select(filename, july24_category1, july24_category2) |>
  pivot_longer(cols = c(july24_category1, july24_category2), 
               names_to = "category_type", 
               values_to = "category") |>
  filter(!is.na(category)) |>
  count(category, name = "july24_count")

# Combine them side by side
comparison <- full_join(current_counts, july24_counts, by = "category") |>
  replace_na(list(current_count = 0, july24_count = 0)) |>
  arrange(desc(current_count))

comparison





```

```{r}
diffs |> 
  count(changed)


```

### Tally the categories
```{r}
left_join(
  diffs |> count(category, name = "category_count"),
  diffs |> count(july24_category, name = "july24_count"),
  by = c("category" = "july24_category")
)

```



#Notes

Prompt guidance from Claude.ai
Suggested Prompt Structure
You are analyzing news articles to identify and categorize business issues.

BUSINESS ISSUES are defined as articles that primarily discuss:
- **Companies**: Corporate entities, their operations, or strategic decisions
- **Executives**: Business leaders, appointments, departures, or leadership actions
- **Business Activity**: Including but not limited to:
  - Product launches and announcements
  - Marketing campaigns
  - Regulatory compliance matters
  - Mergers and acquisitions
  - Partnerships and collaborations
  - Corporate strategy changes
- **Financial Dealings**: Earnings, investments, funding, stock performance, financial results
- **Industry**: Sector trends, market analysis, competitive dynamics

EXCLUSIONS:
- Labor issues (strikes, unionization, worker disputes) should be classified separately as LABOR, not business issues
- Articles that only mention a company tangentially (e.g., "Apple CEO attended a charity gala") without discussing business matters

TASK:
For each article, determine:
1. Does it contain a business issue? (YES/NO)
2. If YES, what is the primary category? (Company, Executive, Business Activity, Financial, Industry)
3. Provide a brief justification (1 sentence)



#old prompt


#used this prompt
```{r}
chat2 <- chat_gemini(
  system_prompt = "You are an academic researcher performing context analysis on selected Newsweek articles. Analyze each article and classify it based on these criteria:
  
  1. If an article contains two or more mentions of an upcoming election or electoral politics, classify it as 'politics'
  2. If an article contains two or more mentions of vote or voter registration, classify it as 'voter_turnout'
  3. If an article contains two or more mentions the economic developments, classify it as 'economy'
  4. If an article contains two or more mentions of business issues, classify it as 'business'
  5. If an article meets multiple criteria, assign multiple categories separated by semicolons
  6. If an article meets none of the criteria, classify it as 'other'
  
  Return ONLY a CSV-formatted result with exactly THREE (3) columns:
  filename, category, explanation
  
  For example:
  filename, category, explanation
  article1.txt,politics, warning about upcoming election
  article2.txt,democrat_critique, these words were critical of democrats: xyz
  article3.txt,business;politics, business mentioned three times voter turnout twice
  article4.txt,other
  
  No additional text, explanations, or summary counts."
)

# Send to the LLM
politics <- chat2$chat(combined_text_with_filenames)
```
