---
title: "Antebellum Lynching Analysis 2025"
author: "Rob Wells"
date: "2026-01-05"
output: html_document
---

```{r}
#install.packages("udpipe")
#install.packages('DT')
#install.packages("arrow")
#install.packages("spacyr")
library(arrow)
library(jsonlite)
library(janitor)
library(lubridate)
library(quanteda)
library(tidytext)
library(tidyverse)
library(udpipe)
library(DT)
library(spacyr)
library(stringr)

```

#Import 444 cleaned articles
```{r}
precivil_complete <- read_csv("../data/precivil_complete_jan12.csv") 



```


#Chart - Results by year
```{r}
#Here is a chart, Figure 1, that describes lynching search results by year
precivil_complete %>% 
   group_by(year) %>% 
  count(year) %>% 
#Sandwich it onto a simple ggplot
  ggplot(aes(x = year, y = n, fill = n)) +
  geom_col(position = "dodge") + 
  theme(legend.position = "none") +
  scale_x_continuous(labels = c(seq(1825, 1865, 5)), breaks = seq(1825, 1865, 5)) +
   # scale_y_continuous(breaks = seq(0, 15, 1)) +
  labs(title = "Precivil War Lynching Coverage By Year, 1835-1861", 
       subtitle = "Count of Lynching Stories Examined",
       caption = "n=491  articles. Graphic by Rob Wells, 1/10/2026",
       y="Count of Articles",
       x="Year")

ggsave("../output/precivil_coverage_year_1_3_2026.png", device = "png",width=9,height=6, dpi=800)
```


year total descending
```{r}
precivil_complete  %>% 
count(year) %>% 
  mutate(pct = formattable::percent(n/sum(n,1))) |> 
  arrange(desc(pct))


```

newspapers total descending
```{r}
newspapers <- precivil_complete  %>% 
count(newspaper_name_clean) %>% 
  arrange(desc(n))
newspapers

```


# Word count metrics
```{r}
#average story length is 449 words, but some articles ran up to 1,618 words

precivil_complete |> 
  summarize(avg = mean(word_count, na.rm=TRUE),
            median = median(word_count, na.rm=TRUE),
            min = min(word_count, na.rm=TRUE),
            max = max(word_count, na.rm=TRUE),
            sd = sd(word_count, na.rm=TRUE))


```


### Count number of lynch search terms in an article
then filter articles with one term for close filtering to that sentence

```{r}


# 1. Define the patterns (as provided in your guide)
lynching_patterns <- c(
# Matches lynch, lynched, lynching, lynchings, lynches
  "\\blynch(ed|ing|ings|es)?\\b", 
  
  # Matches hung, hang, hanged, hanging, hangings
  "\\b(hung|hanging)\\b.{1,20}\\bmob\\b", 
  
  # Matches vigilance, vigilant, vigilante, vigilantes
  "\\bvigilant(e|es|ance|ism)?\\b",
  "lynchings?\\W+of\\W+(the\\W+)?(([[:alnum:]]+\\W+){1,2})?colored",
  "(murderer|fiend|desperado|brute)\\W+(([[:alnum:]]+\\W+){1,2})?lynch(ed|es|ing)?(\\W+|$)",
  "coloreds?\\W+(([[:alnum:]]+\\W+){1,2})?((was|were)\\W+)?lynch(ed|es|ing)?(\\W+|$)",
  "lynchings?\\W+of\\W+(the\\W+)?(([[:alnum:]]+\\W+){1,2})?negro",
  "mob\\W+(([[:alnum:]]+\\W+){1,2})?(hung|hang(ed|ings?|s)|lynch(ed|es|ing)?)",
  "negro(e?s)?\\W+(([[:alnum:]]+\\W+){1,2})?((was|were)\\W+)?lynch(ed|es|ing)?(\\W+|$)"
)

combined_pattern <- paste(lynching_patterns, collapse = "|")

# 2. Process the dataframe and count occurrences
Xprecivil_complete <- precivil_complete %>%
  # Cleaning steps for better matching
  mutate(x_article = str_to_lower(spellchecked)) %>%
  # COUNT the number of times any of the patterns appear
  mutate(search_count = str_count(x_article, regex(combined_pattern, ignore_case = TRUE)))



```



```{r}
fatboys <- Xprecivil_complete |> 
  filter(word_count > 311) |> 
  filter(search_count < 2)


```

Validate
```{r}
wrapped_text <- strwrap(as.character(fatboys$spellchecked[1]), width = 60)
writeLines(wrapped_text, "temp_text.txt")
file.edit("temp_text.txt")
```

# Extract Lynch Sentences - 
```{r}


# 1. Define the patterns (as provided in your guide)
lynching_patterns <- c(
# Matches lynch, lynched, lynching, lynchings, lynches
  "\\blynch(ed|ing|ings|es)?\\b", 
  
  # Matches hung, hang, hanged, hanging, hangings
  "\\b(hung|hanging)\\b.{1,20}\\bmob\\b", 
  
  # Matches vigilance, vigilant, vigilante, vigilantes
  "\\bvigilant(e|es|ance|ism)?\\b",
  "lynchings?\\W+of\\W+(the\\W+)?(([[:alnum:]]+\\W+){1,2})?colored",
  "(murderer|fiend|desperado|brute)\\W+(([[:alnum:]]+\\W+){1,2})?lynch(ed|es|ing)?(\\W+|$)",
  "coloreds?\\W+(([[:alnum:]]+\\W+){1,2})?((was|were)\\W+)?lynch(ed|es|ing)?(\\W+|$)",
  "lynchings?\\W+of\\W+(the\\W+)?(([[:alnum:]]+\\W+){1,2})?negro",
  "mob\\W+(([[:alnum:]]+\\W+){1,2})?(hung|hang(ed|ings?|s)|lynch(ed|es|ing)?)",
  "negro(e?s)?\\W+(([[:alnum:]]+\\W+){1,2})?((was|were)\\W+)?lynch(ed|es|ing)?(\\W+|$)"
)

combined_pattern <- paste(lynching_patterns, collapse = "|")

# Tokenize into sentences
sentences_df <- precivil_complete %>%
  unnest_tokens(sentence, spellchecked, token = "sentences") %>%
  group_by(index) %>% # Assuming you have a unique ID per article
  mutate(sentence_id = row_number())

# Identify sentences containing your terms
matches <- sentences_df %>%
  filter(str_detect(sentence, regex(combined_pattern, ignore_case = TRUE))) %>%
  select(index, sentence_id)

# 2. Create the window boundaries
windows <- matches %>%
  mutate(start = sentence_id - 2,
         end = sentence_id + 3) %>%
  group_by(index) %>%
  # This magic step merges overlapping windows in the same article
  arrange(start) %>%
  mutate(group = cumsum(start > lag(end, default = first(end)))) 
# 3. Extract the merged blocks
final_reports <- windows %>%
  group_by(index, group) %>%
  summarize(start_window = min(start),
            end_window = max(end), .groups = "drop") %>%
  inner_join(sentences_df, by = "index") %>%
  filter(sentence_id >= start_window & sentence_id <= end_window) %>%
  group_by(index, group) %>%
  summarize(report_text = paste(sentence, collapse = " "), .groups = "drop")


write_csv(final_reports, "../output/validate_lynching_jan_10.csv")
```

### Validate

examine this spreadsheet
https://docs.google.com/spreadsheets/d/11m5UJKkkJ5TynC-TKTFZk3km4lZLHe_o-QUlkN2PmXs/edit?gid=0#gid=0
### Examine specific articles
```{r}


# Filter the dataframe for that index and grab the text
report_content <- precivil_complete %>%
  filter(index == "preciv120") %>%
  pull(spellchecked)

wrapped_text <- strwrap(as.character(report_content, width = 60))
writeLines(wrapped_text, "temp_text.txt")
file.edit("temp_text.txt")
```

### Print out articles for verification
```{r}
# Now create the text file with locations highlighted
file_conn <- file("verify_excerpts.txt", "w")

for (i in 1:nrow(final_reports)) {
  row_data <- final_reports[i, ]
  
  writeLines(paste0("Document ID: ", row_data$index), file_conn)
  # Wrap the full text
  wrapped_text <- strwrap(row_data$report_text, width = 80)
  wrapped_text[1] <- paste0("Full Context: ", wrapped_text[1])
  writeLines(wrapped_text, file_conn)
  
  if (i < nrow(final_reports)) {
    writeLines("\n==================================================\n", file_conn)
  }
}

close(file_conn)
```


### single df examination
```{r}
wrapped_text <- strwrap(as.character(context_extracted$context_block[2]), width = 60)
writeLines(wrapped_text, "temp_text.txt")
file.edit("temp_text.txt")
```





attempts
```{r}
death | punished | attempted

attempted to lynch 
they would probably have been lynched
tar - feather
 lashes given them on the bare back
```



### States
```{r}
precivil_complete  %>% 
count(newspaper_state) %>% 
 mutate(pct = formattable::percent(n/sum(n,1))) |> 
  arrange(desc(pct))

```


#KWIC

```{r}

# Create a corpus using the 'article_spellchecked' column as the text field
my_corpus <- corpus(precivil_complete, text_field = "spellchecked", docid_field = "index") # build a new 
# Tokenize the corpus
my_tokens <- tokens(my_corpus)

#15 tokens on both sides of the search term
quanteda_df_llm <- kwic(my_tokens, phrase(c("lynch")), window = 15, valuetype = "regex") %>% as.data.frame() |> 
  distinct(docname,  .keep_all = TRUE)

quanteda_df_llm <- kwic(my_tokens, phrase(c("lynch|hung|hang(ed|ing)?|executed?|execution|mob\\s+law|vigilance\\s+committee")), window = 20, valuetype = "regex") %>% as.data.frame() |> 
  distinct(docname,  .keep_all = TRUE)

df_llm <- precivil_complete |> 
  inner_join(quanteda_df_llm, by=c("index"="docname")) |> 
  mutate(kwic_text = paste0(pre," ", keyword," ",post)) |> 
   select(-c(from, to,pre,post, keyword, pattern)) |> 
  mutate(word_count_select = str_count(select_text, "\\w+"))
#447 processed articles!
#write_csv(df_llm, "../data/precivil_df_llm_dec_26.csv")

```

### Bigrams
Tokenize

```{r}

stories <- str_replace_all(df_llm$kwic_text, "- ", "")
stories_df <- tibble(stories,)

# unnest includes lower, punct removal

stories_tokenized <- stories_df %>%
  unnest_tokens(word,stories)

stories_tokenized
```
 Remove stopwords


```{r}

data(stop_words)

stories_tokenized <- stories_tokenized %>%
  anti_join(stop_words, by = c("word" = "word")) %>%
  filter(word != "temp_file") %>%
  filter(word != "stories_corpus") %>%
  filter(!grepl('[0-9]', word))

# Word Count

story_word_ct <- stories_tokenized %>%
  count(word, sort=TRUE)

```

### Create Bigrams

```{r}
bigrams <- stories_df %>%
  unnest_tokens(bigram, stories, token="ngrams", n=2)


#Filter out stop words.


bigrams1 <- bigrams %>%
  separate(bigram, c("word1", "word2"), sep = " ")

bigrams2 <- bigrams1 %>%
  filter(!word1 %in% stop_words$word) %>%
  filter(!word2 %in% stop_words$word)

bigrams <- bigrams2 %>%
  count(word1, word2, sort = TRUE)


```

# Speech Parsing 
It uses the [udpipe](https://cran.r-project.org/web/packages/udpipe/index.html) package in R:

Tokenization, Parts of Speech Tagging, Lemmatization and Dependency Parsing with the 'UDPipe' 'NLP' Toolkit.
This natural language processing toolkit provides language-agnostic 'tokenization', 'parts of speech tagging', 'lemmatization' and 'dependency parsing' of raw text.

This code loads an English language model to permit tagging of parts of speech.
This version only deals with adjectives.
It then counts all adjectives across all articles and then counts adjectives by article. Credit to Claude.ai for providing the basis of this code.


### Load English language model

```{r}
udmodel <- udpipe_download_model(language = "english")
udmodel <- udpipe_load_model(udmodel$file_model)
```

### Annotate parts of speech on articles

```{r}
# Then run your annotation
annotated_text <- udpipe_annotate(udmodel, 
                                 x = precivil_complete$spellchecked, 
                                 doc_id = precivil_complete$index,
                                   parser = "default",
                                   tagger = "default")
annotated_df <- as.data.frame(annotated_text)
```




The annotated_df columns follow a "CoNLL-U" format for annotated linguistic data in Universal Dependencies.
It permits analysis of grammatical and syntactic structure of text.

Breaking down the details in an annotated dataframe from UDPipe:

-   doc_id: Identifier for the document - this matches harris_articles_df\$filename.

-   paragraph_id: Identifier for paragraphs within each document.

-   sentence_id: Identifier for sentences within each document.

-   sentence: The full text of the sentence.

-   token_id: Identifier for each token (word, punctuation) within a sentence.

-   token: The actual word or punctuation mark.

-   lemma: The base or dictionary form of the token (e.g., "running" → "run").
    \
    upos: Universal Part of Speech tag - a standardized grammatical category (e.g., NOUN, VERB, ADJ).

-   xpos: Language-specific part of speech tag - more detailed than upos, varies by language.
    feats: Morphological features - additional grammatical information like Number=Sing, Tense=Past.

-   head_token_id: The ID of the token that the current token depends on in the dependency parse.

-   dep_rel: Dependency relation - describes the syntactic relationship between a token and its head (e.g., "nsubj" for nominal subject).

-   deps: Enhanced dependency graph in the format of head-relation pairs.

-   misc: Miscellaneous information that doesn't fit elsewhere, often contains spans of tokens or other metadata.

Learn more about [Universal Dependencies format](https://universaldependencies.org/format.html)



A nerdy paper about this process [Tokenizing, POS Tagging, Lemmatizing and Parsing UD 2.0 with UDPipe](https://ufal.mff.cuni.cz/~straka/papers/2017-conll_udpipe.pdf)


## Count adjectives
```{r}
# Filter for adjectives (upos = "ADJ") and count frequencies
adjective_counts <- annotated_df %>%
  filter(upos == "ADJ") %>%
  group_by(token) %>%
  summarise(
    frequency = n(),
    articles = n_distinct(doc_id)
  ) %>%
  arrange(desc(frequency))

# View the top 20 most frequent adjectives
head(adjective_counts, 20)

datatable(adjective_counts,
          caption = "Top adjectives in Precivil War Lynching coverage",
          options = list(pageLength = 50)) 

```

Interesting results:
--large is #8, appears 165 times in 114 articles. Large mob? Large gathering?
--good, #9, 154 times in 106 articles. Good result from the lynching?
--dead, #14, 117 times iun 84 articles. the lynching involved death at some level. did "dead" -- like murder -- preciptate the lynching?
--white, #24, 78 times in 47 articles. white men? racialized marker
--Indian, #28, 62 times in 45 articles. suggesting indians were lynched or involved?
--respectable #30, 58 times in 47 articles. respectable citizens who carried out lynching?
--Mexican, #46, 42 times in 29 articles. Were they being lynched?
--guilty, #59, 38 times in 32 articles 
--notorious, #77, 31 times in 29 articles
--black, #87, 28 times iun 26 articles
--Southern, #93, 27 tuimes in 16 articles
--slave, #157, 17 times in 14 articles



## Count adjectives by article
```{r}
adjectives_by_article <- annotated_df %>%
  filter(upos == "ADJ") %>%
  group_by(doc_id, token) %>%
  summarise(
    frequency = n()
  ) %>%
  arrange(doc_id, desc(frequency))

datatable(adjectives_by_article,
          caption = "Adjectives by article in antebellow lynching coverage",
          options = list(pageLength = 50)) 

```

# Extract adjective-noun pairs

This section extracts adjective-noun pairs where the adjective directly modifies the noun. It then tabulates the pairs by frequency and by article.

```{r}
adj_noun_pairs <- annotated_df %>%
  inner_join(
    annotated_df %>% select(doc_id, sentence_id, token, token_id, upos),
    by = c("doc_id", "sentence_id", "head_token_id" = "token_id")
  ) %>%
  # Filter for adjectives modifying nouns
  filter(
    dep_rel == "amod" &    # amod = adjectival modifier
    upos.x == "ADJ" &      # first token is adjective
    upos.y == "NOUN"       # head word is noun
  ) %>%
  # Select and rename relevant columns
  select(
    article_id = doc_id,
    sentence_id,
    adjective = token.x,
    noun = token.y
  )

# Count frequencies of adjective-noun pairs
adj_noun_counts <- adj_noun_pairs %>%
  group_by(adjective, noun) %>%
  summarise(
    frequency = n(),
    articles = n_distinct(article_id)
  ) %>%
  arrange(desc(frequency))

datatable(adj_noun_counts,
          caption = "Top adjectives-noun pairs in Antebellum Lynching coverage",
          options = list(pageLength = 50)) 

```

Interesting results

--white man, #9, 24 times in 17 articles
--white men, #15, 17 times in 16 articles
--Lynch law #10, 21 times in 16 articles
--armed men, #11, 21 times in 18 articles
--respectable citizens, #13, 20 times in 18 articles
--good citizens, #67, 7 times in 7 articles
--fair trial, #16, 16 times in 16 articles
--cold blood, #26, 12 times in 10 articles
--grand jury, #30, 11 times in 11 articles
--Northern man, #36, 10 times in 3 articles


### Adjective - noun pairs by article
```{r}
pairs_by_article <- adj_noun_pairs %>%
  group_by(article_id, adjective, noun) %>%
  summarise(
    frequency = n()
  ) %>%
  arrange(article_id, desc(frequency))

datatable(pairs_by_article,
          caption = "Adjectives-noun pairs by article in lynching coverage",
          options = list(pageLength = 50)) 
```


## Spacy parsing
```{r}

# Initialize spacy (one-time setup required: spacy_install())
spacy_initialize(model = "en_core_web_sm")
```

```{r}
text <- data.frame(
  doc_id = df_llm$index,
  text = df_llm$kwic_text,
  stringsAsFactors = FALSE
)

# Parse with spacyr
parsed <- spacy_parse(text, entity = TRUE, 
                               lemma = TRUE,
                               pos = TRUE,
                               tag= TRUE,  
                              dependency = TRUE  )


```


### Spacy Adjective-noun pairs
using a much narrower corpus of the kwic text that is fitered to 30 tokens
```{r}

# Create adjective-noun pairs using spaCy output
spacy_adj_noun_pairs <- parsed %>%
  inner_join(
    parsed %>% select(doc_id, sentence_id, token, token_id, pos),
    by = c("doc_id", "sentence_id", "head_token_id" = "token_id"),
    suffix = c("_adj", "_noun")
  ) %>%
  # Filter for adjectives modifying nouns
  filter(
    dep_rel == "amod" &           # amod = adjectival modifier
    pos_adj == "ADJ" &            # first token is adjective
    pos_noun == "NOUN"            # head word is noun
  ) %>%
  # Select and rename relevant columns
  select(
    article_id = doc_id,
    sentence_id,
    adjective = token_adj,
    noun = token_noun
  )

# Count frequencies of adjective-noun pairs
spacy_adj_noun_counts <- spacy_adj_noun_pairs %>%
  group_by(adjective, noun) %>%
  summarise(
    frequency = n(),
    articles = n_distinct(article_id),
    .groups = "drop"
  ) %>%
  arrange(desc(frequency))

# Display results
datatable(spacy_adj_noun_counts,
          caption = "Spacy Top adjective-noun pairs in Antebellum Lynching coverage",
          options = list(pageLength = 50))
```


# Deaths
```{r}
deaths <- precivil_complete |> 
  filter(entry_type == "Primary") |> 
  filter(category=="death") |>
  filter(!is.na(state)) |> 
  group_by(state) |> 
  count(category) |> 
  ungroup() |> 
  mutate(Percent = as.character(formattable::percent(n/sum(n),0))) |> 
  arrange(desc(n)) |> 
  rename(Death_Total = n, State = state) |> 
  select(State, Death_Total, Percent)

datatable(deaths,
          # Use HTML styling directly in the caption string
          caption = htmltools::tags$caption(
            style = 'caption-side: top; text-align: center; color: black; font-size: 20pt; font-weight: bold;',
            'Lynching Deaths in Newspaper Reports, 1835-1865'
          ),
          options = list(
            pageLength = 50,
            searching = FALSE, 
            lengthChange = FALSE
          ))
```


### All Cases

```{r}

dt_total <- new_total |> 
  filter(!is.na(state)) |> 
  rename(State = state, Total = n)

datatable(dt_total,
          # Use HTML styling directly in the caption string
          caption = htmltools::tags$caption(
            style = 'caption-side: top; text-align: center; color: black; font-size: 20pt; font-weight: bold;',
            'Top locations for pre-Civil War lynchings'
          ),
          options = list(
            pageLength = 50,
            searching = FALSE, 
            lengthChange = FALSE
          ))
```


### Map output
```{r}
map <- deaths
write_csv(map, "../output/map.csv")

```

# Lynching Analysis Over Time

```{r}

precivil_complete |> 
    filter(entry_type == "Primary") |> 
  group_by(year, category) |> 
  count()

```

# Time series graphic

```{r}
precivil_complete |> 
    filter(entry_type == "Primary") |> 
    filter(!is.na(category)) |> 
  filter(!category %in% c("other", "unclear")) |> 
  group_by(year, category) |> 
  count() |> 
  mutate(category = factor(category, levels = c("attempt", "punish", "death"))) |>
 ggplot(aes(x = year, y = n, fill = category)) +
  geom_col() +
  scale_x_continuous(breaks = seq(1835, 1865, by = 5)) +
  labs(
    title = "Types of Lynchings, 1835-1865",
    x = "Year",
    y = "Count of Entries",
    caption = "Source: newspaper articles, Library of Congress, Graphic by Rob Wells, 1-13-2026"
  ) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

```


```{r}
state_list <- precivil_complete |> 
  filter(!is.na(state)) |> 
  distinct(state) |> 
  arrange(state) |> 
  pull(state)

precivil_complete |> 
  filter(entry_type == "Primary") |> 
  filter(!is.na(state), !is.na(category)) |>
  filter(!category %in% c("other", "unclear")) |> 
  group_by(state, category) |> 
  count() |> 
  # Reordering so "death" is at the bottom of the Y-axis
  mutate(category = factor(category, levels = c("death", "punish", "attempt"))) |> 
  ggplot(aes(x = category, y = state, fill = n)) +
  geom_tile(color = "white") +
  scale_y_discrete(limits = rev(state_list)) +# White border makes the "tiles" pop
 scale_fill_gradient(low = "#4169E1", high = "#FF0000") + 
  labs(
    title = "Intensity of Mentions by Category (1835-1865)",
    x = NULL,      # Drops X-axis label
    y = NULL,      # Drops Y-axis label
    caption = "Data source: Library of Congress, Graphic by Rob Wells, 1-13-2026"
  ) +
  theme_minimal() +
  theme(
    legend.position = "none", # Drops the legend
    axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1)
  )
```

```{r}
counts <- precivil_complete |> 
  filter(entry_type == "Primary") |> 
  filter(!is.na(state), !is.na(category)) |>
  filter(!category %in% c("other", "unclear")) |> 
  group_by(state, category) |> 
  count() 
```

