---
title: "Antebellum Lynching Analysis 2025"
author: "Rob Wells"
date: "2025-12-22"
output: html_document
---

```{r}
#install.packages("udpipe")
#install.packages('DT')
#install.packages("arrow")
#install.packages("spacyr")
library(arrow)
library(jsonlite)
library(janitor)
library(lubridate)
library(quanteda)
library(tidytext)
library(tidyverse)
library(udpipe)
library(DT)
library(spacyr)
library(stringr)

```

#Import 508 cleaned articles
```{r}
precivil_complete <- read_csv("../data/precivil_complete_jan_3.csv") 



```


#Chart - Results by year
```{r}
#Here is a chart, Figure 1, that describes lynching search results by year
precivil_complete %>% 
   group_by(year) %>% 
  count(year) %>% 
#Sandwich it onto a simple ggplot
  ggplot(aes(x = year, y = n, fill = n)) +
  geom_col(position = "dodge") + 
  theme(legend.position = "none") +
  scale_x_continuous(labels = c(seq(1825, 1865, 5)), breaks = seq(1825, 1865, 5)) +
   # scale_y_continuous(breaks = seq(0, 15, 1)) +
  labs(title = "Precivil War Lynching Coverage By Year, 1835-1861", 
       subtitle = "Count of Lynching Stories Examined",
       caption = "n=509  articles. Graphic by Rob Wells, 1/3/2026",
       y="Count of Articles",
       x="Year")

ggsave("../output/precivil_coverage_year_1_3_2026.png", device = "png",width=9,height=6, dpi=800)
```


year total descending
```{r}
precivil_complete  %>% 
count(year) %>% 
  mutate(pct = formattable::percent(n/sum(n,1))) |> 
  arrange(desc(pct))


```

newspapers total descending
```{r}
newspapers <- precivil_complete  %>% 
count(newspaper_name_clean) %>% 
  arrange(desc(n))
newspapers

```


# Word count metrics
```{r}
#average story length is 449 words, but some articles ran up to 1,618 words

precivil_complete |> 
  summarize(avg = mean(word_count, na.rm=TRUE),
            median = median(word_count, na.rm=TRUE),
            min = min(word_count, na.rm=TRUE),
            max = max(word_count, na.rm=TRUE),
            sd = sd(word_count, na.rm=TRUE))


```

### States
```{r}
precivil_complete  %>% 
count(newspaper_state) %>% 
 mutate(pct = formattable::percent(n/sum(n,1))) |> 
  arrange(desc(pct))

```


#KWIC

```{r}

# Create a corpus using the 'article_spellchecked' column as the text field
my_corpus <- corpus(precivil_complete, text_field = "spellchecked", docid_field = "index") # build a new 
# Tokenize the corpus
my_tokens <- tokens(my_corpus)

#15 tokens on both sides of the search term
quanteda_df_llm <- kwic(my_tokens, phrase(c("lynch")), window = 15, valuetype = "regex") %>% as.data.frame() |> 
  distinct(docname,  .keep_all = TRUE)

quanteda_df_llm <- kwic(my_tokens, phrase(c("lynch|hung|hang(ed|ing)?|executed?|execution|mob\\s+law|vigilance\\s+committee")), window = 20, valuetype = "regex") %>% as.data.frame() |> 
  distinct(docname,  .keep_all = TRUE)

df_llm <- precivil_complete |> 
  inner_join(quanteda_df_llm, by=c("index"="docname")) |> 
  mutate(kwic_text = paste0(pre," ", keyword," ",post)) |> 
   select(-c(from, to,pre,post, keyword, pattern)) |> 
  mutate(word_count_select = str_count(select_text, "\\w+"))
#447 processed articles!
#write_csv(df_llm, "../data/precivil_df_llm_dec_26.csv")

```

### Bigrams
Tokenize

```{r}

stories <- str_replace_all(df_llm$kwic_text, "- ", "")
stories_df <- tibble(stories,)

# unnest includes lower, punct removal

stories_tokenized <- stories_df %>%
  unnest_tokens(word,stories)

stories_tokenized
```
 Remove stopwords


```{r}

data(stop_words)

stories_tokenized <- stories_tokenized %>%
  anti_join(stop_words, by = c("word" = "word")) %>%
  filter(word != "temp_file") %>%
  filter(word != "stories_corpus") %>%
  filter(!grepl('[0-9]', word))

# Word Count

story_word_ct <- stories_tokenized %>%
  count(word, sort=TRUE)

```

### Create Bigrams

```{r}
bigrams <- stories_df %>%
  unnest_tokens(bigram, stories, token="ngrams", n=2)


#Filter out stop words.


bigrams1 <- bigrams %>%
  separate(bigram, c("word1", "word2"), sep = " ")

bigrams2 <- bigrams1 %>%
  filter(!word1 %in% stop_words$word) %>%
  filter(!word2 %in% stop_words$word)

bigrams <- bigrams2 %>%
  count(word1, word2, sort = TRUE)


```

# Speech Parsing 
It uses the [udpipe](https://cran.r-project.org/web/packages/udpipe/index.html) package in R:

Tokenization, Parts of Speech Tagging, Lemmatization and Dependency Parsing with the 'UDPipe' 'NLP' Toolkit.
This natural language processing toolkit provides language-agnostic 'tokenization', 'parts of speech tagging', 'lemmatization' and 'dependency parsing' of raw text.

This code loads an English language model to permit tagging of parts of speech.
This version only deals with adjectives.
It then counts all adjectives across all articles and then counts adjectives by article. Credit to Claude.ai for providing the basis of this code.


### Load English language model

```{r}
udmodel <- udpipe_download_model(language = "english")
udmodel <- udpipe_load_model(udmodel$file_model)
```

### Annotate parts of speech on articles

```{r}
# Then run your annotation
annotated_text <- udpipe_annotate(udmodel, 
                                 x = precivil_complete$spellchecked, 
                                 doc_id = precivil_complete$index,
                                   parser = "default",
                                   tagger = "default")
annotated_df <- as.data.frame(annotated_text)
```




The annotated_df columns follow a "CoNLL-U" format for annotated linguistic data in Universal Dependencies.
It permits analysis of grammatical and syntactic structure of text.

Breaking down the details in an annotated dataframe from UDPipe:

-   doc_id: Identifier for the document - this matches harris_articles_df\$filename.

-   paragraph_id: Identifier for paragraphs within each document.

-   sentence_id: Identifier for sentences within each document.

-   sentence: The full text of the sentence.

-   token_id: Identifier for each token (word, punctuation) within a sentence.

-   token: The actual word or punctuation mark.

-   lemma: The base or dictionary form of the token (e.g., "running" → "run").
    \
    upos: Universal Part of Speech tag - a standardized grammatical category (e.g., NOUN, VERB, ADJ).

-   xpos: Language-specific part of speech tag - more detailed than upos, varies by language.
    feats: Morphological features - additional grammatical information like Number=Sing, Tense=Past.

-   head_token_id: The ID of the token that the current token depends on in the dependency parse.

-   dep_rel: Dependency relation - describes the syntactic relationship between a token and its head (e.g., "nsubj" for nominal subject).

-   deps: Enhanced dependency graph in the format of head-relation pairs.

-   misc: Miscellaneous information that doesn't fit elsewhere, often contains spans of tokens or other metadata.

Learn more about [Universal Dependencies format](https://universaldependencies.org/format.html)



A nerdy paper about this process [Tokenizing, POS Tagging, Lemmatizing and Parsing UD 2.0 with UDPipe](https://ufal.mff.cuni.cz/~straka/papers/2017-conll_udpipe.pdf)


## Count adjectives
```{r}
# Filter for adjectives (upos = "ADJ") and count frequencies
adjective_counts <- annotated_df %>%
  filter(upos == "ADJ") %>%
  group_by(token) %>%
  summarise(
    frequency = n(),
    articles = n_distinct(doc_id)
  ) %>%
  arrange(desc(frequency))

# View the top 20 most frequent adjectives
head(adjective_counts, 20)

datatable(adjective_counts,
          caption = "Top adjectives in Precivil War Lynching coverage",
          options = list(pageLength = 50)) 

```

Interesting results:
--large is #8, appears 165 times in 114 articles. Large mob? Large gathering?
--good, #9, 154 times in 106 articles. Good result from the lynching?
--dead, #14, 117 times iun 84 articles. the lynching involved death at some level. did "dead" -- like murder -- preciptate the lynching?
--white, #24, 78 times in 47 articles. white men? racialized marker
--Indian, #28, 62 times in 45 articles. suggesting indians were lynched or involved?
--respectable #30, 58 times in 47 articles. respectable citizens who carried out lynching?
--Mexican, #46, 42 times in 29 articles. Were they being lynched?
--guilty, #59, 38 times in 32 articles 
--notorious, #77, 31 times in 29 articles
--black, #87, 28 times iun 26 articles
--Southern, #93, 27 tuimes in 16 articles
--slave, #157, 17 times in 14 articles



## Count adjectives by article
```{r}
adjectives_by_article <- annotated_df %>%
  filter(upos == "ADJ") %>%
  group_by(doc_id, token) %>%
  summarise(
    frequency = n()
  ) %>%
  arrange(doc_id, desc(frequency))

datatable(adjectives_by_article,
          caption = "Adjectives by article in antebellow lynching coverage",
          options = list(pageLength = 50)) 

```

# Extract adjective-noun pairs

This section extracts adjective-noun pairs where the adjective directly modifies the noun. It then tabulates the pairs by frequency and by article.

```{r}
adj_noun_pairs <- annotated_df %>%
  inner_join(
    annotated_df %>% select(doc_id, sentence_id, token, token_id, upos),
    by = c("doc_id", "sentence_id", "head_token_id" = "token_id")
  ) %>%
  # Filter for adjectives modifying nouns
  filter(
    dep_rel == "amod" &    # amod = adjectival modifier
    upos.x == "ADJ" &      # first token is adjective
    upos.y == "NOUN"       # head word is noun
  ) %>%
  # Select and rename relevant columns
  select(
    article_id = doc_id,
    sentence_id,
    adjective = token.x,
    noun = token.y
  )

# Count frequencies of adjective-noun pairs
adj_noun_counts <- adj_noun_pairs %>%
  group_by(adjective, noun) %>%
  summarise(
    frequency = n(),
    articles = n_distinct(article_id)
  ) %>%
  arrange(desc(frequency))

datatable(adj_noun_counts,
          caption = "Top adjectives-noun pairs in Antebellum Lynching coverage",
          options = list(pageLength = 50)) 

```

Interesting results

--white man, #9, 24 times in 17 articles
--white men, #15, 17 times in 16 articles
--Lynch law #10, 21 times in 16 articles
--armed men, #11, 21 times in 18 articles
--respectable citizens, #13, 20 times in 18 articles
--good citizens, #67, 7 times in 7 articles
--fair trial, #16, 16 times in 16 articles
--cold blood, #26, 12 times in 10 articles
--grand jury, #30, 11 times in 11 articles
--Northern man, #36, 10 times in 3 articles


### Adjective - noun pairs by article
```{r}
pairs_by_article <- adj_noun_pairs %>%
  group_by(article_id, adjective, noun) %>%
  summarise(
    frequency = n()
  ) %>%
  arrange(article_id, desc(frequency))

datatable(pairs_by_article,
          caption = "Adjectives-noun pairs by article in lynching coverage",
          options = list(pageLength = 50)) 
```


## Spacy parsing
```{r}

# Initialize spacy (one-time setup required: spacy_install())
spacy_initialize(model = "en_core_web_sm")
```

```{r}
text <- data.frame(
  doc_id = df_llm$index,
  text = df_llm$kwic_text,
  stringsAsFactors = FALSE
)

# Parse with spacyr
parsed <- spacy_parse(text, entity = TRUE, 
                               lemma = TRUE,
                               pos = TRUE,
                               tag= TRUE,  
                              dependency = TRUE  )


```


### Spacy Adjective-noun pairs
using a much narrower corpus of the kwic text that is fitered to 30 tokens
```{r}

# Create adjective-noun pairs using spaCy output
spacy_adj_noun_pairs <- parsed %>%
  inner_join(
    parsed %>% select(doc_id, sentence_id, token, token_id, pos),
    by = c("doc_id", "sentence_id", "head_token_id" = "token_id"),
    suffix = c("_adj", "_noun")
  ) %>%
  # Filter for adjectives modifying nouns
  filter(
    dep_rel == "amod" &           # amod = adjectival modifier
    pos_adj == "ADJ" &            # first token is adjective
    pos_noun == "NOUN"            # head word is noun
  ) %>%
  # Select and rename relevant columns
  select(
    article_id = doc_id,
    sentence_id,
    adjective = token_adj,
    noun = token_noun
  )

# Count frequencies of adjective-noun pairs
spacy_adj_noun_counts <- spacy_adj_noun_pairs %>%
  group_by(adjective, noun) %>%
  summarise(
    frequency = n(),
    articles = n_distinct(article_id),
    .groups = "drop"
  ) %>%
  arrange(desc(frequency))

# Display results
datatable(spacy_adj_noun_counts,
          caption = "Spacy Top adjective-noun pairs in Antebellum Lynching coverage",
          options = list(pageLength = 50))
```



### Cities and towns
Captures lynching locations!
```{r}

# Get lynch sentences
lynch_info <- parsed %>%
   filter(grepl("lynch|hung|hang(ed|ing)?|executed?|execution|mob\\s+law|vigilance\\s+committee", 
               token, ignore.case = TRUE)) %>%
  distinct(doc_id, sentence_id, lynch_token_id = token_id)

# For each location entity, check what word comes before it
locations_with_prep <- parsed %>%
  filter(str_detect(entity, "^(GPE|LOC|FAC)_(B|I)$")) %>%
  # Add the previous token for each row
  group_by(doc_id, sentence_id) %>%
  mutate(prev_token = lag(token)) %>%
  ungroup() %>%
  # Join with lynch info
  inner_join(lynch_info, by = c("doc_id", "sentence_id")) %>%
  # Filter for location prepositions and positions after lynch term
  filter(prev_token %in% c("at", "in", "near", "on", "from", "to", "near") |
         token_id > lynch_token_id)
```


### count cities

captures 226 places

```{r}

# Extract location entities and combine multi-word locations
location_counts2 <- locations_with_prep %>%
  filter(str_detect(entity, "^(GPE|LOC|FAC)_(B|I)$")) %>%
  # Group by doc_id and sentence_id to keep entities together
  group_by(doc_id, sentence_id) %>%
  # Create entity groups - each _B starts a new entity
  mutate(entity_group = cumsum(str_detect(entity, "_B$"))) %>%
  # Combine tokens within each entity group
  group_by(doc_id, sentence_id, entity_group) %>%
  summarise(
    location = paste(token, collapse = " "),
    .groups = "drop"
  ) 


```

### Normalize list

```{r}
location_counts2 <- location_counts2 |> 
  mutate(location_clean = tolower(location)) |> 
  mutate(location_clean = case_when(
    # 1. REMOVE NOISE (Returns NA)
    # Added: san, driskells, los
    location_clean %in% c("a", "the memphis enquirer of the", "the yoakum gang", 
                          "rev", "lieut", "jas", "sheriff", "fort", "5th", 
                          "5th september september", "states", "west", "red",
                          "a senor senor badlo badlo", "san", "driskells", "los") ~ NA_character_,
    
    # 2. SPECIFIC COMBINATIONS & CORRECTIONS
    # Rogue River variants -> Rogue River Valley
    location_clean %in% c("rogue", "rogue river", "rogue river river", 
                          "rogue river river valley valley") ~ "Rogue River Valley",
    
    # Alameda -> Alameda County
    location_clean == "alameda" ~ "Alameda County",
    
    # Annador -> Amador County
    location_clean == "annador county" ~ "Amador County",
    
    # Gentry variants -> Gentry County
    location_clean %in% c("gentry", "gentry county county") ~ "Gentry County",
    
    # 3. MAPPINGS & ABBREVIATIONS
    location_clean %in% c("mo", "mo .", "mo . .") ~ "Missouri",
    location_clean %in% c("ky", "ky .")            ~ "Kentucky",
    location_clean %in% c("ill", "ill . .")        ~ "Illinois",
    location_clean %in% c("la .")                  ~ "Louisiana",
    location_clean %in% c("n.c")                   ~ "North Carolina",
    
    # 4. TYPO FIXES (Specific cases)
    location_clean == "san francisco francisco"    ~ "San Francisco",
    location_clean == "los angeles angeles"        ~ "Los Angeles",
    location_clean == "santa barbara barbara"      ~ "Santa Barbara",
    location_clean == "red bluffs bluffs"          ~ "Red Bluffs",
    location_clean == "san luis luis obispo obispo" ~ "San Luis Obispo",
    location_clean == "salmon river river"         ~ "Salmon River",
    location_clean == "sawyer 's 's bar bar"       ~ "Sawyer's Bar",
    
    # 5. DEFAULT: Standardize everything else to Title Case
    # This handles Alabama, Nashville, Iowa, etc., automatically.
    TRUE ~ str_to_title(location_clean)
  )) |> 
  filter(!is.na(location_clean)) |> 
  group_by(location_clean) |> 
  summarise(
    freq_total = n(), 
    articles_total = n_distinct(doc_id), 
    .groups = "drop"
  ) |> 
  arrange(desc(freq_total))
```


### Datatable of locations
```{r}

datatable(location_counts2,
          # Use HTML styling directly in the caption string
          caption = htmltools::tags$caption(
            style = 'caption-side: top; text-align: center; color: black; font-size: 20pt; font-weight: bold;',
            'Top locations for pre-Civil War lynchings'
          ),
          options = list(
            pageLength = 50,
            searching = FALSE, 
            lengthChange = FALSE
          ))
```

### States from localities

```{r}
location_counts2 <- location_counts2 |> 
  mutate(state = case_when(
    # Already state names - pass through
    location_clean %in% c("Alabama", "Arkansas", "California", "Delaware", "Georgia", 
                          "Illinois", "Indiana", "Iowa", "Kansas", "Kentucky", 
                          "Louisiana", "Minnesota", "Mississippi", "Missouri", 
                          "Nebraska", "New Hampshire", "New York", "North Carolina",
                          "Ohio", "Tennessee", "Texas", "Virginia") ~ location_clean,
    
    # California counties and cities
    location_clean %in% c("Alameda", "Alameda County", "Annador County", "Calaveras County",
                          "Merced County", "Napa", "Oakland", "Placer County", 
                          "San Andreas", "San Antonio", "San Felipe", "San Francisco",
                          "San Gabriel", "San Isidro", "San Jose", "San Luis Obispo",
                           "Santa Barbara", "Shasta", "Solano County", 
                          "Stockton", "Tejon", "Tuolumne County", "Yuba County",
                          "Red Bluffs", "Drytown", "Iowa Hill", "Monterey",
                          "Sawyer's Bar", "South Sulphur") ~ "California",
    
    # Texas counties and cities
    location_clean %in% c("Uvalde County", "El Paso", "San Antonio") ~ "Texas",
    
    # Missouri counties and cities
    location_clean %in% c("Buchanan County", "Gentry County") ~ "Missouri",
    
    # Illinois counties and cities
    location_clean %in% c("Ogle County", "Keithsburg", "Rockford") ~ "Illinois",
    
    # Indiana counties and cities
    location_clean %in% c("South Bend") ~ "Indiana",
    #deleted "Noble County" "Allen County",  - too many variations
    
    # Mississippi counties and cities
    location_clean %in% c("Issaquena County", "Panola", "Pontotoc County") ~ "Mississippi",
    
    # Iowa counties and cities
    location_clean %in% c("Black Hawk", "Tama County") ~ "Iowa",
    
    # Kentucky cities
    location_clean %in% c("Lexington", "Louisville") ~ "Kentucky",
    
    # Tennessee cities
    location_clean %in% c("Memphis", "Nashville") ~ "Tennessee",
    
    # Georgia cities
    location_clean %in% c("Savannah") ~ "Georgia",
    
    # Arkansas counties and cities
    location_clean %in% c("Pike County", "Tuskegee") ~ "Alabama",
    
    # Louisiana cities
    location_clean %in% c("Baton Rouge", "New Orleans") ~ "Louisiana",
    
    # Kansas cities
    location_clean %in% c("Lecompton", "Leavenworth City", "Kansas Territory") ~ "Kansas",
    
    # New York cities
    location_clean == "New York" ~ "New York",
    
    # Ohio cities
    location_clean == "Williamstown" ~ "Kentucky",
    
    # Florida cities
    location_clean == "Pensacola" ~ "Florida",
    
    # Nebraska cities
    location_clean == "Omaha City" ~ "Nebraska",
    
    # Colorado cities
    location_clean == "Denver City" ~ "Colorado",
    
    # Oregon counties
    location_clean %in% c("Lewis County", "Rogue River", "Rogue River Valley") ~ "Oregon",
    
    # North Carolina cities
    location_clean == "Wilmington" ~ "North Carolina",
    
    # District of Columbia
    location_clean == "Washington" ~ "District of Columbia",
    
    # Rivers/geographic features - assign to primary state
    location_clean == "Mississippi" ~ "Mississippi",
    location_clean == "Upper Mississippi River" ~ "Multiple States",
    location_clean == "Salmon River" ~ "Idaho",
    
    # Uncertain/generic locations
    location_clean %in% c("Columbia", "Geneva", "Milton", "Alston", "Bargar",
                          "Driskells", "Partridge", "Renick", "Scaggs", 
                          "Shattuck", "Weatherford", "Weston", "Woodstock",
                          "Shelbyville", "Lawrenceville", "Union City",
                          "Grant County", "Randolph County", "Petersburg", "Essex County", "Melbourne", "Mount Alexander") ~ "Unknown",
    
    
    # Default
    TRUE ~ "Unknown"
  ))

# View results
#unknown = 39
map <- 
location_counts2 |> 
  count(state) |> 
  arrange(desc(n)) |> 
  filter(state!="Unknown")

write_csv(map, "../output/map.csv")
  
```

The
whole neighborhood was aroused to a frenzy of excitement so
soon as the bloody deed became known. 


### Quick view of any entry by row
```{r}
wrapped_text <- strwrap(as.character(precivil_complete$spellchecked[162]), width = 60)
writeLines(wrapped_text, "temp_text.txt")
file.edit("temp_text.txt")
```




## validate-read city entries

```{r}
# First, merge locations with your original dataframe
# Assuming your original docs have a doc_id that matches locations3$doc_id

locations_with_context <- locations_with_prep %>%
  group_by(doc_id, sentence_id) %>%
  summarize(
    locations = paste(token, collapse = " "),
    entity_types = paste(unique(entity), collapse = ", "),
    .groups = "drop"
  ) %>%
  right_join(precivil_df, by = c("doc_id" = "index")) # adjust join key as needed

# Now create the text file with locations highlighted
file_conn <- file("lynching_locations.txt", "w")

for (i in 1:nrow(locations_with_context)) {
  row_data <- locations_with_context[i, ]
  
  writeLines(paste0("Document ID: ", row_data$doc_id), file_conn)
  writeLines(paste0("Sentence ID: ", row_data$sentence_id), file_conn)
  writeLines(paste0("LOCATION IDENTIFIED: ", row_data$locations), file_conn)
  writeLines(paste0("Entity Type: ", row_data$entity_types), file_conn)
  writeLines(paste0("Headline: ", row_data$headline), file_conn)
  writeLines(paste0("Newspaper: ", row_data$newspaper_name), file_conn)
  writeLines(paste0("Date: ", row_data$date), file_conn)

  
  # Wrap the full text
  wrapped_text <- strwrap(row_data$kwic_text, width = 80)
  wrapped_text[1] <- paste0("Full Context: ", wrapped_text[1])
  writeLines(wrapped_text, file_conn)
  
  if (i < nrow(locations_with_context)) {
    writeLines("\n==================================================\n", file_conn)
  }
}

close(file_conn)
```



### Print stories for validation
precivil narrow print specific stories in a text file
```{r}
#filter data to examine
precivil_df <- precivil_complete |> 
  filter(index=="preciv446")


# Open a connection to the text file
file_conn <- file("dataframe_entries.txt", "w")

# Loop through each row of the dataframe
for (i in 1:nrow(precivil_df)) {
  # Extract the row
  row_data <- precivil_df[i, c("headline", "newspaper_name", "date", "spellchecked", "article_id", "index")]
  
  # Write each field on a separate line
  writeLines(paste0("Headline: ", row_data$headline), file_conn)
  writeLines(paste0("Newspaper: ", row_data$newspaper_name), file_conn)
  writeLines(paste0("Date: ", row_data$date), file_conn)
  writeLines(paste0("Article ID: ", row_data$article_id), file_conn)
  writeLines(paste0("Index Number: ", row_data$index), file_conn)
   wrapped_explanation <- strwrap(row_data$spellchecked, width = 60)
  wrapped_explanation[1] <- paste0("Spellchecked Text: ", wrapped_explanation[1])
  writeLines(wrapped_explanation, file_conn)
  
  # # Handle the matches list column
  # matches_content <- unlist(row_data$matches)
  # if (length(matches_content) > 0) {
  #   writeLines(paste0("Matches: ", paste(matches_content, collapse = ", ")), file_conn)
  # } else {
  #   writeLines("Matches: None", file_conn)
  # }
  
  # writeLines(paste0("\nArticle:\n", row_data$spellchecked), file_conn)
  
  # Add separator line (except after the last entry)
  if (i < nrow(precivil_df)) {
    writeLines("\n**************************************************\n", file_conn)
  }
}

# Close the file connection
close(file_conn)
```


## print from df_llm
```{r}

#filter data to examine
precivil_df <- df_llm |> 
  filter(str_detect(kwic_text, "California"))


# Open a connection to the text file
file_conn <- file("dataframe_entries.txt", "w")

# Loop through each row of the dataframe
for (i in 1:nrow(precivil_df)) {
  # Extract the row
  row_data <- precivil_df[i, c("headline", "newspaper_name", "date", "kwic_text", "index")]
  
  # Write each field on a separate line
  writeLines(paste0("Headline: ", row_data$headline), file_conn)
  writeLines(paste0("Newspaper: ", row_data$newspaper_name), file_conn)
  writeLines(paste0("Date: ", row_data$date), file_conn)
  writeLines(paste0("Index Number: ", row_data$index), file_conn)
   wrapped_explanation <- strwrap(row_data$kwic_text, width = 60)
  wrapped_explanation[1] <- paste0("KWIC_text: ", wrapped_explanation[1])
  writeLines(wrapped_explanation, file_conn)
  
  if (i < nrow(precivil_df)) {
    writeLines("\n**************************************************\n", file_conn)
  }
}

# Close the file connection
close(file_conn)
```


#-------------------------------------------------------------------------------

#Notes

# missing 15 text files from students
```{r}
missing <- students |> 
  anti_join(files, by=c("file_id_txt"="filename")) 

write_csv(missing, "missing_student_text_dec_21.csv")
```


### Notes on parsing
https://claude.ai/share/9c186ac7-2e5f-4e94-a64d-14e1e79aa49e

using spacyr- needs work
```{r}


# Get sentence-level text
sentences_with_locations <- parsed %>%  # your spacyr parsed object
  filter(doc_id %in% locations3$doc_id,
         sentence_id %in% locations3$sentence_id) %>%
  group_by(doc_id, sentence_id) %>%
  summarize(
    sentence_text = paste(token, collapse = " "),
    .groups = "drop"
  ) %>%
  left_join(locations3 %>% 
              group_by(doc_id, sentence_id) %>%
              summarize(locations = paste(token, collapse = " "), .groups = "drop"))

# Join with your metadata
results <- sentences_with_locations %>%
  left_join(precivil_df, by = c("doc_id" = "index"))

# Write to file
file_conn <- file("lynching_locations_sentences.txt", "w")

for (i in 1:nrow(results)) {
  row <- results[i, ]
  
  writeLines(paste0("=== ENTRY ", i, " ==="), file_conn)
  writeLines(paste0("Location Found: ", row$locations), file_conn)
  writeLines(paste0("Document: ", row$doc_id), file_conn)
  writeLines(paste0("Newspaper: ", row$newspaper_name, " (", row$date, ")"), file_conn)
  writeLines("", file_conn)
  writeLines("Sentence with location:", file_conn)
  wrapped <- strwrap(row$sentence_text, width = 80)
  writeLines(paste0("  ", wrapped), file_conn)
  writeLines("", file_conn)
  writeLines("Full article context:", file_conn)
  wrapped_full <- strwrap(row$kwic_text, width = 80)
  writeLines(paste0("  ", wrapped_full), file_conn)
  writeLines("\n**************************************************\n", file_conn)
}

close(file_conn)
```



## people
```{r}
# For people
person_counts <- annotated_spacy %>%
  filter(str_detect(entity, "^PERSON_(B|I)$")) %>%
  group_by(doc_id, sentence_id) %>%
  mutate(entity_group = cumsum(str_detect(entity, "_B$"))) %>%
  group_by(doc_id, sentence_id, entity_group) %>%
  summarise(
    person = paste(token, collapse = " "),
    .groups = "drop"
  ) %>%
  group_by(person) %>%
  summarise(
    frequency = n(),
    articles = n_distinct(doc_id)
  ) %>%
  arrange(desc(frequency))

# For organizations
org_counts <- annotated_spacy %>%
  filter(str_detect(entity, "^ORG_(B|I)$")) %>%
  group_by(doc_id, sentence_id) %>%
  mutate(entity_group = cumsum(str_detect(entity, "_B$"))) %>%
  group_by(doc_id, sentence_id, entity_group) %>%
  summarise(
    organization = paste(token, collapse = " "),
    .groups = "drop"
  ) %>%
  group_by(organization) %>%
  summarise(
    frequency = n(),
    articles = n_distinct(doc_id)
  ) %>%
  arrange(desc(frequency))
```


```{r}

library(stringr)

# Extract location entities and combine multi-word locations
location_counts2 <- annotated_spacy %>%
  filter(str_detect(entity, "^(GPE|LOC|FAC)_(B|I)$")) %>%
  # Group by doc_id and sentence_id to keep entities together
  group_by(doc_id, sentence_id) %>%
  # Create entity groups - each _B starts a new entity
  mutate(entity_group = cumsum(str_detect(entity, "_B$"))) %>%
  # Combine tokens within each entity group
  group_by(doc_id, sentence_id, entity_group) %>%
  summarise(
    location = paste(token, collapse = " "),
    .groups = "drop"
  ) %>%
  # Now count across all documents
  group_by(location) %>%
  summarise(
    frequency = n(),
    articles = n_distinct(doc_id)
  ) %>%
  arrange(desc(frequency)) |> 
  mutate(pct =  formattable::percent(articles/508,1))

head(location_counts, 20)
```

```{r}
# Get doc_id + sentence_id combinations that contain lynch terms
lynch_sentences <- parsed %>%
  filter(grepl("lynch|hung", token, ignore.case = TRUE)) %>%
  distinct(doc_id, sentence_id)

# Filter locations to those sentences
locations3 <- parsed %>%
  filter(str_detect(entity, "^(GPE|LOC)_(B|I)$")) %>%
  semi_join(lynch_sentences, by = c("doc_id", "sentence_id"))
```


Notes
These are BIO tags (Beginning-Inside-Outside) used in Named Entity Recognition to mark entity boundaries:
Entity Types:

GPE = Geopolitical Entity - locations with governments (cities, states, countries)

Examples: "San Francisco", "California", "United States", "Mexico"


LOC = Location - geographical locations that aren't geopolitical entities

Examples: mountain ranges, bodies of water, regions without government
Examples: "Pacific Ocean", "Sierra Nevada", "the West"



BIO Tags:

B (Beginning) = First token of an entity

GPE_B: 703 entities that start with a geopolitical name
LOC_B: 73 entities that start with a geographical location


I (Inside) = Continuation token of a multi-word entity

GPE_I: 229 tokens that are part of multi-word place names (e.g., "San" is B, "Francisco" is I)
LOC_I: 52 tokens continuing location name



## count cities - tighter filter
```{r}
#text <- df_llm$kwic_text
text <- data.frame(
  doc_id = df_llm$index,
  text = df_llm$kwic_text,
  stringsAsFactors = FALSE
)

# Parse with spacyr
parsed <- spacy_parse(text, entity = TRUE, 
                               lemma = TRUE,
                               pos = TRUE,
                               tag= TRUE)



# Find entities near "lynch" related words
lynch_positions <- which(grepl("lynch|hung", parsed$token, ignore.case = TRUE))

# Look for GPE (geopolitical entities) or LOC near these positions
window_size <- 5  # tokens before/after

# Create a helper to check if position is within window
is_near_lynch <- function(token_id, doc_id, parsed, lynch_positions, window_size) {
  # Get lynch positions in the same document
  doc_lynch_pos <- lynch_positions[parsed$doc_id[lynch_positions] == doc_id]
  
  if(length(doc_lynch_pos) == 0) return(FALSE)
  
  # Get actual token_id values for lynch positions in this doc
  lynch_token_ids <- parsed$token_id[doc_lynch_pos]
  
  # Check if current token_id is within window of any lynch position
  any(abs(token_id - lynch_token_ids) <= window_size)
}

locations3 <- parsed %>%
  filter(str_detect(entity, "^(GPE|LOC)_(B|I)$")) %>%
  filter(sentence_id %in% parsed$sentence_id[lynch_positions])


# Extract location entities and combine multi-word locations
location_counts3 <- parsed %>%
  filter(str_detect(entity, "^(GPE|LOC)_(B|I)$")) %>%
  filter(sentence_id %in% parsed$sentence_id[lynch_positions]) |> 
  # Group by doc_id and sentence_id to keep entities together
  group_by(doc_id, sentence_id) %>%
  # Create entity groups - each _B starts a new entity
  mutate(entity_group = cumsum(str_detect(entity, "_B$"))) %>%
  # Combine tokens within each entity group
  group_by(doc_id, sentence_id, entity_group) %>%
  summarise(
    location = paste(token, collapse = " "),
    .groups = "drop"
  ) %>%
  # Now count across all documents
  group_by(location) %>%
  summarise(
    frequency = n(),
    articles = n_distinct(doc_id)
  ) %>%
  arrange(desc(frequency)) |> 
  mutate(pct =  formattable::percent(articles/508,1))


```
```{r}
# Extract locations (GPE = geopolitical entity, LOC = location)
location_counts2 <- annotated_spacy %>%
    filter(entity %in% c("GPE_B", "GPE_I", "LOC_B", "LOC_I")) %>%
  group_by(token) %>%
  summarise(
    frequency = n(),
    articles = n_distinct(doc_id)
  ) %>%
  arrange(desc(frequency))
```

```{r}
# If you have entity information in your annotated_df
location_counts <- annotated_df %>%
  filter(upos == "PROPN") %>%
  # You can also filter by entity type if available
  filter(entity %in% c("LOC", "GPE")) %>%  # LOC = location, GPE = geopolitical entity
  group_by(token) %>%
  summarise(
    frequency = n(),
    articles = n_distinct(doc_id)
  ) %>%
  arrange(desc(frequency))
```

```{r}


# Then extract locations using entity information
location_counts <- annotated_df %>%
  filter(
    upos == "PROPN" | 
    str_detect(entity, "^B-LOC|^I-LOC|^B-GPE|^I-GPE")  # Named entity tags for locations
  ) %>%
  group_by(token) %>%
  summarise(
    frequency = n(),
    articles = n_distinct(doc_id)
  ) %>%
  arrange(desc(frequency))
```



#-------------------------------------------------------------------------------


# Tolnay and Beck Lynching Inventory
```{r}

tolnay_beck <- read_csv("../data/Bailey_Beck_lynching_list_8_1_2022.csv") %>% 
  as.data.frame()

tolnay_beck <- janitor::clean_names(tolnay_beck)
```


```{r}
tolnay_beck |> 
  count(status_clean)
```


# Filter the dataframe: 9937 articles
filtered_df <- df1 %>%
