---
title: "New_Lynching_Data_2025"
author: "Rob Wells"
date: "2025-12-26"
output: html_document
---

```{r}
library(tidyverse)
#install.packages("arrow")
library(arrow)
library(jsonlite)
library(janitor)
library(lubridate)
#install.packages("readtext")
library(quanteda)
```


# Import LLM scanned and Sean/Rob regex filtered lynching
data from line 996 lynching_all_years_wells_modified
```{r}

#447 probable precivil lynching stories
df_llm <- read_rds("/Users/gizmofo/Code/american_stories_lynching_article_llm/data/output_data/pre_civil_lynch/classified_articles/lynch_keyword_articles_classified.rds") |> 
  filter(probable_specific_lynching_article=="yes - evaluated by llm") |> 
  filter(x01_probably_lynch_story=="1") |> 
  mutate(index = paste0("preciv",row_number())) |> 
    mutate(extracted_string = str_extract(article_id, "(?:[^_]*_){3}([^_]+)")) |> 
  mutate(sn = str_replace(extracted_string, ".*_", "")) |> 
  mutate(date1 = as.Date(date, "%Y-%m-%d")) |> 
  select(-c(extracted_string, extracted_json_1, extracted_json_category_id, 
          extracted_json_article_id, output_json, recheck_valid_json_in_raw_response, 
          repaired_json, folder, x04_unknown, x01_probably_lynch_story, 
          x03_names_and_titles, x02_lynchburg, pass_to_llm, 
          other_lynch_word_matches, model_provider, model_type, 
          raw_response, original_raw_response, valid_json_in_raw_response, date)) |> 
  rename(date = date1, article_spellchecked = extracted_json_spellchecked_text) |> 
  mutate(word_count = str_count(article_spellchecked, "\\w+"))

```



# compare to precivil_complete
```{r}
#31 matches
matches <- df_llm |> 
  inner_join(precivil_complete, by =c("sn", "date"))

#416 new entries after filtering against the precivil_complete
df_llm <- df_llm |> 
  anti_join(precivil_complete, by =c("sn", "date"))
```

### Print out entries
```{r}

df_print <- df_llm |> 
  arrange(date)


# Open a connection to the text file
file_conn <- file("precivil_new_entries.txt", "w")

# Loop through each row of the dataframe
for (i in 1:nrow(df_print)) {
  # Extract the row
  row_data <- df_print[i, c("headline", "newspaper_name", "date", "article_spellchecked", "llm_explanation", "article_id", "index")]
  
  # Write each field on a separate line
  writeLines(paste0("Headline: ", row_data$headline), file_conn)
  writeLines(paste0("Newspaper: ", row_data$newspaper_name), file_conn)
  writeLines(paste0("Date: ", row_data$date), file_conn)
  writeLines(paste0("Article ID: ", row_data$article_id), file_conn)
  wrapped_explanation <- strwrap(row_data$llm_explanation, width = 60)
  wrapped_explanation[1] <- paste0("LLM Explains: ", wrapped_explanation[1])
  writeLines(wrapped_explanation, file_conn)
  writeLines(paste0("Index Number: ", row_data$index), file_conn)
  # Handle the matches list column
  # matches_content <- unlist(row_data$llm_explanation)
  # if (length(matches_content) > 0) {
  #   writeLines(paste0("LLM-explains: ", paste(llm_explanation, collapse = ", ")), file_conn)
  # } else {
  #   writeLines("Matches: None", file_conn)
  # }
  # Wrap article text at 60 characters
  article_text <- row_data$article_spellchecked
  wrapped_text <- strwrap(article_text, width = 60)
  
  writeLines("\nArticle:", file_conn)
  writeLines(wrapped_text, file_conn)
  
  # writeLines(paste0("\nArticle:\n", row_data$article), file_conn)
  
  # Add separator line (except after the last entry)
  if (i < nrow(df_print)) {
    writeLines("\n**************************************************\n", file_conn)
  }
}

# Close the file connection
close(file_conn)

```

### validating entries
```{r}
# valid <- c(325, xxxx)
#15-1 valid
valid <- c(preciv446, preciv445,preciv444,preciv44, preciv442, preciv441, preciv440, preciv439, preciv438, preciv437, preciv436, preciv435, preciv434, preciv433, preciv432 )
notlynch <-c(preciv447, )

df_llm <- df_llm |> 
  mutate(validate = if_else(index %in% valid, "Y", "N"))

write_csv(df_llm, "../output/df_llm_dec_26.csv")

```

#KWIC
### saved precivil_df_llm_dec_26.csv


```{r}

# Create a corpus using the 'article_spellchecked' column as the text field
my_corpus <- corpus(df_llm, text_field = "article_spellchecked", docid_field = "index") # build a new 
# Tokenize the corpus
my_tokens <- tokens(my_corpus)

#100 characters: Approximately 15-20 words, depending on word length.
#200 characters:  6 (avg chars per word + space) ≈ 30-40 words
#300 characters: about 45-60 words
quanteda_df_llm <- kwic(my_tokens, phrase(c("lynch")), window = 300, valuetype = "regex") %>% as.data.frame() |> 
  distinct(docname,  .keep_all = TRUE)

df_llm <- df_llm |> 
  inner_join(quanteda_df_llm, by=c("index"="docname")) |> 
  mutate(select_text = paste0(pre," ", keyword," ",post)) |> 
   select(-c(from, to,pre,post, keyword, pattern)) |> 
  mutate(word_count_select = str_count(select_text, "\\w+"))
#447 processed articles!
write_csv(df_llm, "../data/precivil_df_llm_dec_26.csv")

```

### Load df_llm
```{r}

df_llm <- read_csv("../data/precivil_df_llm_dec_26.csv") 
```

### FIX STATE COLUMN
```{r}
library(dplyr)
library(readr)



# 1. Download the Master Newspaper List from LOC (approx 10MB)
# This file contains the metadata for EVERY newspaper in their system.

loc_master_list <- read_csv("/Users/gizmofo/Code/hcij_lynching_phase_two/Storage_Older_Versions/Older Code/article_extraction/automated_article_identification/visualization_tool/static/newspapers.csv") |> 
  clean_names()

# 2. Look at what we got (Columns: lccn, title, state, etc.)
# Note: 'lccn' in their file is the same as your 'sn'
head(loc_master_list)

# 3. Join it to your dataframe locally
# We match your 'sn' column to their 'lccn' column
df_llm2_fixed <- df_llm |>
  left_join(
    loc_master_list |> select(lccn, state), 
    by = c("sn" = "lccn")
  ) |>
  # Move the new state data into your newspaper_state column
  mutate(newspaper_state = state) |>
  select(-state) # Remove the extra column from the join

# 4. Check how many are still missing
missing_count <- sum(is.na(df_llm2_fixed$newspaper_state))

missing_state <- df_llm2_fixed |> 
  filter(is.na(newspaper_state))

df_llm2_fixed |> 
  filter(is.na(newspaper_state)) |> 
  count(newspaper_name, sn)

message(paste("Remaining missing states:", missing_count))
```


```{r}
df_llm <- df_llm2_fixed |> 
  mutate(state2 = case_when(
    newspaper_name == "Arkansas intelligencer." ~ "Arkansas",
    newspaper_name == "The Van Buren press." ~ "Arkansas",
    newspaper_name == "Washington telegraph." ~ "Arkansas",
    newspaper_name == "The Daily spy." ~ "Massachusetts",
    newspaper_name == "Worcester daily spy." ~ "Massachusetts",
    newspaper_name == "Washington telegraph." ~ "Arkansas",
    newspaper_name == "The weekly Anglo-African." ~ "New York",
   TRUE ~ as.character(newspaper_state)
  )) |> 
  select(-c(newspaper_state)) |> 
  rename(newspaper_state = state2)



```



```{r}
precivil_complete <- read_csv("../data/precivil_complete_dec_28.csv") |> 
  mutate(index = as.character(index))


```

# write precivil_complete_dec_29.csv

```{r}

df_llm2 <- bind_rows(df_llm, precivil_complete)


df_llm2 <- df_llm2 |> 
  mutate(spellchecked = coalesce(spellchecked_text, article_spellchecked)) |> 
  select(-c(probable_specific_lynching_article,era, extracted_json_category_description,source, original_text, spellchecked_text, article_spellchecked, newspaper_state2)) |> 
  select(index, newspaper_name, date, year, spellchecked, article, word_count_select, newspaper_state, everything()) |> 
   mutate(word_count = str_count(spellchecked, "\\w+")) |> 
  distinct(index, .keep_all = TRUE)

#the 1825 entry isnt valid
df_llm2 <- precivil_complete |> 
  filter(!index=="preciv447")

write_csv(df_llm2, "../data/precivil_complete_dec_29.csv")


```






#-----------------------------------------------------
```{r}
df_llm2 <- df_llm |> 
  inner_join(precivil_complete, by=c("sn", "date")) 
```




```{r}
#selected snippets are average 276 words, median 273 words, max 562 words. SD 163
df_llm |> 
  summarize(avg = mean(word_count_select),
            median = median(word_count_select),
            min = min(word_count_select),
            max = max(word_count_select),
            sd = sd(word_count_select))
```


### Print out snippets
```{r}

df_print <- df_llm |> 
  arrange(date)


# Open a connection to the text file
file_conn <- file("df_llm_entries.txt", "w")

# Loop through each row of the dataframe
for (i in 1:nrow(df_print)) {
  # Extract the row
  row_data <- df_print[i, c("headline", "newspaper_name", "date", "select_text", "llm_explanation", "article_id", "index")]
  
  # Write each field on a separate line
  writeLines(paste0("Headline: ", row_data$headline), file_conn)
  writeLines(paste0("Newspaper: ", row_data$newspaper_name), file_conn)
  writeLines(paste0("Date: ", row_data$date), file_conn)
  writeLines(paste0("Article ID: ", row_data$article_id), file_conn)
  wrapped_explanation <- strwrap(row_data$llm_explanation, width = 60)
  wrapped_explanation[1] <- paste0("LLM Explains: ", wrapped_explanation[1])
  writeLines(wrapped_explanation, file_conn)
  writeLines(paste0("Index Number: ", row_data$index), file_conn)
  # Handle the matches list column
  # matches_content <- unlist(row_data$llm_explanation)
  # if (length(matches_content) > 0) {
  #   writeLines(paste0("LLM-explains: ", paste(llm_explanation, collapse = ", ")), file_conn)
  # } else {
  #   writeLines("Matches: None", file_conn)
  # }
  # Wrap article text at 60 characters
  article_text <- row_data$select_text
  wrapped_text <- strwrap(article_text, width = 60)
  
  writeLines("\nArticle:", file_conn)
  writeLines(wrapped_text, file_conn)
  
  # writeLines(paste0("\nArticle:\n", row_data$article), file_conn)
  
  # Add separator line (except after the last entry)
  if (i < nrow(df_print)) {
    writeLines("\n**************************************************\n", file_conn)
  }
}

# Close the file connection
close(file_conn)

```






write.csv(df_short_precivil, "../output/df_short_precivil_dec_26.csv")




#1720 - american sympathizers lynched
Ii may not be known to many of readers that the Colonel
Prince whose summary execution Of four OF the American
"sympathizers' with the Canadian macontens has been broug
under the notice of the House of Peers by Lord Brougham,
was some few years since resident of Cheltenham where he
prac- tsed as an attorney under the firm of Strator rd and
Prince. -JPorcester Herald.

#20 - man lynched for being in favor of union
A RuNoR.--It is reported that Mr.. John Shackeltord.
highly.respectable citizen ol | Dumfries, Virginia, was
taken from his house | on Friday by mob of secessionists,
and hung with grape vine in sight of his dwelling. The
cause assigned is, that Mr.. Shackelford was in favor of
the Union.


#-----------------------
#Creating precivil_complete
#-----------------------

```{r}
precivil_complete <- read_csv("../data/precivil_lynching_complete_12_22.csv")
```

# precivil narrow print specific stories in a text file, created from df1`
```{r}
#precivil tightly filtered
precivil_narrow <- filtered_df |> 
  filter(date < "1861-04-11")

#write.csv(precivil_narrow, "../data/precivil_narrow.csv")
#47 articles, 1844-1861

# Open a connection to the text file
file_conn <- file("dataframe_entries.txt", "w")

# Loop through each row of the dataframe
for (i in 1:nrow(precivil_narrow)) {
  # Extract the row
  row_data <- precivil_narrow[i, c("headline", "newspaper_name", "date", "article", "matches", "article_id", "index")]
  
  # Write each field on a separate line
  writeLines(paste0("Headline: ", row_data$headline), file_conn)
  writeLines(paste0("Newspaper: ", row_data$newspaper_name), file_conn)
  writeLines(paste0("Date: ", row_data$date), file_conn)
  writeLines(paste0("Article ID: ", row_data$article_id), file_conn)
  writeLines(paste0("Index Number: ", row_data$index), file_conn)
  
  # Handle the matches list column
  matches_content <- unlist(row_data$matches)
  if (length(matches_content) > 0) {
    writeLines(paste0("Matches: ", paste(matches_content, collapse = ", ")), file_conn)
  } else {
    writeLines("Matches: None", file_conn)
  }
  
  writeLines(paste0("\nArticle:\n", row_data$article), file_conn)
  
  # Add separator line (except after the last entry)
  if (i < nrow(precivil_narrow)) {
    writeLines("\n**************************************************\n", file_conn)
  }
}

# Close the file connection
close(file_conn)
```

# validating entries

```{r}
# valid <- c(325, 327, 1015, 1016, 1511, 1512, 1513,1514, 3067,3068,3069)

valid <- c(325, 327, 1015, 1016, 1511, 1512, 1513,1514, 3067,3068,3069,3070, 3071, 3072, 3073, 4685, 4686, 4687, 4688, 4689, 5973, 5974, 5975, 5976, 6880, 6881, 6882, 6883, 6885, 6887, 6889, 7519, 7520, 7521, 7522, 7523, 7524, 9645)
notlynch <-c(1014, 5972, 6884,6886, 6888, 6890,7525, 9646)

precivil_narrow <- precivil_narrow |> 
  mutate(validate = if_else(index %in% valid, "Y", "N"))

write.csv(precivil_narrow, "../output/precivil_narrow_dec_20.csv")

```


```{r}
precivil_narrow |> 
  count(validate) |> 
  mutate(pct = n/sum(n))

#81 percent of the articles involved lynching or attempted lynchings, 19 percent did not.
```


```{r}

precivil_narrow2 <- precivil_narrow |> 
  filter(validate == "Y") 
precivil_narrow |> 
  filter(validate == "Y") |> 
  count(year) |> 
  arrange(year)


```

## Read Google Sheet With Student Entries
```{r}
#imports 618 entries
precivil_students_entries <- rio::import("https://docs.google.com/spreadsheets/d/17nL8YBMIys7iJI1eeufiCpI1Plpmid2FSd6zuti5iSk/edit#gid=1208290507", skip=0) |>
   janitor::clean_names() |>
   rename(lynching = 'y_n_describes_a_lynching') |>
   mutate(id =  as.numeric(id)) |>
   mutate(sn = str_extract(url, "sn[^/]+"))

 precivil_students_entries <- precivil_students_entries |>
   slice(-c(619,620))




#46 lynching cases
students <- precivil_students_entries  |> 
  filter(lynching == "Yes") |> 
  mutate(source = "students") |> 
  mutate(date = as.Date(date)) |> 
  mutate(file_id_txt = paste0(file_id2, ".txt")) |> 
   mutate(file_id_txt = str_trim(file_id_txt))

```

# join students to main df by sn and date

```{r}
#matches 23 of 46
#only retrieves two records
joined_students <- df1 |> 
  mutate(
    sn = str_trim(sn),
    date1 = as.Date(date1)
  ) |> 
  inner_join(
    precivil_students_entries |> mutate(
      sn = str_trim(sn),
      date = as.Date(date)
    ), 
    by = c("sn", "date1" = "date")
  ) |> 
  distinct(article_id.x, .keep_all=TRUE)



```



# Compiler of raw text. 
```{r include=FALSE}
#3312 files
files <- list.files("~/Code/hcij_lynching_phase_two/Storage_Older_Versions/latest_articles_aug23/", pattern="*.txt") %>% 
  as.data.frame() %>%
  rename(filename = 1) 

index <- students |> 
  inner_join(files, by=c("file_id_txt"="filename")) %>%
  mutate(filepath = paste0("~/Code/hcij_lynching_phase_two/Storage_Older_Versions/latest_articles_aug23/",file_id_txt)) |> 
  rename(filename = file_id_txt)

#AI to the rescue
library(dplyr)
library(readr)
library(purrr)

# Function to process a chunk of rows
process_chunk <- function(chunk_rows, chunk_index) {
  articles_chunk <- tibble()
  
  for (row_value in chunk_rows) {
    temp <- index %>%
      slice(row_value)
    
    temp_filename <- temp$filename
    
    # Read lines from file
    tryCatch({
      articles_df_temp <- read_lines(temp$filepath) %>%
        as_tibble() %>%
        mutate(filename = temp_filename)
      
      # Append to chunk results
      articles_chunk <- bind_rows(articles_chunk, articles_df_temp)
    }, error = function(e) {
      message(paste("Error processing row", row_value, ":", conditionMessage(e)))
    })
  }
  
  # Write chunk to disk
  write_csv(articles_chunk, paste0("articles_chunk_", chunk_index, ".csv"))
}

# Determine chunk size and create chunks
chunk_size <- 500  # Adjust chunk size as needed
row_values <- split(1:nrow(index), ceiling(seq_along(1:nrow(index))/chunk_size))

# Process each chunk
walk2(row_values, seq_along(row_values), process_chunk)

# Combine all chunks into one data frame
chunk_files <- list.files(pattern = "articles_chunk_.*\\.csv")

articles_df <- map_dfr(chunk_files, read_csv)

# Perform the inner join after all files are processed
joined_df <- articles_df %>%
  select(filename, sentence = value) %>%
  inner_join(index, by = "filename")

write.csv(joined_df, ("students_22_precvil_text_dec_21_2025.csv"))


# View the resulting data frame
print(head(joined_df))

#captures text from 31 articles

```


```{r}
joined_df %>%
  summarize(n_unique_files = n_distinct(filename))
```
 #missing 15 text files from students
```{r}
missing <- students |> 
  anti_join(files, by=c("file_id_txt"="filename")) 

write_csv(missing, "missing_student_text_dec_23.csv")
```


```{r}
missing <- missing |> 
  mutate(seq = str_extract(url, "seq[^/]+"))
  
```


```{r}
# Create the missing pages lookup dataframe (without seq)
missing_pages_lookup <- data.frame(
  date = c("1805-01-14", "1805-01-19", "1805-02-01", "1835-07-31", "1835-08-01",
           "1835-09-05", "1835-09-05", "1835-09-08", "1835-09-12", "1835-10-20",
           "1835-10-20", "1835-10-22", "1836-11-09", "1851-05-10", "1851-06-03"),
  sn = c("sn83045242", "sn84024710", "sn85059525", "sn87062070", "sn84026826",
         "sn84026826", "sn84026826", "sn84024735", "sn84026826", "sn84024735",
         "sn84024735", "sn84024649", "sn83016620", "sn83030313", "sn82014594"),
  stringsAsFactors = FALSE
)

# Remove duplicate date-sn combinations
missing_pages_lookup <- missing_pages_lookup %>%
  distinct(date, sn)


# Filter your df for matching records
# Assuming your date column might need to be converted to character for matching
found_df <- df_precivil_new %>%
  mutate(date_char = as.character(date)) %>%
  semi_join(missing_pages_lookup, by = c("date_char" = "date", "sn" = "sn")) %>%
  select(-date_char)  # Remove the temporary column

# View results
print(found_df)

# Or if you want to see how many matches were found
cat("Found", nrow(found_df), "matching articles\n")
```


collapsed into one df
i want to take joined_df, which is 31 articles, one line per sentence, and produce a dataframe with 31 rows and a new column called "article" that has all of the text from "sentence" 

```{r}
articles_collapsed <- joined_df |>
  group_by(file_id2) |>
  summarize(
    article = paste(sentence, collapse = " "),
    across(everything(), first)
  ) |>
  ungroup() 

articles_collapsed <- articles_collapsed |> 
  select("file_id2", "date", "newspaper_name", "article", "newspaper_state", "filename", "attempted_y_n", "similar_coverage_y_n", "notes_from_te", "sn", "year", "month","day", "page","edition", "mod_id", "source", "filepath", "url" )

write_csv(articles_collapsed, "student_text_antebellum_12_21.csv")
```



# articles_collapsed print specific stories in a text file
```{r}
# Open a connection to the text file
file_conn <- file("articles_collapsed.txt", "w")

# Loop through each row of the dataframe
for (i in 1:nrow(articles_collapsed)) {
  # Extract the row
  row_data <- articles_collapsed[i, c("newspaper_name", "date", "article", "filename", "url")]
  
  # Write each field on a separate line
  # writeLines(paste0("Headline: ", row_data$headline), file_conn)
  writeLines(paste0("Newspaper: ", row_data$newspaper_name), file_conn)
  writeLines(paste0("Date: ", row_data$date), file_conn)
  writeLines(paste0("Filename: ", row_data$filename), file_conn)
  writeLines(paste0("URL: ", row_data$url), file_conn)
  
# Wrap article text at 60 characters
  article_text <- row_data$article
  wrapped_text <- strwrap(article_text, width = 60)
  
  writeLines("\nArticle:", file_conn)
  writeLines(wrapped_text, file_conn)
  
  # Add separator line (except after the last entry)
  if (i < nrow(articles_collapsed)) {
    writeLines("\n**************************************************\n", file_conn)
  }
}

# Close the file connection
close(file_conn)
```


#join precivil_narrow and articles collapsed

```{r}

precivil_narrow <- precivil_narrow |> 
  mutate(date = as.Date(date))
precivil_complete <-bind_rows(precivil_narrow, articles_collapsed)

#eliminates duplicates and irrelevant coverage
precivil_complete <- precivil_complete |> 
  filter(!file_id2 %in% c("16744_0", "16742_0", "16741_0", "16745_0"))

precivil_complete <- precivil_complete |> 
  filter(!validate %in% "N") |> 
  mutate(article_id2 = coalesce(article_id, file_id2)) |> 
  mutate(newspaper_state2 = coalesce(state, newspaper_state)) |> 
   subset(select = -c(article_id, file_id2, byline, date1, state, newspaper_state, newspapers, validate, url )) |> 
  select(newspaper_name, date, headline, article, cleaned_article, newspaper_state2, matches, notes_from_te, similar_coverage_y_n, everything())


write_csv(precivil_complete, "../data/precivil_lynching_complete_12_22.csv")

#74 articles from the narrow search of the feather files, all validates and the student verified files

```


```{r}
precivil_complete |> 
  anti_join(students, by=c("article_id2"="file_id2"))


```



Spot check results from Sean's rds file
rds$article[1]
--basically a threatened lynching 

rds$article[2]
--definitely a lynching

rds$article[74]
The LLM explanation was wrong: The article mentions 'lynch law' being used to justify such actions, indicating it relates to a lynching event." 
rds$llm_explanation[74]

rds$article[544]
--correct, not a lynching article

rds$article[545]
--correct, not a lynching article

rds$article[554]
--? this was a potential lynching article. category 2
the llm was correct: rds$llm_explanation[554]
[1] "This article describes a situation where there was an immediate threat of lynching by the crowd, though the lynching did not occur. The use of the word 'lynching' suggests that there was a potential for it to happen in the future."


```{r}

rds1 <- rio::import("/Users/gizmofo/Library/CloudStorage/GoogleDrive-robwells@umd.edu/My Drive/Current Projects Gdrive/Lynching UMD/Data/updated_lynch_keyword_articles_classified.rds")

```


Notes from the updated text grab: 6872 articles

filtered for mob. hung

rds1$article[8]
print(rds1[8, ])
--correct: definite attempted lynching
--this is not correct: extracted_json_category_description
8                          1 An article that describes a specific lynching event that has already happened.


print(rds1[10, ])
--Correct:  extracted_json_category_description
10                          1 An article that describes a specific lynching event that has already happened.


print(rds1[67, ])
--Correct, but it was one item in a long dispatch about many other misc events.
extracted_json_category_description
67                          1 An article that describes a specific lynching event that has already happened.


print(rds1[70, ])
--Correct.  extracted_json_category_description
70                          1 An article that describes a specific lynching event that has already happened.

(sorting to 1810)

print(rds1[4662, ])
--Correct, not a lynching article


print(rds1[6872, ])
--Correct, not a lynching article


(earliest positive hit: 1825)
print(rds1[788, ])
--INCORRECT. Mentions a murder but does not describe a lynching or mob violence. "The article describes the discovery of a murdered man's body, the suspicion of foul play, and the involvement of a deaf and dumb witness. The mention of apprehending a suspect and the description of violence make it fit into the category of a specific lynching event that has already happened."

print(rds1[775, ]) (1835)
--Correct, describes attempted lynching
--The LLM description is wrong: "An article that describes a specific lynching event that has already happened.
    extracted_json_1" - The person wasn't lynched
    
print(rds1[776, ]) (1835)
--Incorrect. a reference tothe vicksburg lynching in the round up of many events. "An article that describes a specific lynching event that has already happened." - so that is accurate but it lacks context

print(rds1[777, ]) (1835)
--Correct. "An article that describes a specific lynching event that has already happened.
    extracted_json_1"
    
  print(rds1[786, ]) (1835)
  --correct. extracted_json_category_description
786                          1 An article that describes a specific lynching event that has already happened.

print(rds1[787, ]) (1835)
--correct. extracted_json_category_description
787                          1 An article that describes a specific lynching event that has already happened.



