---
title: "Classify Lynching Location"
author: "Wells"
date: "2026-01-07"
output: html_document
---

This notebook takes the precivil_complete and tags geometery using spacyr. 
It then processes results of a manual review of all 508 entries and retags 292 entries with the correct states.


#Import 491 cleaned articles
```{r}
precivil_complete <- read_csv("../data/precivil_complete_jan_7.csv") 
```

## Spacy parsing
```{r}

# Initialize spacy (one-time setup required: spacy_install())
spacy_initialize(model = "en_core_web_sm") 
```

```{r}
text <- data.frame(
  doc_id = precivil_complete$index,
  text = precivil_complete$spellchecked,
  stringsAsFactors = FALSE
)

# Parse with spacyr
parsed <- spacy_parse(text, entity = TRUE, 
                               lemma = TRUE,
                               pos = TRUE,
                               tag= TRUE,  
                              dependency = TRUE  )


```

### Tag geographic entities
```{r}
# Get lynch sentences
lynch_info <- parsed |>
   filter(grepl("lynch|hung|hang(ed|ing)?|executed?|execution|mob\\s+law|vigilance\\s+committee", token, ignore.case = TRUE)) |>
  distinct(doc_id, sentence_id, lynch_token_id = token_id)

#tag entity near the lynching term appearance
locations_with_prep <- parsed |>
  filter(str_detect(entity, "^(GPE|LOC|FAC)_(B|I)$")) |>
  group_by(doc_id, sentence_id) |>
  mutate(prev_token = lag(token)) |>
  ungroup() |>
  inner_join(lynch_info, by = c("doc_id", "sentence_id"), relationship = "many-to-many") |>
  filter(prev_token %in% c("at", "in", "near", "on", "from", "to") |
         (token_id >= lynch_token_id - 5 & token_id < lynch_token_id) |
         token_id > lynch_token_id) |>
  distinct(doc_id, sentence_id, token_id, .keep_all = TRUE)
```


```{r}
location_counts3 <- locations_with_prep |>
  arrange(doc_id, sentence_id, token_id) |>
  group_by(doc_id, sentence_id) |>
  mutate(
    # Check if this token is consecutive with the previous one
    is_consecutive = (token_id - lag(token_id, default = -999)) == 1,
    # Create merged groups: start new group if NOT consecutive
    entity_group_merged = cumsum(!is_consecutive)
  ) |>
  group_by(doc_id, sentence_id, entity_group_merged) |>
  summarise(
    location = paste(token, collapse = " "),
    .groups = "drop"
  ) |> 
  arrange(desc(doc_id))
```


#new classifier

```{r}
new_location <- location_counts3 |> 
  # right_join(precivil_df, by = c("doc_id" = "index")) |> 
mutate(location_clean = tolower(location)) |> 
  # Remove duplicate consecutive words first
  mutate(location_clean = str_replace_all(location_clean, "\\b(\\w+)\\s+\\1\\b", "\\1")) |>
  mutate(location_clean = case_when(
    # 1. REMOVE NOISE (Returns NA)
    location_clean %in% c("a", "the memphis enquirer of the", "the yoakum gang", 
                          "rev", "lieut", "jas", "sheriff", "fort", "5th", 
                          "5th september september", "states", "west", "red",
                          "a senor senor badlo badlo", "san", "driskells", "los") ~ NA_character_,
    
    # 2. SPECIFIC COMBINATIONS & CORRECTIONS
    # Rogue River variants -> Rogue River Valley
    location_clean %in% c("rogue", "rogue river", "rogue river valley") ~ "Rogue River Valley",
    
    # Alameda -> Alameda County
    location_clean == "alameda" ~ "Alameda County",
    
    # Annador -> Amador County
    location_clean == "annador county" ~ "Amador County",
    
    # Gentry variants -> Gentry County
    location_clean %in% c("gentry", "gentry county") ~ "Gentry County",
    
    # 3. MAPPINGS & ABBREVIATIONS
    location_clean %in% c("mo", "mo .", "mo . .", "weston mo") ~ "Missouri",
    location_clean %in% c("ky", "ky .")            ~ "Kentucky",
    location_clean %in% c("ill", "ill . .")        ~ "Illinois",
    location_clean %in% c("la .")                  ~ "Louisiana",
    location_clean %in% c("n.c")                   ~ "North Carolina",
    
    # 4. TYPO FIXES (Specific cases)
    location_clean == "san francisco"              ~ "San Francisco",
    location_clean == "los angeles"                ~ "Los Angeles",
    location_clean == "santa barbara"              ~ "Santa Barbara",
    location_clean == "red bluffs"                 ~ "Red Bluffs",
    location_clean == "san luis obispo"            ~ "San Luis Obispo",
    location_clean == "salmon river"               ~ "Salmon River",
    location_clean == "sawyer 's bar"              ~ "Sawyer's Bar",
    
    # 5. DEFAULT: Standardize everything else to Title Case
    TRUE ~ str_to_title(location_clean)
  )) |> 
  
  ### Extract state from multi-part locations BEFORE state classification
  mutate(
    # Extract last word if it's a state name
    last_word = word(location_clean, -1),
    # Extract second-to-last word
    second_last_word = word(location_clean, -2),
    
    # If last word is a state, extract it
    extracted_state = case_when(
      last_word %in% c("Alabama", "Arkansas", "California", "Colorado", "Delaware", 
                       "Florida", "Georgia", "Idaho", "Illinois", "Indiana", "Iowa", 
                       "Kansas", "Kentucky", "Louisiana", "Minnesota", "Mississippi", 
                       "Missouri", "Nebraska", "New", "North", "Ohio", "Oregon", 
                       "Tennessee", "Texas", "Virginia", "Wisconsin") ~ last_word,
      # Handle "New York", "North Carolina"
      paste(second_last_word, last_word) %in% c("New York", "New Hampshire", 
                                                  "North Carolina") ~ paste(second_last_word, last_word),
      TRUE ~ NA_character_
    ),
    
    # Clean location by removing the state suffix
    location_clean = if_else(
      !is.na(extracted_state),
      str_trim(str_remove(location_clean, paste0("\\s*", extracted_state, "$"))),
      location_clean
    )
  ) |>
  
  ### States from localities
  mutate(state = case_when(
    # Use extracted state first
    !is.na(extracted_state) ~ extracted_state,
    
    # Already state names - pass through
    location_clean %in% c("Alabama", "Arkansas", "California", "Colorado", "Delaware", 
                          "Florida", "Georgia", "Idaho", "Illinois", "Indiana", "Iowa", 
                          "Kansas", "Kentucky", "Louisiana", "Minnesota", "Mississippi", 
                          "Missouri", "Nebraska", "New Hampshire", "New York", "North Carolina", 
                          "Ohio", "Oregon", "Tennessee", "Texas", "Virginia", "Wisconsin") ~ location_clean,
    
    # California counties and cities
    location_clean %in% c("Alameda", "Alameda County", "Annador County", "Calaveras County",
                          "Merced County", "Napa", "Oakland", "Placer County", 
                          "San Andreas", "San Antonio", "San Felipe", "San Francisco",
                          "San Gabriel", "San Isidro", "San Jose", "San Luis Obispo",
                          "Santa Barbara", "Shasta", "Solano County", 
                          "Stockton", "Tejon", "Tuolumne County", "Yuba County",
                          "Red Bluffs", "Drytown", "Iowa Hill", "Monterey",
                          "Sawyer's Bar", "South Sulphur", "Los Angeles", "Sacramento", 
                          "Yreka", "Verandah", "Columbia", "Pancho Daniel") ~ "California",
    str_detect(location_clean, "Tuolumne") ~ "California",
    
    # Texas counties and cities
    location_clean %in% c("Uvalde County", "El Paso", "San Antonio") ~ "Texas",
    
    # Missouri counties and cities
    location_clean %in% c("Buchanan County", "Gentry County", "Grand Glaize Jackson County", 
                          "Renick Randolph County") ~ "Missouri",
    str_detect(location_clean, "Jackson County") ~ "Missouri",
    
    # Illinois counties and cities
    location_clean %in% c("Ogle County", "Keithsburg", "Rockford", "Shelbyville") ~ "Illinois",
    
    # Indiana counties and cities
    location_clean %in% c("South Bend") ~ "Indiana",
    
    # Mississippi counties and cities
    location_clean %in% c("Issaquena County", "Panola", "Pontotoc County") ~ "Mississippi",
    
    # Iowa counties and cities
    location_clean %in% c("Black Hawk", "Tama County", "Bargar") ~ "Iowa",
    
    # Kentucky cities
    location_clean %in% c("Lexington", "Louisville", "Williamstown Grant County") ~ "Kentucky",
    str_detect(location_clean, "Grant County") ~ "Kentucky",
    
    # Tennessee cities
    location_clean %in% c("Memphis", "Nashville") ~ "Tennessee",
    
    # Georgia cities
    location_clean %in% c("Savannah", "Tuskegee") ~ "Georgia",
    
    # Alabama counties and cities
    location_clean %in% c("Pike County") ~ "Alabama",
    
    # Louisiana cities
    location_clean %in% c("Baton Rouge", "New Orleans") ~ "Louisiana",
    
    # Kansas cities
    location_clean %in% c("Lecompton", "Leavenworth City", "Kansas Territory") ~ "Kansas",
    
    # Florida cities
    location_clean %in% c("Pensacola", "Alston", "Milton") ~ "Florida",
    
    # Nebraska cities
    location_clean %in% c("Omaha City") ~ "Nebraska",
    
    # Colorado cities
    location_clean == "Denver City" ~ "Colorado",
    
    # Wisconsin
    location_clean %in% c("Washara County", "Geneva") ~ "Wisconsin",
    
    # New York
    location_clean == "Bushville" ~ "New York",
    
    # Oregon counties
    location_clean %in% c("Lewis County", "Rogue River", "Rogue River Valley") ~ "Oregon",
    
    # North Carolina cities
    location_clean == "Wilmington" ~ "North Carolina",
    
    # District of Columbia
    location_clean == "Washington" ~ "District of Columbia",
    
    # Rivers/geographic features
    location_clean == "Mississippi" ~ "Mississippi",
    location_clean == "Upper Mississippi River" ~ "Multiple States",
    location_clean == "Salmon River" ~ "Idaho",
    
    # Uncertain/generic locations
    location_clean %in% c("Driskells", "Partridge", "Scaggs", 
                          "Shattuck", "Weatherford", "Woodstock",
                          "Lawrenceville", "Union City", "Petersburg", 
                          "Essex County", "Melbourne", "Mount Alexander",
                          "Anastasia Jesus") ~ "Unknown",
    
    # Default
    TRUE ~ "Unknown"
  )) |>
  select(-last_word, -second_last_word, -extracted_state) |> 
  ungroup() |> 
  group_by(doc_id) |>
  filter(
    # Keep this row if either:
    # 1. It has a known state, OR
    # 2. All rows for this doc_id have Unknown state
    state != "Unknown" | all(state == "Unknown")
  ) |>
  # If still multiple rows (multiple known states), keep first
  slice(1) |>
  ungroup()
```

# fix missing states 

```{r}

docs_states <- new_location  |> 
  select(doc_id, state)

precivil_complete <- precivil_complete |> 
  left_join(docs_states, by=c("index"="doc_id"))

```



```{r}
precivil_complete <- precivil_complete |>
mutate(state = case_when(
    index == "1015" ~ "Iowa",
    index == "1016" ~ "Mississippi",
    index == "1512" ~ "California",
    index == "1513" ~ "Illinois",
    index == "7522" ~ "California",
    index == "16747" ~ "Mississippi",
    index == "16748" ~ "Mississippi",
    index == "16751" ~ "South Carolina",
    index == "16756" ~ "Mississippi",
    index == "16774" ~ "Maryland",
    index == "16791" ~ "Arkansas",
    index == "16792" ~ "Illinois",
    index == "16794" ~ "Illinois",
    index == "16795" ~ "Kentucky",
    index == "16804" ~ "Ohio",
    index == "16808" ~ "Texas",
    index == "16812" ~ "Tennessee",
    index == "16822" ~ "Kentucky",
    index == "preciv1" ~ "District of Columbia",
    index == "preciv2" ~ "South Carolina",
    index == "preciv5" ~ "Nevada",
    index == "preciv7" ~ "Delaware",
    index == "preciv8" ~ "Delaware",
    index == "preciv9" ~ "California",
    index == "preciv398" ~ "North Carolina",
    index == "preciv10" ~ "Alabama",
    index == "preciv13" ~ "New York",
    index == "preciv18" ~ "Arkansas",
    index == "preciv19" ~ "South Carolina",
    index == "preciv20" ~ "New Mexico",
    index == "preciv21" ~ "Oregon",
    index == "preciv22" ~ "Colorado",
    index == "preciv24" ~ "Mississippi",
    index == "preciv27" ~ "Georgia",
    index == "preciv29" ~ "Kansas",
    index == "preciv31" ~ "Arizona",
    index == "preciv32" ~ "Arkansas",
    index == "preciv34" ~ "Michigan",
    index == "preciv36" ~ "Tennessee",
    index == "preciv37" ~ "California",
    index == "preciv38" ~ "Texas",
    index == "preciv39" ~ "Nevada",
    index == "preciv40" ~ "Arkansas",
    index == "preciv41" ~ "Colorado",
    index == "preciv42" ~ "Mississippi",
    index == "preciv43" ~ "South Carolina",
    index == "preciv45" ~ "South Carolina",
    index == "preciv46" ~ "California",
    index == "preciv47" ~ "California",
    index == "preciv52" ~ "Virginia",
    index == "preciv53" ~ "Georgia",
    index == "preciv54" ~ "Ohio",
    index == "preciv57" ~ "South Carolina",
    index == "preciv60" ~ "Missouri",
    index == "preciv63" ~ "Arizona",
    index == "preciv66" ~ "West Virginia",
    index == "preciv67" ~ "Kansas",
    index == "preciv68" ~ "Kansas",
    index == "preciv70" ~ "Kansas",
    index == "preciv71" ~ "California",
    index == "preciv73" ~ "California",
    index == "preciv75" ~ "Kentucky",
    index == "preciv76" ~ "Mississippi",
    index == "preciv78" ~ "Illinois",
    index == "preciv79" ~ "California",
    index == "preciv82" ~ "Arkansas",
    index == "preciv84" ~ "Nevada",
    index == "preciv86" ~ "Arkansas",
    index == "preciv87" ~ "Illinois",
    index == "preciv88" ~ "Virginia",
    index == "preciv89" ~ "Minnesota",
    index == "preciv90" ~ "Minnesota",
    index == "preciv91" ~ "Tennessee",
    index == "preciv92" ~ "California",
    index == "preciv93" ~ "Kansas",
    index == "preciv94" ~ "Illinois",
    index == "preciv95" ~ "Colorado",
    index == "preciv97" ~ "Texas",
    index == "preciv99" ~ "Minnesota",
    index == "preciv100" ~ "California",
    index == "preciv101" ~ "California",
    index == "preciv102" ~ "Kentucky",
    index == "preciv103" ~ "California",
    index == "preciv105" ~ "Minnesota",
    index == "preciv107" ~ "Virginia",
    index == "preciv109" ~ "Minnesota",
    index == "preciv110" ~ "California",
    index == "preciv112" ~ "Kentucky",
    index == "preciv115" ~ "Delaware",
    index == "preciv118" ~ "California",
    index == "preciv119" ~ "Kentucky",
    index == "preciv120" ~ "California",
    index == "preciv121" ~ "California",
    index == "preciv122" ~ "Missouri",
    index == "preciv124" ~ "California",
    index == "preciv126" ~ "Maryland",
    index == "preciv132" ~ "California",
    index == "preciv133" ~ "California",
    index == "preciv134" ~ "Illinois",
    index == "preciv136" ~ "California",
    index == "preciv137" ~ "Delaware",
    index == "preciv143" ~ "California",
    index == "preciv146" ~ "California",
    index == "preciv147" ~ "California",
    index == "preciv151" ~ "California",
    index == "preciv152" ~ "California",
    index == "preciv155" ~ "Kansas",
    index == "preciv157" ~ "California",
    index == "preciv158" ~ "California",
    index == "preciv159" ~ "California",
    index == "preciv160" ~ "Iowa",
    index == "preciv162" ~ "Wisconsin",
    index == "preciv164" ~ "Louisiana",
    index == "preciv166" ~ "Kansas",
    index == "preciv167" ~ "Texas",
    index == "preciv170" ~ "Kansas",
    index == "preciv173" ~ "Nebraska",
    index == "preciv174" ~ "Mississippi",
    index == "preciv175" ~ "Kentucky",
    index == "preciv177" ~ "California",
    index == "preciv179" ~ "Kansas",
    index == "preciv180" ~ "California",
    index == "preciv181" ~ "Virginia",
    index == "preciv184" ~ "Nebraska",
    index == "preciv185" ~ "Kansas",
    index == "preciv187" ~ "California",
    index == "preciv189" ~ "California",
    index == "preciv192" ~ "California",
    index == "preciv193" ~ "Alabama",
    index == "preciv195" ~ "Illinois",
    index == "preciv196" ~ "District of Columbia",
    index == "preciv198" ~ "California",
    index == "preciv199" ~ "California",
    index == "preciv202" ~ "California",
    index == "preciv203" ~ "California",
    index == "preciv205" ~ "Texas",
    index == "preciv206" ~ "California",
    index == "preciv207" ~ "Arkansas",
    index == "preciv208" ~ "Virginia",
    index == "preciv209" ~ "California",
    index == "preciv210" ~ "California",
    index == "preciv213" ~ "California",
    index == "preciv215" ~ "California",
    index == "preciv216" ~ "California",
    index == "preciv217" ~ "California",
    index == "preciv218" ~ "California",
    index == "preciv219" ~ "California",
    index == "preciv221" ~ "California",
    index == "preciv223" ~ "California",
    index == "preciv224" ~ "California",
    index == "preciv226" ~ "Arkansas",
    index == "preciv228" ~ "California",
    index == "preciv229" ~ "Arkansas",
    index == "preciv230" ~ "California",
    index == "preciv231" ~ "California",
    index == "preciv233" ~ "Kansas",
    index == "preciv235" ~ "California",
    index == "preciv237" ~ "California",
    index == "preciv238" ~ "Texas",
    index == "preciv239" ~ "Massachusetts",
    index == "preciv240" ~ "Missouri",
    index == "preciv243" ~ "California",
    index == "preciv246" ~ "California",
    index == "preciv247" ~ "Indiana",
    index == "preciv248" ~ "Missouri",
    index == "preciv250" ~ "Missouri",
    index == "preciv253" ~ "California",
    index == "preciv255" ~ "Missouri",
    index == "preciv256" ~ "Missouri",
    index == "preciv264" ~ "California",
    index == "preciv266" ~ "California",
    index == "preciv267" ~ "California",
    index == "preciv268" ~ "California",
    index == "preciv269" ~ "California",
    index == "preciv270" ~ "California",
    index == "preciv271" ~ "California",
    index == "preciv273" ~ "California",
    index == "preciv274" ~ "California",
    index == "preciv276" ~ "Arkansas",
    index == "preciv277" ~ "California",
    index == "preciv279" ~ "Missouri",
    index == "preciv280" ~ "California",
    index == "preciv282" ~ "Missouri",
    index == "preciv283" ~ "California",
    index == "preciv284" ~ "Illinois",
    index == "preciv287" ~ "Texas",
    index == "preciv288" ~ "Illinois",
    index == "preciv289" ~ "Tennessee",
    index == "preciv292" ~ "Iowa",
    index == "preciv293" ~ "California",
    index == "preciv294" ~ "California",
    index == "preciv297" ~ "California",
    index == "preciv298" ~ "Louisiana",
    index == "preciv300" ~ "California",
    index == "preciv301" ~ "Louisiana",
    index == "preciv302" ~ "California",
    index == "preciv303" ~ "California",
    index == "preciv304" ~ "New Mexico",
    index == "preciv306" ~ "California",
    index == "preciv308" ~ "California",
    index == "preciv309" ~ "California",
    index == "preciv310" ~ "Missouri",
    index == "preciv312" ~ "California",
    index == "preciv314" ~ "California",
    index == "preciv317" ~ "California",
    index == "preciv319" ~ "California",
    index == "preciv321" ~ "California",
    index == "preciv322" ~ "California",
    index == "preciv323" ~ "California",
    index == "preciv324" ~ "California",
    index == "preciv325" ~ "California",
    index == "preciv326" ~ "California",
    index == "preciv327" ~ "California",
    index == "preciv331" ~ "California",
    index == "preciv333" ~ "Florida",
    index == "preciv334" ~ "California",
    index == "preciv337" ~ "California",
    index == "preciv338" ~ "California",
    index == "preciv339" ~ "Virginia",
    index == "preciv342" ~ "California",
    index == "preciv344" ~ "Texas",
    index == "preciv349" ~ "California",
    index == "preciv350" ~ "California",
    index == "preciv351" ~ "California",
    index == "preciv352" ~ "California",
    index == "preciv353" ~ "District of Columbia",
    index == "preciv354" ~ "California",
    index == "preciv357" ~ "Florida",
    index == "preciv359" ~ "California",
    index == "preciv360" ~ "California",
    index == "preciv361" ~ "California",
    index == "preciv362" ~ "California",
    index == "preciv363" ~ "California",
    index == "preciv364" ~ "California",
    index == "preciv366" ~ "California",
    index == "preciv367" ~ "California",
    index == "preciv368" ~ "California",
    index == "preciv369" ~ "California",
    index == "preciv370" ~ "California",
    index == "preciv371" ~ "Kentucky",
    index == "preciv373" ~ "Maryland",
    index == "preciv375" ~ "Kentucky",
    index == "preciv377" ~ "Texas",
    index == "preciv378" ~ "Virginia",
    index == "preciv381" ~ "California",
    index == "preciv382" ~ "California",
    index == "preciv383" ~ "Tennessee",
    index == "preciv384" ~ "Mississippi",
    index == "preciv385" ~ "Arkansas",
    index == "preciv385" ~ "Illinois",
    index == "preciv387" ~ "Arkansas",
    index == "preciv387" ~ "New Jersey",
    index == "preciv390" ~ "Illinois",
    index == "preciv391" ~ "Texas",
    index == "preciv392" ~ "Arkansas",
    index == "preciv393" ~ "Illinois",
    index == "preciv394" ~ "Illinois",
    index == "preciv395" ~ "Arkansas",
    index == "preciv396" ~ "Illinois",
    index == "preciv397" ~ "Illinois",
    index == "preciv398" ~ "Missouri",
    index == "preciv399" ~ "New York",
    index == "preciv400" ~ "New York",
    index == "preciv402" ~ "Louisiana",
    index == "preciv403" ~ "Florida",
    index == "preciv404" ~ "Florida",
    index == "preciv404" ~ "North Carolina",
    index == "preciv408" ~ "Mississippi",
    index == "preciv409" ~ "Maryland",
    index == "preciv410" ~ "Pennsylvania",
    index == "preciv412" ~ "Texas",
    index == "preciv415" ~ "Illinois",
    index == "preciv416" ~ "Illinois",
    index == "preciv417" ~ "Mississippi",
    index == "preciv418" ~ "Missouri",
    index == "preciv419" ~ "Arkansas", 
    index == "preciv420" ~ "Massachusetts",
    index == "preciv422" ~ "Alabama", 
    index == "preciv423" ~ "Missouri", 
    index == "preciv424" ~ "Missouri", 
    index == "preciv425" ~ "Mississippi", 
    index == "preciv426" ~ "Arkansas",   
    index == "preciv427" ~ "Kentucky",  
    index == "preciv429" ~ "Mississippi", 
    index == "preciv430" ~ "Mississippi", 
    index == "preciv432" ~ "Mississippi",
    index == "preciv435" ~ "Kentucky", 
    index == "preciv439" ~ "Louisiana", 
    index == "preciv440" ~ "Massachusetts", 
    index == "preciv441" ~ "Alabama", 
    index == "preciv442" ~ "Kentucky", 
    index == "preciv443" ~ "Alabama", 
    index == "preciv446" ~ "Mississippi",
    TRUE ~ state  # Keep existing state value if index not in list
  ))




```


### Validate and check total

```{r}
new_total <- precivil_complete |> 
  count(state) |> 
  arrange(desc(n))
#-- jan 5: 26 states, calif 41 reports. 147 reports total
# jan 6: 36 states, calif 135
# jan 7: 39 states, calif 151
```


### write precivil_complete_jan_7_update.csv
```{r}

#write_csv(precivil_complete, "../data/precivil_complete_jan_7_update.csv")

```


```{r}

dt_total <- new_total |> 
  filter(!is.na(state)) |> 
  rename(State = state, Total = n)

datatable(dt_total,
          # Use HTML styling directly in the caption string
          caption = htmltools::tags$caption(
            style = 'caption-side: top; text-align: center; color: black; font-size: 20pt; font-weight: bold;',
            'Top locations for pre-Civil War lynchings'
          ),
          options = list(
            pageLength = 50,
            searching = FALSE, 
            lengthChange = FALSE
          ))
```


### Map output
```{r}
map <- dt_total
write_csv(map, "../output/map.csv")

```

#-----------------------------------------------------------------

### Quick view of any entry by row
```{r}
wrapped_text <- strwrap(as.character(precivil_complete$spellchecked[426]), width = 60)
writeLines(wrapped_text, "temp_text.txt")
file.edit("temp_text.txt")
```


### Print out articles for verification
```{r}
new_text_location <- new_location |> 
  right_join(precivil_complete, by = c("doc_id" = "index")) |> 
  # For each doc_id, remove locations that are substrings of other locations
  group_by(doc_id) |>
  arrange(doc_id, desc(nchar(location_clean))) |>  # Longest first
  filter(!duplicated(location_clean)) |>  # Remove exact duplicates first
  mutate(
    # Check if this location is a substring of any longer location in same doc
    is_substring = sapply(location_clean, function(loc) {
      if (is.na(loc)) return(FALSE)
      others <- location_clean[location_clean != loc & !is.na(location_clean)]
      any(str_detect(others, fixed(loc)))
    })
  ) |>
  filter(!is_substring) |>  # Remove substrings
  select(-is_substring) |> 
  rename(state = state.y) |> 
  select(-state.x)

# Now create the text file with locations highlighted
file_conn <- file("verify_states.txt", "w")

for (i in 1:nrow(new_text_location)) {
  row_data <- new_text_location[i, ]
  
  writeLines(paste0("Document ID: ", row_data$doc_id), file_conn)
  writeLines(paste0("Sentence ID: ", row_data$sentence_id), file_conn)
  writeLines(paste0("State: ", row_data$state), file_conn)
  writeLines(paste0("Location: ", row_data$location), file_conn)
  # writeLines(paste0("Entity Type: ", row_data$entity_types), file_conn)
  writeLines(paste0("Headline: ", row_data$headline), file_conn)
  writeLines(paste0("Newspaper: ", row_data$newspaper_name), file_conn)
  writeLines(paste0("Date: ", row_data$date), file_conn)

  
  # Wrap the full text
  wrapped_text <- strwrap(row_data$spellchecked, width = 80)
  wrapped_text[1] <- paste0("Full Context: ", wrapped_text[1])
  writeLines(wrapped_text, file_conn)
  
  if (i < nrow(new_text_location)) {
    writeLines("\n==================================================\n", file_conn)
  }
}

close(file_conn)
```
