---
title: "Untitled"
author: "Wells"
date: "2026-01-05"
output: html_document
---
#KWIC

```{r}

# Create a corpus using the 'article_spellchecked' column as the text field
my_corpus <- corpus(precivil_complete, text_field = "spellchecked", docid_field = "index") # build a new 
# Tokenize the corpus
my_tokens <- tokens(my_corpus)

#15 tokens on both sides of the search term
quanteda_df_llm <- kwic(my_tokens, phrase(c("lynch")), window = 15, valuetype = "regex") %>% as.data.frame() |> 
  distinct(docname,  .keep_all = TRUE)

quanteda_df_llm <- kwic(my_tokens, phrase(c("lynch|hung|hang(ed|ing)?|executed?|execution|mob\\s+law|vigilance\\s+committee")), window = 20, valuetype = "regex") %>% as.data.frame() |> 
  distinct(docname,  .keep_all = TRUE)

df_llm <- precivil_complete |> 
  inner_join(quanteda_df_llm, by=c("index"="docname")) |> 
  mutate(kwic_text = paste0(pre," ", keyword," ",post)) |> 
   select(-c(from, to,pre,post, keyword, pattern)) |> 
  mutate(word_count_select = str_count(select_text, "\\w+"))
#447 processed articles!
#write_csv(df_llm, "../data/precivil_df_llm_dec_26.csv")

```

## Spacy parsing
```{r}

# Initialize spacy (one-time setup required: spacy_install())
spacy_initialize(model = "en_core_web_sm")
```

```{r}
text <- data.frame(
  doc_id = precivil_complete$index,
  text = precivil_complete$spellchecked,
  stringsAsFactors = FALSE
)

# Parse with spacyr
parsed <- spacy_parse(text, entity = TRUE, 
                               lemma = TRUE,
                               pos = TRUE,
                               tag= TRUE,  
                              dependency = TRUE  )


```


```{r}
locations_with_prep <- parsed %>%
  filter(str_detect(entity, "^(GPE|LOC|FAC)_(B|I)$")) %>%
  group_by(doc_id, sentence_id) %>%
  mutate(prev_token = lag(token)) %>%
  ungroup() %>%
  inner_join(lynch_info, by = c("doc_id", "sentence_id"), relationship = "many-to-many") %>%
  filter(prev_token %in% c("at", "in", "near", "on", "from", "to") |
         (token_id >= lynch_token_id - 5 & token_id < lynch_token_id) |
         token_id > lynch_token_id) %>%
  # Remove duplicates!
  distinct(doc_id, sentence_id, token_id, .keep_all = TRUE)
```



```{r}
location_counts2 <- locations_with_prep %>%
  arrange(doc_id, sentence_id, token_id) %>%
  group_by(doc_id, sentence_id) %>%
  # Combine ALL consecutive locations
  summarise(
    location = paste(token, collapse = " "),
    .groups = "drop"
  )
```

```{r}
Wlocation_counts2 <- locations_with_prep %>%
  arrange(doc_id, sentence_id, token_id) %>%
  group_by(doc_id, sentence_id) %>%
  mutate(
    # Check if this token is consecutive with the previous one
    is_consecutive = (token_id - lag(token_id, default = -999)) == 1,
    # Create merged groups: start new group if NOT consecutive
    entity_group_merged = cumsum(!is_consecutive)
  ) %>%
  group_by(doc_id, sentence_id, entity_group_merged) %>%
  summarise(
    location = paste(token, collapse = " "),
    .groups = "drop"
  ) |> 
  arrange(desc(doc_id))
```


#new classifier

```{r}
new_location <- Wlocation_counts2 |> 
  # right_join(precivil_df, by = c("doc_id" = "index")) |> 
mutate(location_clean = tolower(location)) |> 
  # Remove duplicate consecutive words first
  mutate(location_clean = str_replace_all(location_clean, "\\b(\\w+)\\s+\\1\\b", "\\1")) |>
  mutate(location_clean = case_when(
    # 1. REMOVE NOISE (Returns NA)
    location_clean %in% c("a", "the memphis enquirer of the", "the yoakum gang", 
                          "rev", "lieut", "jas", "sheriff", "fort", "5th", 
                          "5th september september", "states", "west", "red",
                          "a senor senor badlo badlo", "san", "driskells", "los") ~ NA_character_,
    
    # 2. SPECIFIC COMBINATIONS & CORRECTIONS
    # Rogue River variants -> Rogue River Valley
    location_clean %in% c("rogue", "rogue river", "rogue river valley") ~ "Rogue River Valley",
    
    # Alameda -> Alameda County
    location_clean == "alameda" ~ "Alameda County",
    
    # Annador -> Amador County
    location_clean == "annador county" ~ "Amador County",
    
    # Gentry variants -> Gentry County
    location_clean %in% c("gentry", "gentry county") ~ "Gentry County",
    
    # 3. MAPPINGS & ABBREVIATIONS
    location_clean %in% c("mo", "mo .", "mo . .", "weston mo") ~ "Missouri",
    location_clean %in% c("ky", "ky .")            ~ "Kentucky",
    location_clean %in% c("ill", "ill . .")        ~ "Illinois",
    location_clean %in% c("la .")                  ~ "Louisiana",
    location_clean %in% c("n.c")                   ~ "North Carolina",
    
    # 4. TYPO FIXES (Specific cases)
    location_clean == "san francisco"              ~ "San Francisco",
    location_clean == "los angeles"                ~ "Los Angeles",
    location_clean == "santa barbara"              ~ "Santa Barbara",
    location_clean == "red bluffs"                 ~ "Red Bluffs",
    location_clean == "san luis obispo"            ~ "San Luis Obispo",
    location_clean == "salmon river"               ~ "Salmon River",
    location_clean == "sawyer 's bar"              ~ "Sawyer's Bar",
    
    # 5. DEFAULT: Standardize everything else to Title Case
    TRUE ~ str_to_title(location_clean)
  )) |> 
  
  ### Extract state from multi-part locations BEFORE state classification
  mutate(
    # Extract last word if it's a state name
    last_word = word(location_clean, -1),
    # Extract second-to-last word
    second_last_word = word(location_clean, -2),
    
    # If last word is a state, extract it
    extracted_state = case_when(
      last_word %in% c("Alabama", "Arkansas", "California", "Colorado", "Delaware", 
                       "Florida", "Georgia", "Idaho", "Illinois", "Indiana", "Iowa", 
                       "Kansas", "Kentucky", "Louisiana", "Minnesota", "Mississippi", 
                       "Missouri", "Nebraska", "New", "North", "Ohio", "Oregon", 
                       "Tennessee", "Texas", "Virginia", "Wisconsin") ~ last_word,
      # Handle "New York", "North Carolina"
      paste(second_last_word, last_word) %in% c("New York", "New Hampshire", 
                                                  "North Carolina") ~ paste(second_last_word, last_word),
      TRUE ~ NA_character_
    ),
    
    # Clean location by removing the state suffix
    location_clean = if_else(
      !is.na(extracted_state),
      str_trim(str_remove(location_clean, paste0("\\s*", extracted_state, "$"))),
      location_clean
    )
  ) |>
  
  ### States from localities
  mutate(state = case_when(
    # Use extracted state first
    !is.na(extracted_state) ~ extracted_state,
    
    # Already state names - pass through
    location_clean %in% c("Alabama", "Arkansas", "California", "Colorado", "Delaware", 
                          "Florida", "Georgia", "Idaho", "Illinois", "Indiana", "Iowa", 
                          "Kansas", "Kentucky", "Louisiana", "Minnesota", "Mississippi", 
                          "Missouri", "Nebraska", "New Hampshire", "New York", "North Carolina", 
                          "Ohio", "Oregon", "Tennessee", "Texas", "Virginia", "Wisconsin") ~ location_clean,
    
    # California counties and cities
    location_clean %in% c("Alameda", "Alameda County", "Annador County", "Calaveras County",
                          "Merced County", "Napa", "Oakland", "Placer County", 
                          "San Andreas", "San Antonio", "San Felipe", "San Francisco",
                          "San Gabriel", "San Isidro", "San Jose", "San Luis Obispo",
                          "Santa Barbara", "Shasta", "Solano County", 
                          "Stockton", "Tejon", "Tuolumne County", "Yuba County",
                          "Red Bluffs", "Drytown", "Iowa Hill", "Monterey",
                          "Sawyer's Bar", "South Sulphur", "Los Angeles", "Sacramento", 
                          "Yreka", "Verandah", "Columbia", "Pancho Daniel") ~ "California",
    str_detect(location_clean, "Tuolumne") ~ "California",
    
    # Texas counties and cities
    location_clean %in% c("Uvalde County", "El Paso", "San Antonio") ~ "Texas",
    
    # Missouri counties and cities
    location_clean %in% c("Buchanan County", "Gentry County", "Grand Glaize Jackson County", 
                          "Renick Randolph County") ~ "Missouri",
    str_detect(location_clean, "Jackson County") ~ "Missouri",
    
    # Illinois counties and cities
    location_clean %in% c("Ogle County", "Keithsburg", "Rockford", "Shelbyville") ~ "Illinois",
    
    # Indiana counties and cities
    location_clean %in% c("South Bend") ~ "Indiana",
    
    # Mississippi counties and cities
    location_clean %in% c("Issaquena County", "Panola", "Pontotoc County") ~ "Mississippi",
    
    # Iowa counties and cities
    location_clean %in% c("Black Hawk", "Tama County", "Bargar") ~ "Iowa",
    
    # Kentucky cities
    location_clean %in% c("Lexington", "Louisville", "Williamstown Grant County") ~ "Kentucky",
    str_detect(location_clean, "Grant County") ~ "Kentucky",
    
    # Tennessee cities
    location_clean %in% c("Memphis", "Nashville") ~ "Tennessee",
    
    # Georgia cities
    location_clean %in% c("Savannah", "Tuskegee") ~ "Georgia",
    
    # Alabama counties and cities
    location_clean %in% c("Pike County") ~ "Alabama",
    
    # Louisiana cities
    location_clean %in% c("Baton Rouge", "New Orleans") ~ "Louisiana",
    
    # Kansas cities
    location_clean %in% c("Lecompton", "Leavenworth City", "Kansas Territory") ~ "Kansas",
    
    # Florida cities
    location_clean %in% c("Pensacola", "Alston", "Milton") ~ "Florida",
    
    # Nebraska cities
    location_clean %in% c("Omaha City") ~ "Nebraska",
    
    # Colorado cities
    location_clean == "Denver City" ~ "Colorado",
    
    # Wisconsin
    location_clean %in% c("Washara County", "Geneva") ~ "Wisconsin",
    
    # New York
    location_clean == "Bushville" ~ "New York",
    
    # Oregon counties
    location_clean %in% c("Lewis County", "Rogue River", "Rogue River Valley") ~ "Oregon",
    
    # North Carolina cities
    location_clean == "Wilmington" ~ "North Carolina",
    
    # District of Columbia
    location_clean == "Washington" ~ "District of Columbia",
    
    # Rivers/geographic features
    location_clean == "Mississippi" ~ "Mississippi",
    location_clean == "Upper Mississippi River" ~ "Multiple States",
    location_clean == "Salmon River" ~ "Idaho",
    
    # Uncertain/generic locations
    location_clean %in% c("Driskells", "Partridge", "Scaggs", 
                          "Shattuck", "Weatherford", "Woodstock",
                          "Lawrenceville", "Union City", "Petersburg", 
                          "Essex County", "Melbourne", "Mount Alexander",
                          "Anastasia Jesus") ~ "Unknown",
    
    # Default
    TRUE ~ "Unknown"
  )) |>
  select(-last_word, -second_last_word, -extracted_state) |> 
  ungroup() |> 
  group_by(doc_id) |>
  filter(
    # Keep this row if either:
    # 1. It has a known state, OR
    # 2. All rows for this doc_id have Unknown state
    state != "Unknown" | all(state == "Unknown")
  ) |>
  # If still multiple rows (multiple known states), keep first
  slice(1) |>
  ungroup()
```

# fix missing states 
```{r}
x <- Xprecivil_complete |>
  rename(doc_id = index) |>
#   
# x <- new_location|> 
mutate(state = case_when(
    doc_id == "1015" ~ "Iowa",
    doc_id == "1016" ~ "Mississippi",
    doc_id == "1512" ~ "California",
    doc_id == "1513" ~ "Illinois",
    doc_id == "7522" ~ "California",
    doc_id == "16747" ~ "Mississippi",
    doc_id == "16748" ~ "Mississippi",
    doc_id == "16751" ~ "South Carolina",
    doc_id == "16756" ~ "Mississippi",
    doc_id == "16774" ~ "Maryland",
    doc_id == "16791" ~ "Arkansas",
    doc_id == "16792" ~ "Illinois",
    doc_id == "16794" ~ "Illinois",
    doc_id == "16795" ~ "Kentucky",
    doc_id == "16804" ~ "Ohio",
    doc_id == "16808" ~ "Texas",
    doc_id == "16812" ~ "Tennessee",
    doc_id == "16822" ~ "Kentucky",
    doc_id == "preciv1" ~ "District of Columbia",
    doc_id == "preciv2" ~ "South Carolina",
    doc_id == "preciv5" ~ "Nevada",
    doc_id == "preciv7" ~ "Delaware",
    doc_id == "preciv8" ~ "Delaware",
    doc_id == "preciv9" ~ "California",
    doc_id == "preciv398" ~ "North Carolina",
    doc_id == "preciv10" ~ "Alabama",
    doc_id == "preciv13" ~ "New York",
    doc_id == "preciv18" ~ "Arkansas",
    doc_id == "preciv19" ~ "South Carolina",
    doc_id == "preciv20" ~ "New Mexico",
    doc_id == "preciv21" ~ "Oregon",
    doc_id == "preciv22" ~ "Colorado",
    doc_id == "preciv24" ~ "Mississippi",
    doc_id == "preciv27" ~ "Georgia",
    doc_id == "preciv29" ~ "Kansas",
    doc_id == "preciv31" ~ "Arizona",
    doc_id == "preciv32" ~ "Arkansas",
    doc_id == "preciv34" ~ "Michigan",
    doc_id == "preciv36" ~ "Tennessee",
    doc_id == "preciv37" ~ "California",
    doc_id == "preciv38" ~ "Texas",
    doc_id == "preciv39" ~ "Nevada",
    doc_id == "preciv40" ~ "Arkansas",
    doc_id == "preciv41" ~ "Colorado",
    doc_id == "preciv42" ~ "Mississippi",
    doc_id == "preciv43" ~ "South Carolina",
    doc_id == "preciv45" ~ "South Carolina",
    doc_id == "preciv46" ~ "California",
    doc_id == "preciv47" ~ "California",
    doc_id == "preciv52" ~ "Virginia",
    doc_id == "preciv53" ~ "Georgia",
    doc_id == "preciv54" ~ "Ohio",
    doc_id == "preciv57" ~ "South Carolina",
    doc_id == "preciv60" ~ "Missouri",
    doc_id == "preciv63" ~ "Arizona",
    doc_id == "preciv66" ~ "West Virginia",
    doc_id == "preciv67" ~ "Kansas",
    doc_id == "preciv68" ~ "Kansas",
    doc_id == "preciv70" ~ "Kansas",
    doc_id == "preciv71" ~ "California",
    doc_id == "preciv73" ~ "California",
    doc_id == "preciv75" ~ "Kentucky",
    doc_id == "preciv76" ~ "Mississippi",
    doc_id == "preciv78" ~ "Illinois",
    doc_id == "preciv79" ~ "California",
    doc_id == "preciv82" ~ "Arkansas",
    doc_id == "preciv84" ~ "Nevada",
    doc_id == "preciv86" ~ "Arkansas",
    doc_id == "preciv87" ~ "Illinois",
    doc_id == "preciv88" ~ "Virginia",
    doc_id == "preciv89" ~ "Minnesota",
    doc_id == "preciv90" ~ "Minnesota",
    doc_id == "preciv91" ~ "Tennessee",
    doc_id == "preciv92" ~ "California",
    doc_id == "preciv93" ~ "Kansas",
    doc_id == "preciv94" ~ "Illinois",
    doc_id == "preciv95" ~ "Colorado",
    doc_id == "preciv97" ~ "Texas",
    doc_id == "preciv99" ~ "Minnesota",
    doc_id == "preciv100" ~ "California",
    doc_id == "preciv101" ~ "California",
    doc_id == "preciv102" ~ "Kentucky",
    doc_id == "preciv103" ~ "California",
    doc_id == "preciv105" ~ "Minnesota",
    doc_id == "preciv107" ~ "Virginia",
    doc_id == "preciv109" ~ "Minnesota",
    doc_id == "preciv110" ~ "California",
    doc_id == "preciv112" ~ "Kentucky",
    doc_id == "preciv115" ~ "Delaware",
    doc_id == "preciv118" ~ "California",
    doc_id == "preciv119" ~ "Kentucky",
    doc_id == "preciv120" ~ "California",
    doc_id == "preciv121" ~ "California",
    doc_id == "preciv122" ~ "Missouri",
    doc_id == "preciv124" ~ "California",
    doc_id == "preciv126" ~ "Maryland",
    doc_id == "preciv132" ~ "California",
    doc_id == "preciv133" ~ "California",
    doc_id == "preciv134" ~ "Illinois",
    doc_id == "preciv136" ~ "California",
    doc_id == "preciv137" ~ "Delaware",
    doc_id == "preciv143" ~ "California",
    doc_id == "preciv146" ~ "California",
    doc_id == "preciv147" ~ "California",
    doc_id == "preciv151" ~ "California",
    doc_id == "preciv152" ~ "California",
    doc_id == "preciv155" ~ "Kansas",
    doc_id == "preciv157" ~ "California",
    doc_id == "preciv158" ~ "California",
    doc_id == "preciv159" ~ "California",
    doc_id == "preciv160" ~ "Iowa",
    doc_id == "preciv162" ~ "Wisconsin",
    doc_id == "preciv164" ~ "Louisiana",
    doc_id == "preciv166" ~ "Kansas",
    doc_id == "preciv167" ~ "Texas",
    doc_id == "preciv170" ~ "Kansas",
    doc_id == "preciv173" ~ "Nebraska",
    doc_id == "preciv174" ~ "Mississippi",
    doc_id == "preciv175" ~ "Kentucky",
    doc_id == "preciv177" ~ "California",
    doc_id == "preciv179" ~ "Kansas",
    doc_id == "preciv180" ~ "California",
    doc_id == "preciv181" ~ "Virginia",
    doc_id == "preciv184" ~ "Nebraska",
    doc_id == "preciv185" ~ "Kansas",
    doc_id == "preciv187" ~ "California",
    doc_id == "preciv189" ~ "California",
    doc_id == "preciv192" ~ "California",
    doc_id == "preciv193" ~ "Alabama",
    doc_id == "preciv195" ~ "Illinois",
    doc_id == "preciv196" ~ "District of Columbia",
    doc_id == "preciv198" ~ "California",
    doc_id == "preciv199" ~ "California",
    doc_id == "preciv202" ~ "California",
    doc_id == "preciv203" ~ "California",
    doc_id == "preciv205" ~ "Texas",
    doc_id == "preciv206" ~ "California",
    doc_id == "preciv207" ~ "Arkansas",
    doc_id == "preciv208" ~ "Virginia",
    doc_id == "preciv209" ~ "California",
    doc_id == "preciv210" ~ "California",
    doc_id == "preciv213" ~ "California",
    doc_id == "preciv215" ~ "California",
    doc_id == "preciv216" ~ "California",
    doc_id == "preciv217" ~ "California",
    doc_id == "preciv218" ~ "California",
    doc_id == "preciv219" ~ "California",
    doc_id == "preciv221" ~ "California",
    doc_id == "preciv223" ~ "California",
    doc_id == "preciv224" ~ "California",
    doc_id == "preciv226" ~ "Arkansas",
    doc_id == "preciv228" ~ "California",
    doc_id == "preciv229" ~ "Arkansas",
    doc_id == "preciv230" ~ "California",
    doc_id == "preciv231" ~ "California",
    doc_id == "preciv233" ~ "Kansas",
    doc_id == "preciv235" ~ "California",
    doc_id == "preciv237" ~ "California",
    doc_id == "preciv238" ~ "Texas",
    doc_id == "preciv239" ~ "Massachusetts",
    doc_id == "preciv240" ~ "Missouri",
    doc_id == "preciv243" ~ "California",
    doc_id == "preciv246" ~ "California",
    doc_id == "preciv247" ~ "Indiana",
    doc_id == "preciv248" ~ "Missouri",
    doc_id == "preciv250" ~ "Missouri",
    doc_id == "preciv253" ~ "California",
    doc_id == "preciv255" ~ "Missouri",
    doc_id == "preciv256" ~ "Missouri",
    doc_id == "preciv264" ~ "California",
    doc_id == "preciv266" ~ "California",
    doc_id == "preciv267" ~ "California",
    doc_id == "preciv268" ~ "California",
    doc_id == "preciv269" ~ "California",
    doc_id == "preciv270" ~ "California",
    doc_id == "preciv271" ~ "California",
    doc_id == "preciv273" ~ "California",
    doc_id == "preciv274" ~ "California",
    doc_id == "preciv276" ~ "Arkansas",
    doc_id == "preciv277" ~ "California",
    doc_id == "preciv279" ~ "Missouri",
    doc_id == "preciv280" ~ "California",
    doc_id == "preciv282" ~ "Missouri",
    doc_id == "preciv283" ~ "California",
    doc_id == "preciv284" ~ "Illinois",
    doc_id == "preciv287" ~ "Texas",
    doc_id == "preciv288" ~ "Illinois",
    doc_id == "preciv289" ~ "Tennessee",
    doc_id == "preciv292" ~ "Iowa",
    doc_id == "preciv293" ~ "California",
    doc_id == "preciv294" ~ "California",
    doc_id == "preciv297" ~ "California",
    doc_id == "preciv298" ~ "Louisiana",
    doc_id == "preciv300" ~ "California",
    doc_id == "preciv301" ~ "Louisiana",
    doc_id == "preciv302" ~ "California",
    doc_id == "preciv303" ~ "California",
    doc_id == "preciv304" ~ "New Mexico",
    doc_id == "preciv306" ~ "California",
    doc_id == "preciv308" ~ "California",
    doc_id == "preciv309" ~ "California",
    doc_id == "preciv310" ~ "Missouri",
    doc_id == "preciv312" ~ "California",
    doc_id == "preciv314" ~ "California",
    doc_id == "preciv317" ~ "California",
    doc_id == "preciv319" ~ "California",
    doc_id == "preciv321" ~ "California",
    doc_id == "preciv322" ~ "California",
    doc_id == "preciv323" ~ "California",
    doc_id == "preciv324" ~ "California",
    doc_id == "preciv325" ~ "California",
    doc_id == "preciv326" ~ "California",
    doc_id == "preciv327" ~ "California",
    doc_id == "preciv331" ~ "California",
    doc_id == "preciv333" ~ "Florida",
    doc_id == "preciv334" ~ "California",
    doc_id == "preciv337" ~ "California",
    doc_id == "preciv338" ~ "California",
    doc_id == "preciv339" ~ "Virginia",
    doc_id == "preciv342" ~ "California",
    doc_id == "preciv344" ~ "Texas",
    doc_id == "preciv349" ~ "California",
    doc_id == "preciv350" ~ "California",
    doc_id == "preciv351" ~ "California",
    doc_id == "preciv352" ~ "California",
    doc_id == "preciv353" ~ "District of Columbia",
    doc_id == "preciv354" ~ "California",
    doc_id == "preciv357" ~ "Florida",
    doc_id == "preciv359" ~ "California",
    doc_id == "preciv360" ~ "California",
    doc_id == "preciv361" ~ "California",
    doc_id == "preciv362" ~ "California",
    doc_id == "preciv363" ~ "California",
    doc_id == "preciv364" ~ "California",
    doc_id == "preciv366" ~ "California",
    doc_id == "preciv367" ~ "California",
    doc_id == "preciv368" ~ "California",
    doc_id == "preciv369" ~ "California",
    doc_id == "preciv370" ~ "California",
    doc_id == "preciv371" ~ "Kentucky",
    doc_id == "preciv373" ~ "Maryland",
    doc_id == "preciv375" ~ "Kentucky",
    doc_id == "preciv377" ~ "Texas",
    doc_id == "preciv378" ~ "Virginia",
    doc_id == "preciv381" ~ "California",
    doc_id == "preciv382" ~ "California",
    doc_id == "preciv383" ~ "Tennessee",
    doc_id == "preciv384" ~ "Mississippi",
    doc_id == "preciv385" ~ "Arkansas",
    doc_id == "preciv385" ~ "Illinois",
    doc_id == "preciv387" ~ "Arkansas",
    doc_id == "preciv387" ~ "New Jersey",
    doc_id == "preciv390" ~ "Illinois",
    doc_id == "preciv391" ~ "Texas",
    doc_id == "preciv392" ~ "Arkansas",
    doc_id == "preciv393" ~ "Illinois",
    doc_id == "preciv394" ~ "Illinois",
    doc_id == "preciv395" ~ "Arkansas",
    doc_id == "preciv396" ~ "Illinois",
    doc_id == "preciv397" ~ "Illinois",
    doc_id == "preciv398" ~ "Missouri",
    doc_id == "preciv399" ~ "New York",
    doc_id == "preciv400" ~ "New York",
    doc_id == "preciv402" ~ "Louisiana",
    doc_id == "preciv403" ~ "Florida",
    doc_id == "preciv404" ~ "Florida",
    doc_id == "preciv404" ~ "North Carolina",
    doc_id == "preciv408" ~ "Mississippi",
    doc_id == "preciv409" ~ "Maryland",
    doc_id == "preciv410" ~ "Pennsylvania",
    doc_id == "preciv412" ~ "Texas",
    doc_id == "preciv415" ~ "Illinois",
    doc_id == "preciv416" ~ "Illinois",
    doc_id == "preciv417" ~ "Mississippi",
    doc_id == "preciv418" ~ "Missouri",
    doc_id == "preciv419" ~ "Arkansas", 
    doc_id == "preciv420" ~ "Massachusetts",
    doc_id == "preciv422" ~ "Alabama", 
    doc_id == "preciv423" ~ "Missouri", 
    doc_id == "preciv424" ~ "Missouri", 
    doc_id == "preciv425" ~ "Mississippi", 
    doc_id == "preciv426" ~ "Arkansas",   
    doc_id == "preciv427" ~ "Kentucky",  
    doc_id == "preciv429" ~ "Mississippi", 
    doc_id == "preciv430" ~ "Mississippi", 
    doc_id == "preciv432" ~ "Mississippi",
    doc_id == "preciv435" ~ "Kentucky", 
    doc_id == "preciv439" ~ "Louisiana", 
    doc_id == "preciv440" ~ "Massachusetts", 
    doc_id == "preciv441" ~ "Alabama", 
    doc_id == "preciv442" ~ "Kentucky", 
    doc_id == "preciv443" ~ "Alabama", 
    doc_id == "preciv446" ~ "Mississippi",
    TRUE ~ state  # Keep existing state value if doc_id not in list
  ))




```


```{r}
docs_states <- x |> 
  select(doc_id, state)

Xprecivil_complete <- precivil_complete |> 
  left_join(docs_states, by=c("index"="doc_id"))

```


# 491 cases - deleted irrelevant entries
```{r}
x <- x |> 
  filter(!doc_id %in% c("preciv14", "preciv17", "preciv201", "preciv211", 
                        "preciv25", "preciv251", "preciv260", "preciv290", 
                        "preciv291", "preciv328", "preciv332", "preciv434", 
                        "preciv49", "preciv50", "preciv58", "preciv59", "preciv6"))



```


```{r}
new_total <- x |> 
  count(state) |> 
  arrange(desc(n))
#-- jan 5: 26 states, calif 41 reports. 147 reports total
# jan 6: 36 states, calif 135
# jan 7: 37 states, calif 140
```

```{r}
new_text_location <- new_location |> 
  right_join(precivil_df, by = c("doc_id" = "index")) |> 
  # For each doc_id, remove locations that are substrings of other locations
  group_by(doc_id) |>
  arrange(doc_id, desc(nchar(location_clean))) |>  # Longest first
  filter(!duplicated(location_clean)) |>  # Remove exact duplicates first
  mutate(
    # Check if this location is a substring of any longer location in same doc
    is_substring = sapply(location_clean, function(loc) {
      if (is.na(loc)) return(FALSE)
      others <- location_clean[location_clean != loc & !is.na(location_clean)]
      any(str_detect(others, fixed(loc)))
    })
  ) |>
  filter(!is_substring) |>  # Remove substrings
  select(-is_substring) 

# Now create the text file with locations highlighted
file_conn <- file("third_verify_states.txt", "w")

for (i in 1:nrow(new_text_location)) {
  row_data <- new_text_location[i, ]
  
  writeLines(paste0("Document ID: ", row_data$doc_id), file_conn)
  writeLines(paste0("Sentence ID: ", row_data$sentence_id), file_conn)
  writeLines(paste0("State: ", row_data$state), file_conn)
  writeLines(paste0("Location: ", row_data$location), file_conn)
  # writeLines(paste0("Entity Type: ", row_data$entity_types), file_conn)
  writeLines(paste0("Headline: ", row_data$headline), file_conn)
  writeLines(paste0("Newspaper: ", row_data$newspaper_name), file_conn)
  writeLines(paste0("Date: ", row_data$date), file_conn)

  
  # Wrap the full text
  wrapped_text <- strwrap(row_data$spellchecked, width = 80)
  wrapped_text[1] <- paste0("Full Context: ", wrapped_text[1])
  writeLines(wrapped_text, file_conn)
  
  if (i < nrow(new_text_location)) {
    writeLines("\n==================================================\n", file_conn)
  }
}

close(file_conn)
```

```{r}
new_text_location |> 
  count(doc_id) |> 
  arrange(desc(n))
```





```{r}
# See what's being filtered OUT
locations_filtered_out <- parsed %>%
  filter(str_detect(entity, "^(GPE|LOC|FAC)_(B|I)$")) %>%
  group_by(doc_id, sentence_id) %>%
  mutate(prev_token = lag(token)) %>%
  ungroup() %>%
  inner_join(lynch_info, by = c("doc_id", "sentence_id")) %>%
  # OPPOSITE of your filter
  filter(!(prev_token %in% c("at", "in", "near", "on", "from", "to", "near") |
           token_id > lynch_token_id))

# Look at examples
locations_filtered_out %>% 
  select(doc_id, token, prev_token, token_id, lynch_token_id) %>%
  head(20)
```

```{r}
locations_filtered_out %>%
  filter(token_id >= lynch_token_id - 5 & token_id < lynch_token_id) %>%
  nrow()
```

```{r}
parsed %>%
  filter(doc_id == "preciv69") %>%
  select(token_id, token,  entity)
```

```{r}
locations_with_prep %>%
  filter(doc_id == "preciv69", sentence_id == 1) %>%
  mutate(entity_group = cumsum(str_detect(entity, "_B$"))) %>%
  select(token_id, token, entity, entity_group)
```


##--------------------------------------------------------------------------
# Notes from original spacyr geographic parsing
##--------------------------------------------------------------------------


### Cities and towns
Captures lynching locations!
```{r}

# Get lynch sentences
lynch_info <- parsed %>%
   filter(grepl("lynch|hung|hang(ed|ing)?|executed?|execution|mob\\s+law|vigilance\\s+committee", 
               token, ignore.case = TRUE)) %>%
  distinct(doc_id, sentence_id, lynch_token_id = token_id)

# For each location entity, check what word comes before it
locations_with_prep <- parsed %>%
  filter(str_detect(entity, "^(GPE|LOC|FAC)_(B|I)$")) %>%
  # Add the previous token for each row
  group_by(doc_id, sentence_id) %>%
  mutate(prev_token = lag(token)) %>%
  ungroup() %>%
  # Join with lynch info
  inner_join(lynch_info, by = c("doc_id", "sentence_id")) %>%
  # old Filter for location prepositions and positions after lynch term
  # filter(prev_token %in% c("at", "in", "near", "on", "from", "to", "near") |
  #        token_id > lynch_token_id)
   # Filter for:
  # 1. Location prepositions, OR
  # 2. Within 5 tokens before lynch term, OR
  # 3. After lynch term
  filter(prev_token %in% c("at", "in", "near", "on", "from", "to") |
         (token_id >= lynch_token_id - 5 & token_id < lynch_token_id) |
         token_id > lynch_token_id)
```


### count cities

captures 226 places

```{r}

# Extract location entities and combine multi-word locations
location_counts2 <- locations_with_prep %>%
 # filter(str_detect(entity, "^(GPE|LOC|FAC)_(B|I)$")) %>%
  # Group by doc_id and sentence_id to keep entities together
  group_by(doc_id, sentence_id) %>%
  # Create entity groups - each _B starts a new entity
  mutate(entity_group = cumsum(str_detect(entity, "_B$"))) %>%
  # Combine tokens within each entity group
  group_by(doc_id, sentence_id, entity_group) %>%
  summarise(
    location = paste(token, collapse = " "),
    .groups = "drop"
  ) 


```

### Normalize list

```{r}
location_counts2 <- location_counts2 |> 
  mutate(location_clean = tolower(location)) |> 
  mutate(location_clean = case_when(
    # 1. REMOVE NOISE (Returns NA)
    # Added: san, driskells, los
    location_clean %in% c("a", "the memphis enquirer of the", "the yoakum gang", 
                          "rev", "lieut", "jas", "sheriff", "fort", "5th", 
                          "5th september september", "states", "west", "red",
                          "a senor senor badlo badlo", "san", "driskells", "los") ~ NA_character_,
    
    # 2. SPECIFIC COMBINATIONS & CORRECTIONS
    # Rogue River variants -> Rogue River Valley
    location_clean %in% c("rogue", "rogue river", "rogue river river", 
                          "rogue river river valley valley") ~ "Rogue River Valley",
    
    # Alameda -> Alameda County
    location_clean == "alameda" ~ "Alameda County",
    
    # Annador -> Amador County
    location_clean == "annador county" ~ "Amador County",
    
    # Gentry variants -> Gentry County
    location_clean %in% c("gentry", "gentry county county") ~ "Gentry County",
    
    # 3. MAPPINGS & ABBREVIATIONS
    location_clean %in% c("mo", "mo .", "mo . .") ~ "Missouri",
    location_clean %in% c("ky", "ky .")            ~ "Kentucky",
    location_clean %in% c("ill", "ill . .")        ~ "Illinois",
    location_clean %in% c("la .")                  ~ "Louisiana",
    location_clean %in% c("n.c")                   ~ "North Carolina",
    
    # 4. TYPO FIXES (Specific cases)
    location_clean == "san francisco francisco"    ~ "San Francisco",
    location_clean == "los angeles angeles"        ~ "Los Angeles",
    location_clean == "santa barbara barbara"      ~ "Santa Barbara",
    location_clean == "red bluffs bluffs"          ~ "Red Bluffs",
    location_clean == "san luis luis obispo obispo" ~ "San Luis Obispo",
    location_clean == "salmon river river"         ~ "Salmon River",
    location_clean == "sawyer 's 's bar bar"       ~ "Sawyer's Bar",
    
    # 5. DEFAULT: Standardize everything else to Title Case
    # This handles Alabama, Nashville, Iowa, etc., automatically.
    TRUE ~ str_to_title(location_clean)
  )) |> 
  filter(!is.na(location_clean)) |> 
  group_by(location_clean) |> 
  summarise(
    freq_total = n(), 
    articles_total = n_distinct(doc_id), 
    .groups = "drop"
  ) |> 
  arrange(desc(freq_total))
```


### Datatable of locations
```{r}

datatable(location_counts2,
          # Use HTML styling directly in the caption string
          caption = htmltools::tags$caption(
            style = 'caption-side: top; text-align: center; color: black; font-size: 20pt; font-weight: bold;',
            'Top locations for pre-Civil War lynchings'
          ),
          options = list(
            pageLength = 50,
            searching = FALSE, 
            lengthChange = FALSE
          ))
```

### States from localities

```{r}
location_counts2 <- location_counts2 |> 
  mutate(state = case_when(
    # Already state names - pass through
    location_clean %in% c("Alabama", "Arkansas", "California", "Delaware", "Florida", "Georgia", 
                          "Illinois", "Indiana", "Iowa", "Kansas", "Kentucky", 
                          "Louisiana", "Minnesota", "Mississippi", "Missouri", 
                          "Nebraska", "New Hampshire", "New York", "North Carolina", "Oregon",
                          "Ohio", "Tennessee", "Texas", "Virginia") ~ location_clean,
    
    # California counties and cities
    location_clean %in% c("Alameda", "Alameda County", "Annador County", "Calaveras County",
                          "Merced County", "Napa", "Oakland", "Placer County", 
                          "San Andreas", "San Antonio", "San Felipe", "San Francisco",
                          "San Gabriel", "San Isidro", "San Jose", "San Luis Obispo",
                           "Santa Barbara", "Shasta", "Solano County", 
                          "Stockton", "Tejon", "Tuolumne County", "Yuba County",
                          "Red Bluffs", "Drytown", "Iowa Hill", "Monterey",
                          "Sawyer's Bar", "South Sulphur", "Los Angeles", "Sacramento", "Yreka") ~ "California",
    
    # Texas counties and cities
    location_clean %in% c("Uvalde County", "El Paso", "San Antonio") ~ "Texas",
    
    # Missouri counties and cities
    location_clean %in% c("Buchanan County", "Gentry County") ~ "Missouri",
    
    # Illinois counties and cities
    location_clean %in% c("Ogle County", "Keithsburg", "Rockford") ~ "Illinois",
    
    # Indiana counties and cities
    location_clean %in% c("South Bend") ~ "Indiana",
    #deleted "Noble County" "Allen County",  - too many variations
    
    # Mississippi counties and cities
    location_clean %in% c("Issaquena County", "Panola", "Pontotoc County") ~ "Mississippi",
    
    # Iowa counties and cities
    location_clean %in% c("Black Hawk", "Tama County") ~ "Iowa",
    
    # Kentucky cities
    location_clean %in% c("Lexington", "Louisville") ~ "Kentucky",
    
    # Tennessee cities
    location_clean %in% c("Memphis", "Nashville") ~ "Tennessee",
    
    # Georgia cities
    location_clean %in% c("Savannah") ~ "Georgia",
    
    # Arkansas counties and cities
    location_clean %in% c("Pike County", "Tuskegee") ~ "Alabama",
    
    # Louisiana cities
    location_clean %in% c("Baton Rouge", "New Orleans") ~ "Louisiana",
    
    # Kansas cities
    location_clean %in% c("Lecompton", "Leavenworth City", "Kansas Territory") ~ "Kansas",
    
    # New York cities
    location_clean == "New York" ~ "New York",
    
    # Ohio cities
    location_clean == "Williamstown" ~ "Kentucky",
    
    # Florida cities
    location_clean == "Pensacola" ~ "Florida",
    
    # Nebraska cities
    location_clean == "Omaha City" ~ "Nebraska",
    
    # Colorado cities
    location_clean == "Denver City" ~ "Colorado",
    
    # Oregon counties
    location_clean %in% c("Lewis County", "Rogue River", "Rogue River Valley") ~ "Oregon",
    
    # North Carolina cities
    location_clean == "Wilmington" ~ "North Carolina",
    
    # District of Columbia
    location_clean == "Washington" ~ "District of Columbia",
    
    # Rivers/geographic features - assign to primary state
    location_clean == "Mississippi" ~ "Mississippi",
    location_clean == "Upper Mississippi River" ~ "Multiple States",
    location_clean == "Salmon River" ~ "Idaho",
    
    # Uncertain/generic locations
    location_clean %in% c("Columbia", "Geneva", "Milton", "Alston", "Bargar",
                          "Driskells", "Partridge", "Renick", "Scaggs", 
                          "Shattuck", "Weatherford", "Weston", "Woodstock",
                          "Shelbyville", "Lawrenceville", "Union City",
                          "Grant County", "Randolph County", "Petersburg", "Essex County", "Melbourne", "Mount Alexander") ~ "Unknown",
    
    
    # Default
    TRUE ~ "Unknown"
  ))

# View results
#unknown = 39
map <- 
location_counts2 |> 
  count(state) |> 
  arrange(desc(n)) |> 
  filter(state!="Unknown")

write_csv(map, "../output/map.csv")
  
```

The
whole neighborhood was aroused to a frenzy of excitement so
soon as the bloody deed became known. 


### Quick view of any entry by row
```{r}
wrapped_text <- strwrap(as.character(precivil_complete$spellchecked[426]), width = 60)
writeLines(wrapped_text, "temp_text.txt")
file.edit("temp_text.txt")
```


#***********************

## validate-read city entries


Add a cleaned state column to locations_prep and then join to precivil and then view
```{r}
Xlocations_with_context <- locations_with_prep %>%
  group_by(doc_id, sentence_id) %>%
  summarize(
    locations = paste(token, collapse = " "),
    entity_types = paste(unique(entity), collapse = ", "),
    .groups = "drop"
  ) %>%
  right_join(precivil_df, by = c("doc_id" = "index")) |> 
   mutate(location_clean = tolower(locations)) |> 
  mutate(location_clean = case_when(
    # 1. REMOVE NOISE (Returns NA)
    # Added: san, driskells, los
    location_clean %in% c("a", "the memphis enquirer of the", "the yoakum gang", 
                          "rev", "lieut", "jas", "sheriff", "fort", "5th", 
                          "5th september september", "states", "west", "red",
                          "a senor senor badlo badlo", "san", "driskells", "los") ~ NA_character_,
    
    # 2. SPECIFIC COMBINATIONS & CORRECTIONS
    # Rogue River variants -> Rogue River Valley
    location_clean %in% c("rogue", "rogue river", "rogue river river", 
                          "rogue river river valley valley") ~ "Rogue River Valley",
    
    # Alameda -> Alameda County
    location_clean == "alameda" ~ "Alameda County",
    
    # Annador -> Amador County
    location_clean == "annador county" ~ "Amador County",
    
    # Gentry variants -> Gentry County
    location_clean %in% c("gentry", "gentry county county") ~ "Gentry County",
    
    # 3. MAPPINGS & ABBREVIATIONS
    location_clean %in% c("mo", "mo .", "mo . .") ~ "Missouri",
    location_clean %in% c("ky", "ky .")            ~ "Kentucky",
    location_clean %in% c("ill", "ill . .")        ~ "Illinois",
    location_clean %in% c("la .")                  ~ "Louisiana",
    location_clean %in% c("n.c")                   ~ "North Carolina",
    
    # 4. TYPO FIXES (Specific cases)
    location_clean == "san francisco francisco"    ~ "San Francisco",
    location_clean == "los angeles angeles"        ~ "Los Angeles",
    location_clean == "santa barbara barbara"      ~ "Santa Barbara",
    location_clean == "red bluffs bluffs"          ~ "Red Bluffs",
    location_clean == "san luis luis obispo obispo" ~ "San Luis Obispo",
    location_clean == "salmon river river"         ~ "Salmon River",
    location_clean == "sawyer 's 's bar bar"       ~ "Sawyer's Bar",
    
    # 5. DEFAULT: Standardize everything else to Title Case
    # This handles Alabama, Nashville, Iowa, etc., automatically.
    TRUE ~ str_to_title(location_clean)
  )) |> 
### States from localities
  mutate(state = case_when(
    # Already state names - pass through
    location_clean %in% c("Alabama", "Arkansas", "California", "Delaware", "Florida", "Georgia", 
                          "Illinois", "Indiana", "Iowa", "Kansas", "Kentucky", 
                          "Louisiana", "Minnesota", "Mississippi", "Missouri", 
                          "Nebraska", "New Hampshire", "New York", "North Carolina", "Oregon",
                          "Ohio", "Tennessee", "Texas", "Virginia") ~ location_clean,
    
    # California counties and cities
    location_clean %in% c("Alameda", "Alameda County", "Annador County", "Calaveras County",
                          "Merced County", "Napa", "Oakland", "Placer County", 
                          "San Andreas", "San Antonio", "San Felipe", "San Francisco",
                          "San Gabriel", "San Isidro", "San Jose", "San Luis Obispo",
                           "Santa Barbara", "Shasta", "Solano County", 
                          "Stockton", "Tejon", "Tuolumne County", "Yuba County",
                          "Red Bluffs", "Drytown", "Iowa Hill", "Monterey",
                          "Sawyer's Bar", "South Sulphur", "Los Angeles", "Sacramento", "Yreka") ~ "California",
    
    # Texas counties and cities
    location_clean %in% c("Uvalde County", "El Paso", "San Antonio") ~ "Texas",
    
    # Missouri counties and cities
    location_clean %in% c("Buchanan County", "Gentry County") ~ "Missouri",
    
    # Illinois counties and cities
    location_clean %in% c("Ogle County", "Keithsburg", "Rockford") ~ "Illinois",
    
    # Indiana counties and cities
    location_clean %in% c("South Bend") ~ "Indiana",
    #deleted "Noble County" "Allen County",  - too many variations
    
    # Mississippi counties and cities
    location_clean %in% c("Issaquena County", "Panola", "Pontotoc County") ~ "Mississippi",
    
    # Iowa counties and cities
    location_clean %in% c("Black Hawk", "Tama County") ~ "Iowa",
    
    # Kentucky cities
    location_clean %in% c("Lexington", "Louisville") ~ "Kentucky",
    
    # Tennessee cities
    location_clean %in% c("Memphis", "Nashville") ~ "Tennessee",
    
    # Georgia cities
    location_clean %in% c("Savannah") ~ "Georgia",
    
    # Arkansas counties and cities
    location_clean %in% c("Pike County", "Tuskegee") ~ "Alabama",
    
    # Louisiana cities
    location_clean %in% c("Baton Rouge", "New Orleans") ~ "Louisiana",
    
    # Kansas cities
    location_clean %in% c("Lecompton", "Leavenworth City", "Kansas Territory") ~ "Kansas",
    
    # New York cities
    location_clean == "New York" ~ "New York",
    
    # Ohio cities
    location_clean == "Williamstown" ~ "Kentucky",
    
    # Florida cities
    location_clean == "Pensacola" ~ "Florida",
    
    # Nebraska cities
    location_clean == "Omaha City" ~ "Nebraska",
    
    # Colorado cities
    location_clean == "Denver City" ~ "Colorado",
    
    # Oregon counties
    location_clean %in% c("Lewis County", "Rogue River", "Rogue River Valley") ~ "Oregon",
    
    # North Carolina cities
    location_clean == "Wilmington" ~ "North Carolina",
    
    # District of Columbia
    location_clean == "Washington" ~ "District of Columbia",
    
    # Rivers/geographic features - assign to primary state
    location_clean == "Mississippi" ~ "Mississippi",
    location_clean == "Upper Mississippi River" ~ "Multiple States",
    location_clean == "Salmon River" ~ "Idaho",
    
    # Uncertain/generic locations
    location_clean %in% c("Columbia", "Geneva", "Milton", "Alston", "Bargar",
                          "Driskells", "Partridge", "Renick", "Scaggs", 
                          "Shattuck", "Weatherford", "Weston", "Woodstock",
                          "Shelbyville", "Lawrenceville", "Union City",
                          "Grant County", "Randolph County", "Petersburg", "Essex County", "Melbourne", "Mount Alexander") ~ "Unknown",
    
    
    # Default
    TRUE ~ "Unknown"
  ))
```



# Extract location entities and combine multi-word locations
locations_normalized <- locations_with_prep %>%
  filter(str_detect(entity, "^(GPE|LOC|FAC)_(B|I)$")) %>%
  # Group by doc_id and sentence_id to keep entities together
  group_by(doc_id, sentence_id) %>%
  # Create entity groups - each _B starts a new entity
  mutate(entity_group = cumsum(str_detect(entity, "_B$"))) %>%
  # Combine tokens within each entity group
  group_by(doc_id, sentence_id, entity_group) %>%
  summarise(
    location = paste(token, collapse = " "),
    .groups = "drop"
  ) 




x <- locations_with_prep |> 
  inner_join(precivil_complete, by=c("doc_id"="index")

```{r}
# Now create the text file with locations highlighted
file_conn <- file("verify_states.txt", "w")

for (i in 1:nrow(Xlocations_with_context)) {
  row_data <- Xlocations_with_context[i, ]
  
  writeLines(paste0("Document ID: ", row_data$doc_id), file_conn)
  writeLines(paste0("Sentence ID: ", row_data$sentence_id), file_conn)
  writeLines(paste0("State: ", row_data$state), file_conn)
  writeLines(paste0("Location: ", row_data$locations), file_conn)
  writeLines(paste0("Entity Type: ", row_data$entity_types), file_conn)
  writeLines(paste0("Headline: ", row_data$headline), file_conn)
  writeLines(paste0("Newspaper: ", row_data$newspaper_name), file_conn)
  writeLines(paste0("Date: ", row_data$date), file_conn)

  
  # Wrap the full text
  wrapped_text <- strwrap(row_data$kwic_text, width = 80)
  wrapped_text[1] <- paste0("Full Context: ", wrapped_text[1])
  writeLines(wrapped_text, file_conn)
  
  if (i < nrow(Xlocations_with_context)) {
    writeLines("\n==================================================\n", file_conn)
  }
}

close(file_conn)
```

#**********************

```{r}
# First, merge locations with your original dataframe
# Assuming your original docs have a doc_id that matches locations3$doc_id
precivil_df <- df_llm
#filter here if needed


locations_with_context <- locations_with_prep %>%
  group_by(doc_id, sentence_id) %>%
  summarize(
    locations = paste(token, collapse = " "),
    entity_types = paste(unique(entity), collapse = ", "),
    .groups = "drop"
  ) %>%
  right_join(precivil_df, by = c("doc_id" = "index")) # adjust join key as needed

# Now create the text file with locations highlighted
file_conn <- file("lynching_locations.txt", "w")

for (i in 1:nrow(locations_with_context)) {
  row_data <- locations_with_context[i, ]
  
  writeLines(paste0("Document ID: ", row_data$doc_id), file_conn)
  writeLines(paste0("Sentence ID: ", row_data$sentence_id), file_conn)
  writeLines(paste0("LOCATION IDENTIFIED: ", row_data$locations), file_conn)
  writeLines(paste0("Entity Type: ", row_data$entity_types), file_conn)
  writeLines(paste0("Headline: ", row_data$headline), file_conn)
  writeLines(paste0("Newspaper: ", row_data$newspaper_name), file_conn)
  writeLines(paste0("Date: ", row_data$date), file_conn)

  
  # Wrap the full text
  wrapped_text <- strwrap(row_data$kwic_text, width = 80)
  wrapped_text[1] <- paste0("Full Context: ", wrapped_text[1])
  writeLines(wrapped_text, file_conn)
  
  if (i < nrow(locations_with_context)) {
    writeLines("\n==================================================\n", file_conn)
  }
}

close(file_conn)
```



### Print stories for validation
precivil narrow print specific stories in a text file
```{r}
#filter data to examine
precivil_df <- precivil_complete |> 
  filter(index=="preciv446")


# Open a connection to the text file
file_conn <- file("dataframe_entries.txt", "w")

# Loop through each row of the dataframe
for (i in 1:nrow(precivil_df)) {
  # Extract the row
  row_data <- precivil_df[i, c("headline", "newspaper_name", "date", "spellchecked", "article_id", "index")]
  
  # Write each field on a separate line
  writeLines(paste0("Headline: ", row_data$headline), file_conn)
  writeLines(paste0("Newspaper: ", row_data$newspaper_name), file_conn)
  writeLines(paste0("Date: ", row_data$date), file_conn)
  writeLines(paste0("Article ID: ", row_data$article_id), file_conn)
  writeLines(paste0("Index Number: ", row_data$index), file_conn)
   wrapped_explanation <- strwrap(row_data$spellchecked, width = 60)
  wrapped_explanation[1] <- paste0("Spellchecked Text: ", wrapped_explanation[1])
  writeLines(wrapped_explanation, file_conn)
  
  # # Handle the matches list column
  # matches_content <- unlist(row_data$matches)
  # if (length(matches_content) > 0) {
  #   writeLines(paste0("Matches: ", paste(matches_content, collapse = ", ")), file_conn)
  # } else {
  #   writeLines("Matches: None", file_conn)
  # }
  
  # writeLines(paste0("\nArticle:\n", row_data$spellchecked), file_conn)
  
  # Add separator line (except after the last entry)
  if (i < nrow(precivil_df)) {
    writeLines("\n**************************************************\n", file_conn)
  }
}

# Close the file connection
close(file_conn)
```


## print from df_llm
```{r}

#filter data to examine
precivil_df <- df_llm |> 
  filter(str_detect(kwic_text, "California"))


# Open a connection to the text file
file_conn <- file("dataframe_entries.txt", "w")

# Loop through each row of the dataframe
for (i in 1:nrow(precivil_df)) {
  # Extract the row
  row_data <- precivil_df[i, c("headline", "newspaper_name", "date", "kwic_text", "index")]
  
  # Write each field on a separate line
  writeLines(paste0("Headline: ", row_data$headline), file_conn)
  writeLines(paste0("Newspaper: ", row_data$newspaper_name), file_conn)
  writeLines(paste0("Date: ", row_data$date), file_conn)
  writeLines(paste0("Index Number: ", row_data$index), file_conn)
   wrapped_explanation <- strwrap(row_data$kwic_text, width = 60)
  wrapped_explanation[1] <- paste0("KWIC_text: ", wrapped_explanation[1])
  writeLines(wrapped_explanation, file_conn)
  
  if (i < nrow(precivil_df)) {
    writeLines("\n**************************************************\n", file_conn)
  }
}

# Close the file connection
close(file_conn)
```


#-------------------------------------------------------------------------------

#Notes

# missing 15 text files from students
```{r}
missing <- students |> 
  anti_join(files, by=c("file_id_txt"="filename")) 

write_csv(missing, "missing_student_text_dec_21.csv")
```


### Notes on parsing
https://claude.ai/share/9c186ac7-2e5f-4e94-a64d-14e1e79aa49e

using spacyr- needs work
```{r}


# Get sentence-level text
sentences_with_locations <- parsed %>%  # your spacyr parsed object
  filter(doc_id %in% locations3$doc_id,
         sentence_id %in% locations3$sentence_id) %>%
  group_by(doc_id, sentence_id) %>%
  summarize(
    sentence_text = paste(token, collapse = " "),
    .groups = "drop"
  ) %>%
  left_join(locations3 %>% 
              group_by(doc_id, sentence_id) %>%
              summarize(locations = paste(token, collapse = " "), .groups = "drop"))

# Join with your metadata
results <- sentences_with_locations %>%
  left_join(precivil_df, by = c("doc_id" = "index"))

# Write to file
file_conn <- file("lynching_locations_sentences.txt", "w")

for (i in 1:nrow(results)) {
  row <- results[i, ]
  
  writeLines(paste0("=== ENTRY ", i, " ==="), file_conn)
  writeLines(paste0("Location Found: ", row$locations), file_conn)
  writeLines(paste0("Document: ", row$doc_id), file_conn)
  writeLines(paste0("Newspaper: ", row$newspaper_name, " (", row$date, ")"), file_conn)
  writeLines("", file_conn)
  writeLines("Sentence with location:", file_conn)
  wrapped <- strwrap(row$sentence_text, width = 80)
  writeLines(paste0("  ", wrapped), file_conn)
  writeLines("", file_conn)
  writeLines("Full article context:", file_conn)
  wrapped_full <- strwrap(row$kwic_text, width = 80)
  writeLines(paste0("  ", wrapped_full), file_conn)
  writeLines("\n**************************************************\n", file_conn)
}

close(file_conn)
```



## people
```{r}
# For people
person_counts <- annotated_spacy %>%
  filter(str_detect(entity, "^PERSON_(B|I)$")) %>%
  group_by(doc_id, sentence_id) %>%
  mutate(entity_group = cumsum(str_detect(entity, "_B$"))) %>%
  group_by(doc_id, sentence_id, entity_group) %>%
  summarise(
    person = paste(token, collapse = " "),
    .groups = "drop"
  ) %>%
  group_by(person) %>%
  summarise(
    frequency = n(),
    articles = n_distinct(doc_id)
  ) %>%
  arrange(desc(frequency))

# For organizations
org_counts <- annotated_spacy %>%
  filter(str_detect(entity, "^ORG_(B|I)$")) %>%
  group_by(doc_id, sentence_id) %>%
  mutate(entity_group = cumsum(str_detect(entity, "_B$"))) %>%
  group_by(doc_id, sentence_id, entity_group) %>%
  summarise(
    organization = paste(token, collapse = " "),
    .groups = "drop"
  ) %>%
  group_by(organization) %>%
  summarise(
    frequency = n(),
    articles = n_distinct(doc_id)
  ) %>%
  arrange(desc(frequency))
```


```{r}

library(stringr)

# Extract location entities and combine multi-word locations
location_counts2 <- annotated_spacy %>%
  filter(str_detect(entity, "^(GPE|LOC|FAC)_(B|I)$")) %>%
  # Group by doc_id and sentence_id to keep entities together
  group_by(doc_id, sentence_id) %>%
  # Create entity groups - each _B starts a new entity
  mutate(entity_group = cumsum(str_detect(entity, "_B$"))) %>%
  # Combine tokens within each entity group
  group_by(doc_id, sentence_id, entity_group) %>%
  summarise(
    location = paste(token, collapse = " "),
    .groups = "drop"
  ) %>%
  # Now count across all documents
  group_by(location) %>%
  summarise(
    frequency = n(),
    articles = n_distinct(doc_id)
  ) %>%
  arrange(desc(frequency)) |> 
  mutate(pct =  formattable::percent(articles/508,1))

head(location_counts, 20)
```

```{r}
# Get doc_id + sentence_id combinations that contain lynch terms
lynch_sentences <- parsed %>%
  filter(grepl("lynch|hung", token, ignore.case = TRUE)) %>%
  distinct(doc_id, sentence_id)

# Filter locations to those sentences
locations3 <- parsed %>%
  filter(str_detect(entity, "^(GPE|LOC)_(B|I)$")) %>%
  semi_join(lynch_sentences, by = c("doc_id", "sentence_id"))
```


Notes
These are BIO tags (Beginning-Inside-Outside) used in Named Entity Recognition to mark entity boundaries:
Entity Types:

GPE = Geopolitical Entity - locations with governments (cities, states, countries)

Examples: "San Francisco", "California", "United States", "Mexico"


LOC = Location - geographical locations that aren't geopolitical entities

Examples: mountain ranges, bodies of water, regions without government
Examples: "Pacific Ocean", "Sierra Nevada", "the West"



BIO Tags:

B (Beginning) = First token of an entity

GPE_B: 703 entities that start with a geopolitical name
LOC_B: 73 entities that start with a geographical location


I (Inside) = Continuation token of a multi-word entity

GPE_I: 229 tokens that are part of multi-word place names (e.g., "San" is B, "Francisco" is I)
LOC_I: 52 tokens continuing location name



## count cities - tighter filter
```{r}
#text <- df_llm$kwic_text
text <- data.frame(
  doc_id = df_llm$index,
  text = df_llm$kwic_text,
  stringsAsFactors = FALSE
)

# Parse with spacyr
parsed <- spacy_parse(text, entity = TRUE, 
                               lemma = TRUE,
                               pos = TRUE,
                               tag= TRUE)



# Find entities near "lynch" related words
lynch_positions <- which(grepl("lynch|hung", parsed$token, ignore.case = TRUE))

# Look for GPE (geopolitical entities) or LOC near these positions
window_size <- 5  # tokens before/after

# Create a helper to check if position is within window
is_near_lynch <- function(token_id, doc_id, parsed, lynch_positions, window_size) {
  # Get lynch positions in the same document
  doc_lynch_pos <- lynch_positions[parsed$doc_id[lynch_positions] == doc_id]
  
  if(length(doc_lynch_pos) == 0) return(FALSE)
  
  # Get actual token_id values for lynch positions in this doc
  lynch_token_ids <- parsed$token_id[doc_lynch_pos]
  
  # Check if current token_id is within window of any lynch position
  any(abs(token_id - lynch_token_ids) <= window_size)
}

locations3 <- parsed %>%
  filter(str_detect(entity, "^(GPE|LOC)_(B|I)$")) %>%
  filter(sentence_id %in% parsed$sentence_id[lynch_positions])


# Extract location entities and combine multi-word locations
location_counts3 <- parsed %>%
  filter(str_detect(entity, "^(GPE|LOC)_(B|I)$")) %>%
  filter(sentence_id %in% parsed$sentence_id[lynch_positions]) |> 
  # Group by doc_id and sentence_id to keep entities together
  group_by(doc_id, sentence_id) %>%
  # Create entity groups - each _B starts a new entity
  mutate(entity_group = cumsum(str_detect(entity, "_B$"))) %>%
  # Combine tokens within each entity group
  group_by(doc_id, sentence_id, entity_group) %>%
  summarise(
    location = paste(token, collapse = " "),
    .groups = "drop"
  ) %>%
  # Now count across all documents
  group_by(location) %>%
  summarise(
    frequency = n(),
    articles = n_distinct(doc_id)
  ) %>%
  arrange(desc(frequency)) |> 
  mutate(pct =  formattable::percent(articles/508,1))


```
```{r}
# Extract locations (GPE = geopolitical entity, LOC = location)
location_counts2 <- annotated_spacy %>%
    filter(entity %in% c("GPE_B", "GPE_I", "LOC_B", "LOC_I")) %>%
  group_by(token) %>%
  summarise(
    frequency = n(),
    articles = n_distinct(doc_id)
  ) %>%
  arrange(desc(frequency))
```

```{r}
# If you have entity information in your annotated_df
location_counts <- annotated_df %>%
  filter(upos == "PROPN") %>%
  # You can also filter by entity type if available
  filter(entity %in% c("LOC", "GPE")) %>%  # LOC = location, GPE = geopolitical entity
  group_by(token) %>%
  summarise(
    frequency = n(),
    articles = n_distinct(doc_id)
  ) %>%
  arrange(desc(frequency))
```

```{r}


# Then extract locations using entity information
location_counts <- annotated_df %>%
  filter(
    upos == "PROPN" | 
    str_detect(entity, "^B-LOC|^I-LOC|^B-GPE|^I-GPE")  # Named entity tags for locations
  ) %>%
  group_by(token) %>%
  summarise(
    frequency = n(),
    articles = n_distinct(doc_id)
  ) %>%
  arrange(desc(frequency))
```



#-------------------------------------------------------------------------------


# Tolnay and Beck Lynching Inventory
```{r}

tolnay_beck <- read_csv("../data/Bailey_Beck_lynching_list_8_1_2022.csv") %>% 
  as.data.frame()

tolnay_beck <- janitor::clean_names(tolnay_beck)
```


```{r}
tolnay_beck |> 
  count(status_clean)
```


# Filter the dataframe: 9937 articles
filtered_df <- df1 %>%
